### 2024-03-01 appendix/index-definition.qmd
add definition to indexes in the appendix

---
title: "index"
format: pdf
---

For data $X_{n \times p}$, the projection basis $A_{p * d}$ gives the projected data, $y = XA$. Projection pursuit finds the projection direction $A$ that maximises the index function $f$: $$ \max f(XA) ~~ A^\prime A = 1.$$

The following sections summarise the index functions used in the simulation:

# hole index

-   smooth
-   formula: $$I_{holes}(A) = \frac{1-1/n \sum^n_{i = 1} \exp(-1/2y_iy_i^\prime)}{1-\exp(-p/2)}$$
-   reference:
    -   Cook and Swayne (2007) Interactive and Dynamic Graphics for Data Analysis page 30: [link](https://link.springer.com/chapter/10.1007/978-0-387-71762-3_2)
    -   code from the [tourr package](https://github.com/ggobi/tourr/blob/master/R/interesting-indices.r)

# Stringy

-   non-smooth

-   formula: $$I_{stringy}(A) = \frac{\text{number of vertices with 2 edges}}{\text{number of total vertices with more than one edge}} $$ (probably need some graph theory notation to write it mathematically)

-   reference:

    -   Wilkinson et al (2005) Graph-Theoretic Scagnostics: page 160: [link](https://ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=1532142)

    -   code from the [cassowaryr page](https://github.com/numbats/cassowaryr/blob/HEAD/R/mst.R)

# spline

-   smooth
-   formula:
    -   $\hat{y}$ is derived from fitting a penalised spline model of y \~ x,
    -   $\hat{y}^\prime$ is derived from fitting a penalised spline model of x \~ y,
    -   the variance of the residuals is compared to the variance of the original data

$$I_{spline}(A) = \max \left(1 - \frac{var(y - \hat{y})}{var(x)}, 1 - \frac{var(y - \hat{y}^\prime)}{var(y)}\right)$$

-   reference:

    -   Ursula and Di (2020) Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics, page 1176: [link](https://arxiv.org/abs/1902.00181)
    -   code from the [cassowaryr page](https://github.com/numbats/cassowaryr/blob/HEAD/R/association.R)