### 2024-03-18 appendix/index-definition.qmd
dcor2d_2 index definition

---
title: "index"
format: pdf
---

For data $X_{n \times p}$, the projection basis $A_{p * d}$ gives the projected data, $y = XA$. Projection pursuit finds the projection direction $A$ that maximises the index function $f$: $$ \max f(XA) ~~ A^\prime A = 1.$$

For 2D projections, we can also write $y_{n \times 2} = (y_1, y_2)$, where $y_1$ and $y_2$ are the two columns of the projected data.

The following sections summarise the index functions used in the simulation:

# hole index

-   smooth
-   formula: $$I_{holes}(A) = \frac{1-1/n \sum^n_{i = 1} \exp(-1/2y_iy_i^\prime)}{1-\exp(-p/2)}$$
-   reference:
    -   Cook and Swayne (2007) Interactive and Dynamic Graphics for Data Analysis page 30: [link](https://link.springer.com/chapter/10.1007/978-0-387-71762-3_2)
    -   code from the [tourr package](https://github.com/ggobi/tourr/blob/master/R/interesting-indices.r)

# dcor2d_2 index

-   smooth

-   formula:

    -   define pair-wised distance: $a_{ij} = \|y_1^i  - y_1^j\|$ and $b_{ij} = \|y_2^i  - y_2^j\|$ for $i, j = 1, 2, \ldots, n$.

    -   define the column sum, row sum, and overall sum: $$a_{i \cdot} = \sum_{l = 1}^n a_{il}, \quad a_{\cdot j} = \sum_{k = 1}^n a_{kj}, \quad a_{\cdot \cdot} = \sum_{k,l = 1}^n a_{kl} \quad b_{i \cdot} = \sum_{l = 1}^n b_{il}, \quad b_{\cdot j} = \sum_{k = 1}^n b_{kj}, \quad b_{\cdot \cdot} = \sum_{k,l = 1}^n b_{kl}$$

    -   define the distance covariance and variance: $$\text{dCov}(a_{ij}, b_{ij}) = \frac{1}{n(n-3)} \sum_{i \neq j} a_{ij} b_{ij} -\frac{2}{n (n-2)(n-3)}\sum_{i = 1}^n a_{i \cdot} b_{i \cdot} + \frac{1}{n(n - 1)(n-2)(n-3)} a_{\cdot \cdot} b_{\cdot \cdot}$$

    -   define the distance correlation: $$\text{dCor} (a_{ij}, b_{ij}) = \frac{\text{dCov}(a_{ij}, b_{ij})}{\sqrt{\text{dVar}(a_{ij}) \text{dVar}(b_{ij})}}$$

-   Reference:

    -   Xiaoming Huo &Gábor J. Székely (2016) Fast Computing for Distance Covariance: [link](https://www.tandfonline.com/doi/full/10.1080/00401706.2015.1054435). The papre proves that the covariance formula above (eq 3.3 in the paper) is an unbaised estimator of the population covariance in the original formula: $\displaystyle \operatorname {dCov} _{n}^{2}(X,Y):={\frac {1}{n^{2}}}\sum _{j=1}^{n}\sum _{k=1}^{n}A_{j,k}\,B_{j,k}$. The fast algorithm realises the new formula is in $O(n \log n)$ time.
    -   Maria L. Rizzo, Gábor J. Székely (2015) Energy distance: [link](https://wires.onlinelibrary.wiley.com/doi/abs/10.1002/wics.1375) WIREs Computational Statistics review paper
    -   Gábor J. Székely, Maria L. Rizzo, Nail K. Bakirov (2007) Measuring and testing dependence by correlation of distances: [link](https://projecteuclid.org/journals/annals-of-statistics/volume-35/issue-6/Measuring-and-testing-dependence-by-correlation-of-distances/10.1214/009053607000000505.full): original paper of distance correlation

# MIC/ TIC index

-   smooth

-   formula

    -   Let $g(k, l)$ define a partition of the space $(y_1, y_2)$ into $k \times l$ rectangles, for example, $g(2, 3)$ means dividing the data space into 2 rectangles in the $y_2$ direction and 3 rectangles in the $y_1$ direction. Let $G$ denotes all the possible partition.
    -   MIC finds the maximum mutual information over G where k and l is bounded by the grid size: $k \times l < B(n) = n^{\alpha}$ where alpha = 0.3 in our simulation

    $$I_{MIC}(A) = \max_{g \in G} \frac{I(y_1, y_2 | g)}{\log(\min(k^*, l^*))}  = \max_{g \in G} \frac{\sum_{y1} \sum_{y2} P(y_1, y_2) \log \frac{P(y_1)}{P(y_1) P(y_2)}}{\log(\min(k^*, l^*))}$$

    where $k^*$ and $l^*$ are the number of rectangles in the optimal partition $g$.

    -   TIC calculates the sum of mutual information over G

$$I_{TIC}(A) = \sum_{g \in G} \frac{I(y_1, y_2 | g)}{\log(\min(k^*, l^*))}  = \sum_{g \in G} \frac{\sum_{y1} \sum_{y2} P(y_1, y_2) \log \frac{P(y_1)}{P(y_1) P(y_2)}}{\log(\min(k^*, l^*))}$$

-   Reference:

    -   Reshef (2011) Detecting Novel Associations in Large Datasets: [link](https://pubmed.ncbi.nlm.nih.gov/22174245/) - Figure 1 is usually to understand the general idea

    -   Reshef (2016) Measuring Dependence Powerfully and Equitably: [link](https://jmlr.csail.mit.edu/papers/volume17/15-308/15-308.pdf).

-   Note:

    -   The original Definition 1 on page 6 of Reshef (2016) uses supremum, here I just use the maximum.
    -   There are different version of MIC and we use $MIC_e$ and $TIC_e$ (See Sec 4 in the paper in Reshef (2016)), which uses the equicharacteristic matrix rather than the original characteristic matrix in both MIC/ TIC calculation, see page 14 Figure 1 for there difference (equicharacteristic matrix parts the space equally hence faster).
    -   Section 4.3/4.4 of Reshef (2016) talks about the time complexity

# loess/ spline index

-   smooth

-   formula:

    -   let $e^{\text{model}}_{y_1 \sim y_2}$ denote the residual from model $y_1 \sim y_2$

    -   for the loess index, we fit a loess model to $y_1 \sim y_2$ (use $\alpha = 0.05$ for the smoothing parameter in the loess index)

    $$I_{\text{loess}}(A) = \max \left(1 - \frac{var(e^{\text{loess}}_{y_1 \sim y_2})}{var(y_1)}, 1 - \frac{var(e^{\text{loess}}_{y_2 \sim y_1})}{var(y_2)}\right)$$

    -   for the spline index, a spline model is fitted (use cubic regression spline with k = 15 for the diemnsion of the basis in the spline model).

$$I_{\text{spline}}(A) = \max \left(1 - \frac{var(e^{\text{spline}}_{y_1 \sim y_2})}{var(y_1)}, 1 - \frac{var(e^{\text{spline}}_{y_2 \sim y_1})}{var(y_2)}\right)$$

-   reference:
    -   loess: (none) I create it myself, inspired by the spline index
    -   spline:
        -   Ursula and Di (2020) Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics, page 1176: [link](https://arxiv.org/abs/1902.00181)
        -   code from the [cassowaryr page](https://github.com/numbats/cassowaryr/blob/HEAD/R/association.R)

# Stringy

-   non-smooth

-   formula: $$I_{stringy}(A) = \frac{\text{number of vertices with 2 edges}}{\text{number of total vertices with more than one edge}} $$ (probably need some graph theory notation to write it mathematically)

-   reference:

    -   Wilkinson et al (2005) Graph-Theoretic Scagnostics: page 160: [link](https://ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=1532142)

    -   code from the [cassowaryr page](https://github.com/numbats/cassowaryr/blob/HEAD/R/mst.R)