### 2024-06-23 appendix/appendix.qmd
update appendix

---
title: 'Appendix to "Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit"'
author:
  - name: H. Sherry Zhang
    email: huize.zhang@austin.utexas.edu
    affiliations: 
        - id: 1
          name: University of Texas at Austin
          department: Department of Statistics and Data Sciences
          city: Austin
          country: United States
          postal-code: 78751
    attributes:
        corresponding: true
  - name: Dianne Cook
    email: dicook@monash.edu
    affiliations:
        - id: 2
          name: Monash University
          department: Department of Econometrics and Business Statistics
          city: Melbourne
          country: Australia
          postal-code: 3800
  - name: Nicolas  Langrené
    email: nicolaslangrene@uic.edu.cn
    affiliations:
        - id: 3
          name: Guangdong Provincial Key Laboratory of Interdisciplinary Research and Application for Data Science, BNU-HKBU United International College
          department: Department of Mathematical Sciences
          city: Zhuhai
          country: China
          postal-code: 519087
  - name: Jessica Wai Yin Leung
    email: Jessica.Leung@monash.edu
    affiliations:
        - id: 2
          name: Monash University
          department: Department of Econometrics and Business Statistics
          city: Melbourne
          country: Australia
          postal-code: 3800
date: last-modified
abstract: |
  Projection pursuit (PP) is a dimension reduction technique that identifies low-dimensional projections in high-dimensional data by optimising a criteria function known as the PP index. The optimisation in PP can be non-smooth and requires identifying optima with a small "squint angle", detectable only from close proximity. To address these challenges, this study investigates the performance of a recent swarm-based algorithm, Jellyfish Search Optimiser (JSO), for optimising PP indexes. The performance of JSO is evaluated across various hyper-parameter settings and compared with existing optimisers. Additionally, this work proposes novel methods to quantify two properties of the PP index -- smoothness and squintability -- that capture the complexities inherent in PP optimisation problems. These two metrics are evaluated with JSO hyper-parameters to determine their effect on JSO success rate. The JSO algorithm has been implemented in the `tourr` package, while calculations for smoothness and squintability are available in the `ferrn` package. 
format:
  elsevier-pdf:
    journal:
      name: Journal of Multivariate Analysis
      formatting: preprint
      model: 3p
      cite-style: numbername
bibliography: ../bibliography.bib
---

Given high-dimensional data $X \in \mathbf{R}^{n\times p}$ and the index function $f(\cdot)$, projection pursuit finds the orthonormal projection basis $A \in \mathbf{R}^{p \times d}$ by solving the following optimisation problem:

$$
\underset{A}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I
$$ {#eq-optimization}

This appendix presents definition of all indexes $f(\cdot)$ used in the paper. These indexes are defined in 2D cases ($d = 2$), where $Y \in \mathbf{R}^{n \times 2} = (y_1, y_2)$ represents the projected data.

## The `holes` index

The `holes` index [@cook1993projection] is a smooth index for detecting the presence of multi-modality in the projection. The index is defined as:

$$I_{\text{holes}}(A) = \frac{1-\frac{1}{n} \sum^n_{i = 1} \exp(-\frac{1}{2y_iy_i^\prime})}{1-\exp(-\frac{d}{2})}$$

## The `dcor2d` index

The `dcor2d` index [@Grimm2016] is a smooth index that measures the distance correlation between the two projection axes. The index uses pair-wise distance to define column sum, row sum, and overall sum and then calculates distance correlation from distance variance and covariance.

  -  pair-wised distance: $a_{ij} = \|y_1^i - y_1^j\|$ and $b_{ij} = \|y_2^i - y_2^j\|$ for $i, j = 1, 2, \ldots, n$

   -  column sum, row sum, and overall sum: $$a_{i \cdot} = \sum_{l = 1}^n a_{il}, \quad a_{\cdot j} = \sum_{k = 1}^n a_{kj}, \quad a_{\cdot \cdot} = \sum_{k,l = 1}^n a_{kl} \quad b_{i \cdot} = \sum_{l = 1}^n b_{il}, \quad b_{\cdot j} = \sum_{k = 1}^n b_{kj}, \quad b_{\cdot \cdot} = \sum_{k,l = 1}^n b_{kl}$$

  -   distance covariance (and variance defined similarly): $$\text{dCov}(a_{ij}, b_{ij}) = \frac{1}{n(n-3)} \sum_{i \neq j} a_{ij} b_{ij} -\frac{2}{n (n-2)(n-3)}\sum_{i = 1}^n a_{i \cdot} b_{i \cdot} + \frac{1}{n(n - 1)(n-2)(n-3)} a_{\cdot \cdot} b_{\cdot \cdot}$$

   -   distance correlation: $$I_{\text{dcor}}(A)  = \text{dCor} (a_{ij}, b_{ij}) = \frac{\text{dCov}(a_{ij}, b_{ij})}{\sqrt{\text{dVar}(a_{ij}) \text{dVar}(b_{ij})}}$$

# The `MIC` and `TIC` index

Both `MIC` and `TIC` are information-based indexes derived from mutual information. The indexes are defined based on a partition, $g(k, l)$, of the space $(y_1, y_2)$ into $k \times l$ rectangles, For example, $g(2, 3)$ divides the data space into 2 rectangles in the $y_2$ direction and 3 rectangles in the $y_1$ direction. Let $G$ denotes all the possible partition.

MIC finds the **maximum** mutual information over G where $k$ and $l$ are bounded by the grid size: $k \times l < B(n) = n^{\alpha}$. $alpha = 0.3$ is used in our simulation.

$$I_{MIC}(A) = \max_{g \in G} \frac{I(y_1, y_2 | g)}{\log(\min(k^*, l^*))}  = \max_{g \in G} \frac{\sum_{y1} \sum_{y2} P(y_1, y_2) \log \frac{P(y_1)}{P(y_1) P(y_2)}}{\log(\min(k^*, l^*))}$$

where $k^*$ and $l^*$ are the number of rectangles in the optimal partition $g$.

TIC calculates the **sum** of mutual information over G

$$I_{TIC}(A) = \sum_{g \in G} \frac{I(y_1, y_2 | g)}{\log(\min(k^*, l^*))}  = \sum_{g \in G} \frac{\sum_{y1} \sum_{y2} P(y_1, y_2) \log \frac{P(y_1)}{P(y_1) P(y_2)}}{\log(\min(k^*, l^*))}$$

# The `loess` and `splines` index

The `loess` and `splines` indexes detect non-linear structure in the projection, as captured by their respective model. These indexes are computed by regressing both axes against each other, and find the maximum variance from the residual that can be explained by the model. Let $e^{\text{model}}_{y_1 \sim y_2}$ denote the residual from the loess/splines model $y_1 \sim y_2$, and $e^{\text{model}}_{y_2 \sim y_1}$ denote the residual from the loess/spliens model $y_2 \sim y_1$. The two indexes are calculated as

 $$I_{\text{loess}}(A) = \max \left(1 - \frac{var(e^{\text{loess}}_{y_1 \sim y_2})}{var(y_1)}, 1 - \frac{var(e^{\text{loess}}_{y_2 \sim y_1})}{var(y_2)}\right)$$

$$I_{\text{spline}}(A) = \max \left(1 - \frac{var(e^{\text{spline}}_{y_1 \sim y_2})}{var(y_1)}, 1 - \frac{var(e^{\text{spline}}_{y_2 \sim y_1})}{var(y_2)}\right)$$

# The `stringy` index

The `stringy` [@scag; @WW08] index is a non-smooth index based on scagnostics. It measures the proportion of vertices with two edges in the minimum spanning tree (MST) of the projection to detect whether the projection forms a straight line. The index is calculated as

$$I_{stringy}(A) = \frac{\text{number of vertices with 2 edges}}{\text{number of total vertices with more than one edge}} $$