### 2024-06-24 optim.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  number,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{algorithm}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{Journal of Multivariate Analysis}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num-names}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit},
  pdfauthor={H. Sherry Zhang; Dianne Cook; Nicolas Langrené; Jessica Wai Yin Leung},
  pdfkeywords={projection pursuit, Jellyfish Search Optimiser
(JSO), Optimisation},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Studying the Performance of the Jellyfish Search Optimiser for
the Application of Projection Pursuit}
\author[1]{H. Sherry Zhang%
\corref{cor1}%
}
 \ead{huize.zhang@austin.utexas.edu} 
\author[2]{Dianne Cook%
%
}
 \ead{dicook@monash.edu} 
\author[3]{Nicolas Langrené%
%
}
 \ead{nicolaslangrene@uic.edu.cn} 
\author[2]{Jessica Wai Yin Leung%
%
}
 \ead{Jessica.Leung@monash.edu}

\affiliation[1]{organization={University of Texas at Austin, Department
of Statistics and Data Sciences},city={Austin},country={United
States},countrysep={,},postcode={78751},postcodesep={}}
\affiliation[2]{organization={Monash University, Department of
Econometrics and Business
Statistics},city={Melbourne},country={Australia},countrysep={,},postcode={3800},postcodesep={}}
\affiliation[3]{organization={Guangdong Provincial Key Laboratory of
Interdisciplinary Research and Application for Data Science, BNU-HKBU
United International College, Department of Mathematical
Sciences},city={Zhuhai},country={China},countrysep={,},postcode={519087},postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{abstract}
Projection pursuit (PP) is a dimension reduction technique that
identifies low-dimensional projections in high-dimensional data by
optimising a criteria function known as the PP index. The optimisation
in PP can be non-smooth and requires identifying optima with a small
``squint angle'', detectable only from close proximity. To address these
challenges, this study investigates the performance of a recent
swarm-based algorithm, Jellyfish Search Optimiser (JSO), for optimising
PP indexes. The performance of JSO is evaluated across various
hyper-parameter settings and compared with existing optimisers.
Additionally, this work proposes novel methods to quantify two
properties of the PP index -- smoothness and squintability -- that
capture the complexities inherent in PP optimisation problems. These two
metrics are evaluated with JSO hyper-parameters to determine their
effect on JSO success rate. The JSO algorithm has been implemented in
the \texttt{tourr} package, while calculations for smoothness and
squintability are available in the \texttt{ferrn} package.
\end{abstract}

\begin{keyword}
    projection pursuit \sep Jellyfish Search Optimiser (JSO) \sep 
    Optimisation
\end{keyword}
\end{frontmatter}

\textbf{README}:

\begin{itemize}
\item
  British English (``American or British usage is accepted, but not a
  mixture of these'')
\item
  No ``we'' - always third person
\item
  Check if the affiliation information is correct
\item
  word use:

  \begin{itemize}
  \tightlist
  \item
    Use jellyfish search optimiser, or JSO
  \item
    ``hyper-parameter'' rather than ``hyperparameter''
  \item
    PP index rather than PPI, plural PP indexes
  \item
    datasets rather than data sets
  \end{itemize}
\end{itemize}

\section{Introduction {[}Nicolas and
Jessica{]}}\label{introduction-nicolas-and-jessica}

Projection Pursuit (PP) (\citet{kr69}, \citet{FT74}) is a dimension
reduction technique aimed at identifying informative linear projections
of data. The method involves optimising a objective function known as PP
index (\citet{hall1989polynomial}, \citet{cook1993projection},
\citet{lee2010projection}, \citet{Loperfido2018},
\citet{Loperfido2020}), which defines the criteria for what constitutes
interesting or informative projections. Given high-dimensional data
\(X \in \mathbf{R}^{n\times p}\) and an index function \(f(\cdot)\), PP
finds the orthonormal projection matrix
\(A \in \mathbf{R}^{p \times d}\) that maximises the index value of the
projection, \(Y = XA\):

\begin{equation}\phantomsection\label{eq-optimization}{
\underset{A}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I
}\end{equation}

The index functions can be non-linear and non-convex, hence an effective
and efficient optimisation procedure is essential to explore the data
landscape and achieving a globally optimal viewpoint of the data.

Significant efforts have been devoted to enhancing the functionality of
PP and facilitating a better experience in the visualisation process.
\citet{cook1995grand} introduces the PP guided tour, which enabled
interactive visualisation of the optimisation in order to visually
explore the high-dimensional data. It is implemented in the R \citep{R}
package \texttt{tourr} \citep{tourr}. \citet{RJ-2021-105} highlighted
potential problems of the optimisation routine implemented. While
improving the quality of the optimisation solutions in the tour is
essential, it is also important to be able to watch the projected data
as the optimisation progresses. As such, it is pertinent to integrate
the guided tour with a global optimisation algorithm that is efficient
in finding the global optimal and enables viewing of the projected data
during the exploration process.

This study explores the potential for a Jellyfish Search Optimiser (JSO)
to be integrated with projection pursuit guided tour. JSO
\citep{chou_novel_2021} -\citep{rajwar_exhaustive_2023}, inspired by the
search behaviour of jellyfish in the ocean, is a swarm-based
metaheuristic designed to solve global optimisation problems. Compared
to traditional methods, JSO demonstrated stronger search ability and
faster convergence, and requires fewer tuning parameters. These
practical advantages make JSO as a promising candidate for enhancing PP
optimisation. As such, the study aims to explore JSO's potential in
enhancing projection pursuit and examines the characteristics and
advantages of such implementation.

The primary goal of this study is to investigate the performance of JSO
in PP optimisation. A series of simulation experiments using
well-recognised datasets and various PP indexes are conducted to yield
insights regarding the behaviour of JSO under different settings and its
sensitivity to hyper-parameter. Second, to observe the performance of
JSO with different types of PP indexes, metrics are introduced to
capture specific properties of the index including squintability and
smoothness as introduced in \citet{laa_using_2020}. To the best of the
authors' knowledge, this is the first attempt to quantitatively measure
squintability and smoothness within this context. Finally, the
relationship between the JSO performance, hyper-parameters tuning and
various properties of PP indexes is analysed to provide helpful guidance
for practitioners using projection pursuit.

The rest of this paper is structured as follows.
Section~\ref{sec-background} introduces the background of PP guided tour
and reviews existing optimisers and index functions in the literature.
Section~\ref{sec-theory} details JSO and introduces metrics that measure
different properties of PP indexes. Section~\ref{sec-vis} outlines the
data structures, numerical summaries, and visualisations used in the
simulation studies conducted in Section~\ref{sec-app} to assess JSO's
perforamnce. Section~\ref{sec-app} presents two sets of simulation
experiments, comparing JSO with another optimiser, search better, and
studies the impact of different PP index properties on the optimisation
performance. Section~\ref{sec-conclusion} summarises the work and
provides suggestions for future directions.

\section{Projection pursuit, index functions and optimisation {[}Di and
Sherry{]}}\label{sec-background}

A tour on high-dimensional data is constructed by geodesically
interpolating between pairs of planes. Any plane is described by an
orthonormal basis, \(A_t\), where \(t\) represents time in the sequence.
The term ``geodesic'' refers to maintaining the orthonormality
constraint so that each view shown is correctly a projection of the
data. The PP guided tour operates by geodesically interpolating to
target planes (projections) which have high PP index values, as provided
by the optimiser. The geodesic interpolation means that the viewer sees
a continuous sequence of projections of the data, so they can watch
patterns of interest forming as the function is optimised. There are
five optimisation methods implemented in the \texttt{tourr} package:

\begin{itemize}
\tightlist
\item
  \texttt{search\_geodesic()}: provides a pseudo-derivative
  optimisation. It searches locally for the best direction, based on
  differencing the index values for very close projections. Then it
  follows the direction along the geodesic path between planes, stopping
  when the next index value fails to increase.
\item
  \texttt{search\_better()}: is a brute-force optimisation searching
  randomly for projections with higher index values.
\item
  \texttt{search\_better\_random()}: is essentially simulated annealing
  \citep{Bertsimas93} where the search space is reduced as the
  optimisation progresses.
\item
  \texttt{search\_posse()}: implements the algorithm described in
  \citet{posse95}.
\item
  \texttt{search\_polish()}: is a very localised search, to take tiny
  steps to get closer to the local maximum.
\end{itemize}

There are several PP index functions available: \texttt{holes()} and
\texttt{cmass()} \citep{cook1993projection}; \texttt{lda\_pp()}
\citep{lee2005projection}; \texttt{pda\_pp()} \citep{lee2010projection};
\texttt{dcor2d()} and \texttt{splines2d()} \citep{Grimm2016};
\texttt{norm\_bin()} and \texttt{norm\_kol()} \citep{huber85};
\texttt{slice\_index()} \citep{Laa:2020wkm}. Most are relatively simply
defined, for any projection dimension, and implemented because they are
relatively easy to optimise. A goal is to be able to incorporate more
complex PP indexes, for example based on scagnostics (\citet{scag},
\citet{WW08}).

An initial investigation of PP indexes, and the potential for
scagnostics is described in \citet{laa_using_2020}. To be useful here an
optimiser needs to be able to handle index functions which are possibly
not very smooth. In addition, because data structures might be
relatively fine, the optimiser needs to be able to find maxima that
occur with a small squint angle, that can only be seen from very close
by. One last aspect that is useful is for an optimiser to return local
maxima in addition to global because data can contain many different and
interesting features.

\section{The jellyfish optimiser and properties of PP indexes {[}Nicolas
and Jessica{]}}\label{sec-theory}

JSO mimics the natural movements of jellyfish, which include passive and
active motions driven by ocean currents and their swimming patterns,
respectively. In the context of optimization, these movements are
abstracted to explore the search space in a way that balances
exploration (searching new areas) and exploitation (focusing on
promising areas). The algorithm aims to find the optimal solution by
adapting the behaviour of jellyfish to navigate towards the best
solution over iterations \citep{chou_novel_2021}.

To solve the optimisation problem embedded in the PP guided tour, a
starting projection, an index function and the maximum number of
iterations are provided as input. Then, the current projection is
evaluated by the index function. The projection is then moved in a
direction determined by a random factor, influenced by how far along we
are in the optimization process. Occasionally, completely new directions
may be taken like a jellyfish might with ocean currents. A new
projection is accepted if it is an improvement compared to the current
one, rejected otherwise. This process continues and iteratively improves
the projection, until the pre-specified maximum number of trials is
reached.

\begin{tcolorbox}[enhanced jigsaw, opacityback=0, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, coltitle=black, titlerule=0mm, breakable, arc=.35mm, toprule=.15mm, left=2mm, bottomrule=.15mm, bottomtitle=1mm, colframe=quarto-callout-note-color-frame, colbacktitle=quarto-callout-note-color!10!white, colback=white, toptitle=1mm, rightrule=.15mm]

\textbf{Input}: \texttt{current\_projections}, \texttt{index\_function},
\texttt{trial\_id}, \texttt{max\_trial}

\textbf{Output}: \texttt{optimized\_projection}

\textbf{Initialize} \texttt{current\_best} as the projection with the
best index value from \texttt{current\_projections}, and
\texttt{current\_idx} as the array of index function values for each
projection in \texttt{current\_projections}

\textbf{for} each trial\_id in 1 to max\_tries \textbf{do}

\begin{quote}
Calculate \(c_t\) based on the current\_idx and max\_trial
\end{quote}

\begin{quote}
\textbf{if} \(c_t\) is greater than or equal to \(0.5\) \textbf{then}

\begin{quote}
Define trend based on the current\_best and current\_projections
\end{quote}

\begin{quote}
Update each projection towards the trend using a random factor and
orthonormalisation
\end{quote}

\textbf{else}

\begin{quote}
\textbf{if} a random number is greater than \(1 - c_t\) \textbf{then}

\begin{quote}
Slightly adjust each projection with a small random factor (passive)
\end{quote}

\textbf{else}

\begin{quote}
For each projection, compare with a random jelly and adjust towards or
away from it (active)
\end{quote}
\end{quote}

Update the orientation of each projection to maintain consistency

Evaluate the new projections using the index function
\end{quote}

\begin{quote}
\textbf{if} any new projection is worse than the current, revert to the
\texttt{current\_projections} for that case

\begin{quote}
Determine the projection with the best index value as the new
current\_best
\end{quote}
\end{quote}

\begin{quote}
\textbf{if} trial\_id \(\geq\) max\_trial, print the last best
projection \textbf{exit}
\end{quote}

\textbf{return} the set of projections with the updated current\_best as
the optimized\_projection

\end{tcolorbox}

The JSO implementation involves several key parameters that control its
search process in optimization problems. These parameters are designed
to guide the exploration and exploitation phases of the algorithm. While
the specific implementation details can vary depending on the version of
the algorithm or its application, we focus on two main parameters that
are most relevant to our application: the number of jellyfish and the
maximum number of trials.

\citet{laa_using_2020} has proposed five criteria for assessing
projection pursuit indexes (smoothness, squintability, flexibility,
rotation invariance, and speed). Since not all the properties affects
the execution of the optimisation, here we consider the three relevant
properties (smoothness, squintability, and speed), and propose three
metrics to evaluate these three properties.

\subsection{Smoothness}\label{sec-smoothness}

If we evaluate the index function at some random points (like the random
initialization of the jellyfish optimizer), then we can interpret these
random index values as a random field, indexed by a space parameter: the
random projection angle. This analogy suggests to use this random
training sample to fit a spatial model, a simple one being a (spatial)
Gaussian process.

How can we define a measure of smoothness from this? The distribution of
a Gaussian process is fully determined by its mean function and
covariance function. The way the covariance function is defined is where
smoothness comes into play: if an index is very smooth, then two close
projection angles should produce close index values (strong
correlation); by contrast, if an index is not smooth, then two close
projection angles might give very different index values (fast decay of
correlations with respect to distance between angles).

Popular covariance functions are parametric positive semi-definite
functions, some of which have a parameter to capture the smoothness of
the Gaussian field. In particular, consider the Matérn class of
covariance functions, defined by

\[
K(u):=\frac{(\sqrt{2\nu}u)^{\nu}}{\Gamma(\nu)2^{\nu-1}}\mathcal{K}_{\nu}(\sqrt{2\nu}u)
\]

where \(\nu>0\) is the smoothness parameter and where
\(\mathcal{K}_\nu\) is the modified Bessel function. The Matérn
covariance function can be expressed analytically when \(\nu\) is a
half-integer, the most popular values in the literature being \(1/2\),
\(3/2\) and \(5/2\) . The parameter \(\nu\), called smoothness
parameter, controls the decay of the covariance function. As such, it is
an appropriate measure of smoothness of a random field.

In our context, we suggest to use this parameter as a measure of the
smoothness of the index function by fitting a Gaussian process prior
with Matérn covariance on a dataset generated by random evaluations of
the index function, as in the initial stage of the jellyfish random
search. There exist several R packages, such as GpGp or ExaGeoStatR, to
fit the hyperparameters of a GP covariance function on data. In this
project, we make use of the GpGp package.

The fitted value \(\nu>0\) can be interpreted as follows: the higher
\(\nu\), the smoother the index function.

\subsection{Squintability}\label{sec-squintability}

From the literature, it is commonly understood that a large squint angle
implies that the objective function value is close to optimal even when
we are not very close to the perfect view to see the structure. A small
squint angle means that index function value improves substantially only
when we are very close to the perfect view. As such, low squintability
implies rapid improvement in the index value when near the perfect view.

In this subsection, a metric to capture the notion of squintability is
proposed, and two approaches to compute this metric numerically are
detailed.

Let \(A \in \mathbf{R}^{p \times 2}\) be the 2D projection basis and
\(A^*\) the optimal basis achieving the maximum index value for a given
projection pursuit problem defined by the shape to find, data dimension,
and the index function \(f(.)\). The projection distance \(d(A, A^*)\),
is defined as the Frobenius norm \(\lVert . \rVert _F\), between the
squared difference of these two bases:

\begin{equation}\phantomsection\label{eq-proj-dist}{d(A, A^*) = \lVert M \rVert _F = \sum_{ij} M_{ij}^2, \text{ where } M = AA^\prime - A^*A^{*\prime}}\end{equation}

The index value \(f(XA)\) is defined as described in the introduction.

\textcolor{orange}{Here we need to add a precise, mathematical
definition of the $x$ and $y$ variables ("angle distance" to optimal?
and PP index values) used in the figures of this Section.}

{[}We generate random points that is beyond 1.5 projection distance and
interpolate. Then we fit a kernel or use nonlinear least squares.{]}

Let \(f\) represent the average PP index value with respect to the
projection angle distance to optimal, over the course of the JSO. It is
expected that for a PP index with high squint angle, the optimization
(\ref{eq-optimization}) should make substantial progress early on.
Conversely, for a PP index with low squint angle, it might take a long
while for the optimization to make substantial progress, as the
candidate projections would need to be very close to the optimal one for
the structure of the index function to be visible enough to be amenable
to efficient optimization. This observation suggests that the extreme
values of \(f'\) (the ones for which \(f''=0\), assuming that \(f\) is
twice differentiable), and the angles for which these values are
attained, should play an important role in the mathematical definition
of squintability. Noticing that \(f\) is expected to be a decreasing
function, we propose the following squintability metric:

\begin{equation}\phantomsection\label{eq-squintability}{
\varsigma:=-4\min_{x}f'(x)\times\underset{x}{\arg\min}f'(x)
}\end{equation}

On the one hand, \(-\min_{x}f'(x)\) represents the largest value of the
gradient of \(-f\). On the other hand, \(\underset{x}{\arg\min}f'(x)\)
represents the angle value at which this largest gradient is attained.
Based on the above discussion, it is expected that these two values
should be both high in the case of high squintability (fast increase in
\(f\) early on), and both low in the case of low squintability (any
substantial increase in \(f\) happens very late, close to the optimal
angle). This suggest that their product (\ref{eq-squintability}) should
provide a sensible measure of squintability. The multiplicative constant
\(4\), which can be deemed arbitrary, does not change the interpretation
of the squintability metric \(\varsigma\); it is here to adjust the
range of values of \(\varsigma\) and simplify the explicit formula for
\(\varsigma\) obtained later on.

To compute the squintability metric (\ref{eq-squintability}) in
practice, several approaches are possible. The first one is to propose a
parametric model for \(f\), and use it to obtain an explicit formula for
\(\varsigma\). Numerical experiments suggest a scaled sigmoid shape as
described below. Define

\begin{equation}\phantomsection\label{eq-logistic}{
\ell(x):=\frac{1}{1+\exp(\theta_{3}(x-\theta_{2}))}\ ,
}\end{equation}

which is a decreasing logistic function depending on two parameters
\(\theta_2\) and \(\theta_3\) , such that
\(\ell(\theta_{2})=\frac{1}{2}\). Then, define

\begin{equation}\phantomsection\label{eq-parametric}{
f(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(x_{\max})}{\ell(0)-\ell(x_{\max})}+\theta_{4}\ ,
}\end{equation}

which depends on three additional parameters, \(\theta_1\), \(\theta_2\)
, and \(x_{\max}\), such that \(f(0)=\theta_1\) and
\(f(x_{\max})=\theta_4\). Under the parametric model
(\ref{eq-parametric}), the squintability metric (\ref{eq-squintability})
can be shown to be equal to

\begin{equation}\phantomsection\label{eq-squintability-parametric}{
\varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}
}\end{equation}

In practice, the parameters of this model (\ref{eq-parametric}) can be
estimated numerically, for example by non-linear least squares, and then
used to evaluate \(\varsigma\) as in equation
(\ref{eq-squintability-parametric}).

Alternatively, one can estimate (\ref{eq-squintability}) in a
nonparametric way, for example by fitting \(f\) using kernel regression,
then numerically estimate the angle at which \(-f'\) attains its highest
value.

\textcolor{orange}{See whether further details are needed for the nonparametric approach.}

To the best of our knowledge, this is the first attempt to measure the
notion of squintability.

\section{Visualisation of jellyfish optimiser}\label{sec-vis}

The data structure proposed in \citet{RJ-2021-105} is implemented for
JSO. An example data collected from finding the sine-wave structure in
6D data using a spline index function (\texttt{splines2d}) is printed
below:

\begin{verbatim}
Rows: 10
Columns: 12
$ index     <chr> "splines2d", "splines2d", "splines2d", "splines2d", "splines~
$ d         <dbl> 6, 6, 6, 6, 6, 6, 6, 6, 6, 6
$ n_jellies <dbl> 50, 50, 50, 50, 50, 50, 50, 50, 50, 50
$ max_tries <dbl> 50, 50, 50, 50, 50, 50, 50, 50, 50, 50
$ sim       <int> 1, 1, 1, 1, 1, 1, 1, 1, 1, 1
$ seed      <int> 3462, 3462, 3462, 3462, 3462, 3462, 3462, 3462, 3462, 3462
$ basis     <list> <<matrix[6 x 2]>>, <<matrix[6 x 2]>>, <<matrix[6 x 2]>>, <<m~
$ index_val <dbl> 0.036664532, 0.005121912, 0.058683104, 0.056527991, 0.015734~
$ method    <chr> "search_jellyfish", "search_jellyfish", "search_jellyfish", ~
$ tries     <dbl> 1, 1, 1, 1, 1, 1, 1, 1, 1, 1
$ loop      <dbl> 1, 2, 3, 4, 5, 6, 7, 8, 9, 10
$ time      <drtn> 4.362133 mins, 4.362133 mins, 4.362133 mins, 4.362133 mins, ~
\end{verbatim}

The JSO hyper-parameters used are 50 jellyfishes (\texttt{n\_jellies})
and a maximum of 50 tries (\texttt{max\_tries}). The full data contains
50 simulations, indexed by the variable \texttt{sim}, with the
associated \texttt{seed}s. Recorded in columns \texttt{basis},
\texttt{index\_val}, \texttt{tries}, \texttt{loop} and \texttt{time},
are the sampled projection basis, index value, iteration ID, jellyfish
ID, and computational time respectively. The basis column records every
visited basis before comparing with the current (last) basis. If a
visited basis has a smaller index value than the current (last) one, JSO
will keep the current basis for the next iteration in the search but
still records the visited inferior basis. This data can be used to
derive numerical summaries and plots:

\begin{itemize}
\item
  \textbf{Success rate}: The success rate is defined as the proportion
  of simulations that achieves a final index value within 0.05 of the
  best index value found among all 50 simulations. In the dataset above,
  the optimal index value is 0.999. Out of the 50 simulations, 44
  achieved an index value within {[}0.949, 0.999{]}, resulting in a
  success rate of 0.88. Figure~\ref{fig-projection} shows the best
  projection found in each simulation, sorted by the index value. Most
  of the projections show a clear sine-wave structure, while the last
  six show only a straight line with clusters at the end. When local
  optima exist in the projection pursuit problem, plotting the
  projections found in each simulation or by each jellyfish can help
  identify these local optima.
\item
  \textbf{Projection distance}: The projection distance between
  projection bases and the optimal best basis (or the empirical best
  basis if the optimal is unknown) measures how close a visited basis is
  to the best basis and it is calculated as the Frobenius norm in
  Equation

  \begin{enumerate}
  \def\labelenumi{\arabic{enumi}.}
  \tightlist
  \item
    Plotting projection distances against index values for the bases
    visited reveals how the index value changes as the basis approaches
    the optimal basis, helping to calculate the squintability measure.
    Figure~\ref{fig-index-value-proj-dist} shows the index value against
    the projection distance for the given dataset and its binned
    version. As the projection distance decreases, the index value
    initially increases with an increasing slope and then levels off.
    The index value shows greater variation at medium projection
    distances compared to small and large distances.
  \end{enumerate}
\end{itemize}

\begin{figure}

\centering{

\includegraphics{optim_files/figure-pdf/fig-projection-1.pdf}

}

\caption{\label{fig-projection}Best projections found by the JSO in each
of the 50 simulations, sorted by index values. The projection pursuit
problem is to detect the 6D sine-wave structure using the spline index.
Most simulations have detected a clear sine-wave pattern, while the last
six show a straight line with clustering at the ends. In this
simulation, the success rate is 0.88 (44/50).}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{optim_files/figure-pdf/fig-index-value-proj-dist-1.pdf}

}

\caption{\label{fig-index-value-proj-dist}Projection distance plotted
against the index value for detecting the 6D sine-wave structure using
the spline index function. a) The index values start around 0 with low
variance at large projection distances. As the projection distance
decreases, the index values increase with a larger variance. Towards the
minimum projection distance, the index values increase and cluster
around 1, indicating the optimiser is approaching to the optimum. b)
Projection distances are binned with a bin width of 0.1, and index
values are averaged over each bin. The relationship between the two
variables forms a sigmoid curve.}

\end{figure}%

\section{Application {[}Di and Sherry{]}}\label{sec-app}

The JSO provides a new swarm-based search strategy to optimise
projection pursuit problems. Two examples are provided in this section
to demonstrate its performance. The first example compares the
performance of JSO with an exist optimiser, search better, and explores
how JSO behaves with different combinations of hyper-parameters. The
second example studies the effect of optimisation properties defined in
Section~\ref{sec-theory}, along with JSO hyper-parameters, on the
outcome (success rate) of the optimisation.

\subsection{Performance of JSO in pipe-finding
problems}\label{sec-app-1}

The performance of JSO is investigated in two folds: 1) compare it with
a commonly used optimiser: the search-better optimiser
\citep{RJ-2021-105, laa_using_2020}, and 2) examine its success rate
under different hyper-parameters (number of jellyfishes and maximum
number of tries). The projection pursuit problem used is finding the
pipe shape using the holes index, investigated by
\citet{laa_using_2020}.

\begin{figure}

\centering{

\includegraphics{optim_files/figure-pdf/fig-proj-1.pdf}

}

\caption{\label{fig-proj}Projections found by the jellyfish and search
better optimisers at each 10th quantile across 50 simulations. The
projection pursuit problem is to find the pipe shape using the holes
index in the 6, 8, 10, and 12-dimensional spaces. The JSO uses 100
jellyfishes and a maximum number of tries of 100. The search better
optimiser uses a maximum of 1000 tries in each step of random sampling
step before the algorithm terminates. In the 6-D data space, JSO always
finds a clear pipe shape while the search better optimiser also finds
the pipe shape but with a wide rim. At higher data dimensions, JSO finds
a higher index value and a clearer pipe shape across all the quantiles
than the search better optimiser.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{optim_files/figure-pdf/fig-proportion-1.pdf}

}

\caption{\label{fig-proportion}Proportion of simulations reaches
near-optimal index values in the pipe-finding problem using the holes
index. The proportion is calculated based on the number of simulations,
out of 50, that achieve an index value within 0.05 of the
best-performing simulation. As the dimensionality increases, the
proportion of simulations reaching the optimal index value increases.}

\end{figure}%

Fifty simulations are conducted with both JSO and the search better
optimal, in four dimensions (\(d = 6, 8, 10, 12\)). JSO uses 100
jellyfishes and 100 maximum number of tries, while the search-better
optimiser allows a maximum of 1000 samples at each iteration before the
algorithm terminates. Figure~\ref{fig-proj} presents the final
projections found by the two optimisers, broken down by 10th quantile,
faceted by the data dimension. In the 6-dimensional data scenario, JSO
consistently identifies a clear pipe shape. The search better optimiser
also finds the pipe shape but with a wide rim, suggesting a further
polish search may be required. With increasing dimensions, JSO may not
always identify the pipe shape due to random sampling, but it still
presents a pipe shape in over 50\% of cases. When compared to the search
better optimiser, JSO reaches higher index values and clearer pipe
shapes across all quantiles.

In the second experiment, fifty simulations are conducted at each
hyper-parameter combination to analyse its effects on the JSO success
rate. The hyper-parameters tested include: 1) 20, 50, and 100
jellyfishes, and 2) 50 and 100 maximum number of tries. In each
simulation, the success rate is calculated as defined in
Section~\ref{sec-vis}. Figure~\ref{fig-proportion} presents the
proportion of success for each hyper-parameter combinations. As the
number of jellyfishes and maximum tries increase, the success rate also
increases. For simpler problems (6 dimensions), small parameter values
(20 jellyfishes and a maximum number of tries of 50) can already result
in high success rate, while for higher-dimensional problems (8, 10, and
12 dimensions), a combination of 100 jellyfishes and 100 maximum number
of tries is necessary for at least half of the jellyfishes to find the
projection close to the optimal.

\subsection{Factors affecting JSO success rate: optimisation properties
and jellyfish hyper-parameters}\label{sec-app-2}

The optimisation properties, including smoothness, squintability, and
speed, offer numerical metrics to characterise the complexity of
projection pursuit optimisation problems. This example investigates how
these metrics, along with the hyper-parameters affect the JSO success
rate. Simulations are conducted to obtain the success rate across a
wider array of projection pursuit problems, characterised by
shape-to-find, data dimension, and the index function, as well as
different hyper-parameter combinations. Smoothness and squintability are
calculated for each projection pursuit problem as described in
Section~\ref{sec-theory}. A generalised linear model is used to
construct the relationship between the success rate, JSO
hyper-parameters and optimisation properties.

In addition to the pipe-finding problem, a new shape, sine wave, is
investigated in 6D and 8D spaces with six indexes considered:
\texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
\texttt{spline}, and \texttt{stringy}. Combining with two JSO
hyper-parameters, a total of 52 cases is produced, comprising of 24
pipe-finding cases and 28 sine-wave finding cases. For each case, JSO is
run fifty times and the success rate is calculated as in
Section~\ref{sec-app-1}.

Smoothness and squintability are computed for each case, following the
procedures outlined in Section~\ref{sec-smoothness} and
Section~\ref{sec-squintability}. To calculate smoothness, three hundred
random bases are simulated. Index values and projection distance (to the
optimal basis) are calculated for each random basis before fitting into
a Gaussian process model to obtain the smoothness measure. For
squintability, fifty random bases are sampled and interpolated to the
optimal basis with a step size of 0.005. Index values and projection
distances are calculated for these interpolated bases and the index
values are averaged with a bin width of 0.005. The squintability measure
is then calculated from fitting a 4-parameter scaled logistic function
of index value against projection distance, estimated by non-linear
least squares.

Table~\ref{tbl-smoothness-squintability} presents the parameters
estimated from the Gaussian process (variance, range, smooth, and
nugget) and the scaled logistic function (\(\theta_1\) to \(\theta_4\))
for each case. The column ``smooth'' is used as the smoothness measure
and the column ``squint'' is calculated as equation
\ref{eq-squintability-parametric} as squintability. The low squint value
of MIC and TIC index is due to its convex shape of the index value
against the projection distance, as shown in
Figure~\ref{fig-idx-proj-dist}. Comparing to other concave shapes, the
maximum first derivative happens at a smaller projection distance
(\(\theta_2\)), requiring the optimiser to get closer to the optimal to
see a significant change in the index value, hence a small squintability
measure.

\begin{figure}

\centering{

\includegraphics{optim_files/figure-pdf/fig-idx-proj-dist-1.pdf}

}

\caption{\label{fig-idx-proj-dist}Index values versus projection
distance for the 12 pipe/sine-wave finding problem, after the binning
procedure for calculating the squintability measure. The index values,
averaged at bin width of 0.005, are scaled from 0 to 1 for comparison
(holes, TIC, and stringy). The \texttt{MIC} and \texttt{TIC} index
curves are convex while others are concave. The stringy curve shows an
instantaneous jump to the optimum when aproaching the best basis.}

\end{figure}%

\begin{table}

\caption{\label{tbl-smoothness-squintability}Parameters estimated from
the Gaussian process (including variance, range, smoothness, and nugget)
and scaled logistic function (\(\theta_1\) to \(\theta_4\)) for the
pipe-finding and sine-wave finding problems. The column ``smooth'' and
``squint'' represent the smoothness and squintability measures.}

\centering{

\centering\begingroup\fontsize{7}{9}\selectfont

\begin{tabular}{|>{}lll>{}r|rrr>{}r|rrrr>{}r|}
\toprule
  & shape & index & d & variance & range & smooth & nugget & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\theta_4$ & squint\\
\midrule
1 & pipe & holes & 6 & 0.002 & 0.371 & 2.355 & 0.221 & 1.015 & 0.860 & 3.368 & 0.018 & 3.053\\
2 & pipe & holes & 8 & 0.000 & 0.178 & 2.189 & 0.820 & 1.014 & 0.869 & 3.264 & 0.029 & 2.960\\
3 & pipe & holes & 10 & 0.000 & 0.107 & 2.192 & 3.027 & 1.016 & 0.885 & 3.151 & 0.022 & 2.948\\
4 & pipe & holes & 12 & 0.000 & 0.148 & 2.295 & 1.581 & 1.011 & 0.878 & 3.345 & 0.004 & 3.117\\
5 & sine & MIC & 6 & 0.016 & 0.092 & 2.457 & 0.083 & 0.894 & 0.571 & 1.623 & -0.024 & 1.257\\
\addlinespace
6 & sine & MIC & 8 & 0.016 & 0.081 & 2.636 & 0.084 & 0.932 & 0.328 & 1.314 & -0.030 & 0.772\\
7 & sine & TIC & 6 & 0.124 & 0.110 & 2.444 & 0.086 & 0.951 & 0.536 & 1.719 & -0.027 & 1.322\\
8 & sine & TIC & 8 & 0.125 & 0.099 & 2.475 & 0.085 & 0.949 & 0.564 & 1.723 & -0.028 & 1.369\\
9 & sine & dcor2d & 6 & 0.034 & 0.167 & 2.663 & 0.114 & 0.954 & 1.039 & 2.742 & -0.019 & 2.947\\
10 & sine & loess2d & 6 & 0.079 & 0.336 & 1.988 & 0.309 & 1.016 & 1.039 & 2.648 & 0.080 & 2.756\\
\addlinespace
11 & sine & splines2d & 6 & 0.042 & 0.242 & 2.537 & 0.102 & 1.014 & 1.051 & 2.730 & -0.009 & 3.118\\
12 & sine & stringy & 6 & 0.000 & 0.007 & 1.536 & 38.166 & 1.045 & 0.011 & 254.734 & 0.053 & 2.957\\
\bottomrule
\end{tabular}
\endgroup{}

}

\end{table}%

\begingroup\fontsize{7}{9}\selectfont

\begin{longtable}[t]{lrrrrrrl}

\caption{\label{tbl-mod-data}The first 7 rows of the datasets processed
for modelling.}

\tabularnewline

\toprule
Index & D & Smoothness & Squintability & Num. of jellyfish & Max. Num. of tries & Prop. of success & Time\\
\midrule
MIC & 6 & 2.457 & 1.257 & 20 & 50 & 0.12 & 2.479 secs\\
MIC & 6 & 2.457 & 1.257 & 20 & 100 & 0.24 & 8.950 secs\\
MIC & 6 & 2.457 & 1.257 & 50 & 50 & 0.52 & 5.651 secs\\
MIC & 6 & 2.457 & 1.257 & 50 & 100 & 0.64 & 13.223 secs\\
MIC & 6 & 2.457 & 1.257 & 100 & 50 & 0.76 & 19.453 secs\\
\addlinespace
MIC & 8 & 2.636 & 0.772 & 20 & 50 & 0.08 & 2.566 secs\\
MIC & 8 & 2.636 & 0.772 & 20 & 100 & 0.08 & 4.960 secs\\
\bottomrule

\end{longtable}

\endgroup{}

\begingroup\fontsize{7}{9}\selectfont

\begin{longtable}[t]{|>{}lrrr>{}r|}

\caption{\label{tbl-mod-output}Model estimates of proportion of
jellyfish success on optimisation properties and jellyfish
hyper-parameters from the generalised linear model with a binomial
family and a logit link function. The variable smoothness,
squintability, number of jellyfish and maximum number of tries are
positively associated with JSO success rate while data dimension and
being flagged as long runtime are negatively associated with the success
rate. The variable squintability and dimension are significant,
suggesting their importance relative to jellyfish hyper-parameters in
the optimisation success.}

\tabularnewline

\toprule
term & estimate & std.error & statistic & p.value\\
\midrule
Intercept & -4.523 & 5.332 & -0.848 & 0.396\\
Smoothness & 1.195 & 1.915 & 0.624 & 0.533\\
Squintability & 2.060 & 0.684 & 3.011 & 0.003\\
dimension (d) & -0.628 & 0.255 & -2.462 & 0.014\\
long time & -0.874 & 1.286 & -0.680 & 0.497\\
\addlinespace
number of jellyfish & 0.216 & 0.127 & 1.701 & 0.089\\
maximum number of tries & 0.113 & 0.150 & 0.752 & 0.452\\
\bottomrule

\end{longtable}

\endgroup{}

Table~\ref{tbl-mod-data} combines all the variables calculated above
(JSO success rate, hyper-parameters, and optimisation properties) into
one table. Three pre-processing steps are applied to the data: 1)
scaling the JSO hyper-parameters by a factor of 10 for interpretation,
2) creating a new binary variable \texttt{long\_time} to indicate cases
with an average run time over 30 seconds, and 3) encoding the success
rate for the stringy case as 0, because upon visual inspection of the
final projection, none of the 50 simulations identified the sine-wave
shape. A generalised linear model with a binomial family and a logit
link function is used to fit the data and Table~\ref{tbl-mod-output}
presents the model outputs. The model suggests that JSO success rate is
positively associated with the two hyper-parameters, as well as with the
optimisation properties -- smoothness and squintability. Specifically,
using 10 more jellyfishes and 10 more tries will increase the odd ratio
of success by 24.11\% and 11.93\%, respectively. However, being flagged
with long runtime and an increase of data dimension will reduce the
success rate by 41.72\% and 53.36\%, respectively. The variable
\texttt{squintability} and \texttt{dimension} are significant,
suggesting their importance relative to JSO hyper-parameters in the
optimisation success. Users should consider increasing the number of
jellyfish and the number of tries to improve the success rate, until JSO
exceeds a 30-second runtime.

\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}

\renewcommand\refname{References}
  \bibliography{bibliography.bib}

\end{document}