### 2024-06-30 optim.tex
conclusions drafted

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  number,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{algorithm}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{Journal of Multivariate Analysis}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num-names}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit},
  pdfauthor={H. Sherry Zhang; Dianne Cook; Nicolas Langrené; Jessica Wai Yin Leung},
  pdfkeywords={projection pursuit, jellyfish search optimiser
(JSO), optimisation, grand tour, high-dimensional data, exploratory data
analysis},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Studying the Performance of the Jellyfish Search Optimiser for
the Application of Projection Pursuit}
\author[1]{H. Sherry Zhang%
\corref{cor1}%
}
 \ead{huize.zhang@austin.utexas.edu} 
\author[2]{Dianne Cook%
%
}
 \ead{dicook@monash.edu} 
\author[3]{Nicolas Langrené%
%
}
 \ead{nicolaslangrene@uic.edu.cn} 
\author[2]{Jessica Wai Yin Leung%
%
}
 \ead{Jessica.Leung@monash.edu}

\affiliation[1]{organization={University of Texas at Austin, Department
of Statistics and Data Sciences},city={Austin},country={United
States},countrysep={,},postcode={78751},postcodesep={}}
\affiliation[2]{organization={Monash University, Department of
Econometrics and Business
Statistics},city={Melbourne},country={Australia},countrysep={,},postcode={3800},postcodesep={}}
\affiliation[3]{organization={Guangdong Provincial Key Laboratory of
Interdisciplinary Research and Application for Data Science, BNU-HKBU
United International College, Department of Mathematical
Sciences},city={Zhuhai},country={China},countrysep={,},postcode={519087},postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{abstract}
The projection pursuit (PP) guided tour interactively optimises a
criteria function known as the PP index, to explore high-dimensional
data by revealing interesting projections. The optimisation in PP can be
non-trivial, involving non-smooth functions and optima with small a
``squint angle'', detectable only from close proximity. To address these
challenges, this study investigates the performance of a recently
introduced swarm-based algorithm, Jellyfish Search Optimiser (JSO), for
optimising PP indexes. The performance of JSO for visualising data is
evaluated across various hyper-parameter settings and compared with
existing optimisers. Additionally, this work proposes novel methods to
quantify two properties of the PP index -- smoothness and squintability
-- that capture the complexities inherent in PP optimisation problems.
These two metrics are evaluated with JSO hyper-parameters to determine
their effect on JSO success rate. The JSO algorithm has been implemented
in the \texttt{tourr} package, while functions to calculate smoothness
and squintability of indexes are available in the \texttt{ferrn}
package.
\end{abstract}

\begin{keyword}
    projection pursuit \sep jellyfish search optimiser
(JSO) \sep optimisation \sep grand tour \sep high-dimensional data \sep 
    exploratory data analysis
\end{keyword}
\end{frontmatter}

\section{Introduction {[}Nicolas and
Jessica{]}}\label{introduction-nicolas-and-jessica}

Projection Pursuit (PP) (\citet{kr69}, \citet{FT74}) is a dimension
reduction technique aimed at identifying informative linear projections
of data. The method involves optimising a objective function known as
the PP index (e.g. \citet{hall1989polynomial},
\citet{cook1993projection}, \citet{lee2010projection},
\citet{Loperfido2018}, \citet{Loperfido2020}), which defines the
criteria for what constitutes interesting or informative projections.
Given high-dimensional data \(X \in \mathbf{R}^{n\times p}\) and an
index function \(f(\cdot)\), PP finds the orthonormal basis
\(A \in \mathbf{R}^{p \times d}\) that maximises the index value of the
projection, \(Y = XA\):

\begin{equation}\phantomsection\label{eq-optimization}{
\underset{A}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I
}\end{equation}

The index functions can be non-linear and non-convex, hence an effective
and efficient optimisation procedure is essential to explore the data
landscape and achieving a globally optimal viewpoint of the data.

Significant efforts have been devoted to enhancing the functionality of
PP, improving optimisation (e.g. \citet{posse1995},
\citet{marie-sainte2010}, \citet{grochowski2011}) and extending it's use
to visually explore high-dimensional data. \citet{cook1995grand}
introduced the PP guided tour, which monitors the optimisation visually
to see the data projections leading in and out of the optima. An
implementation is available in the R \citep{R} package \texttt{tourr}
\citep{tourr}. \citet{RJ-2021-105} illustrated how to diagnose
optimisation processes, particularly focusing on the guided tour, and
revealed a need to for improved optimisation. While improving the
quality of the optimisation solutions in the tour is essential, it is
also important to be able to view the data projections as the
optimisation progresses. Integrating the guided tour with a global
optimisation algorithm that is efficient in finding the global optimal
and enables viewing of the projected data during the exploration process
is a goal.

Here, the potential for a Jellyfish Search Optimiser (JSO) to be
integrated with the projection pursuit guided tour is explored. JSO
\citep{chou_novel_2021} -\citep{rajwar_exhaustive_2023}, inspired by the
search behaviour of jellyfish in the ocean, is a swarm-based
metaheuristic designed to solve global optimisation problems. Compared
to traditional methods, JSO demonstrated stronger search ability and
faster convergence, and requires fewer tuning parameters. These
practical advantages make JSO a promising candidate for enhancing PP
optimisation.

The primary goal of the study reported here is to investigate the
performance of JSO in PP optimisation for the guided tour. It is of
interest to assess how quickly and closely the optimiser reaches a gobal
optima, for various PP indexes that may have differing complexities. To
observe the performance of JSO with different types of PP indexes,
metrics are introduced to capture specific properties of the index
including squintability (based on \citet{barnett1981interpreting}'s
squint angle) and smoothness as introduced in \citet{laa_using_2020}.
Here we mathematically define metrics for squintability and smoothness,
which is new for the field. A series of simulation experiments using
specially structured datasets and various PP indexes are conducted to
assess the behaviour of JSO under different settings and its sensitivity
to hyper-parameter choices (number of jellies and maximum number of
tries). The relationship between the JSO performance, hyper-parameter
tuning and properties of PP indexes is analysed to provide guidance for
practitioners using projection pursuit, and developers of new indexes.

The paper is structured as follows. Section~\ref{sec-background}
introduces the background of PP guided tour and reviews existing
optimisers and index functions in the literature.
Section~\ref{sec-theory} details JSO and introduces metrics that measure
different properties of PP indexes. Section~\ref{sec-sim-deets} outlines
the data structures, numerical summaries, and visualisations used in the
simulation studies conducted in \textbf{?@sec-app} to assess JSO's
performance. Section~\ref{sec-sim-res} presents two sets of simulation
experiments, comparing JSO permformance improvements relative to an
existing optimiser, and studies the impact of different PP index
properties on the optimisation performance. Section~\ref{sec-conclusion}
summarises the work and provides suggestions for future directions.

\section{Projection pursuit, tours, index functions and optimisation
{[}Di and Sherry{]}}\label{sec-background}

A tour on high-dimensional data is constructed by geodesically
interpolating between pairs of planes. Any plane is described by an
orthonormal basis, \(A_t\), where \(t\) represents time in the sequence.
The term ``geodesic'' refers to maintaining the orthonormality
constraint so that each view shown is correctly a projection of the
data. The PP guided tour operates by geodesically interpolating to
target planes (projections) which have high PP index values, as provided
by the optimiser. The geodesic interpolation means that the viewer sees
a continuous sequence of projections of the data, so they can watch
patterns of interest forming as the function is optimised. There are
five (unsophisticated) optimisation methods implemented in the
\texttt{tourr} package:

\begin{itemize}
\tightlist
\item
  \texttt{search\_geodesic()}: provides a pseudo-derivative
  optimisation. It searches locally for the best direction, based on
  differencing the index values for very close projections. Then it
  follows the direction along the geodesic path between planes, stopping
  when the next index value fails to increase.
\item
  \texttt{search\_better()}: is a brute-force optimisation searching
  randomly for projections with higher index values.
\item
  \texttt{search\_better\_random()}: is essentially simulated annealing
  \citep{Bertsimas93} where the search space is reduced as the
  optimisation progresses.
\item
  \texttt{search\_posse()}: implements the algorithm described in
  \citet{posse95}.
\item
  \texttt{search\_polish()}: is a very localised search, to take tiny
  steps to get closer to the local maximum.
\end{itemize}

There are several PP index functions available: \texttt{holes()} and
\texttt{cmass()} \citep{cook1993projection}; \texttt{lda\_pp()}
\citep{lee2005projection}; \texttt{pda\_pp()} \citep{lee2010projection};
\texttt{dcor2d()} and \texttt{splines2d()} \citep{Grimm2016};
\texttt{norm\_bin()} and \texttt{norm\_kol()} \citep{huber85};
\texttt{slice\_index()} \citep{Laa:2020wkm}. Most are relatively simply
defined, for any projection dimension, and implemented because they are
relatively easy to optimise. A goal is to be able to incorporate more
complex PP indexes, for example based on scagnostics (\citet{scag},
\citet{WW08}).

An initial investigation of PP indexes, and the potential for
scagnostics is described in \citet{laa_using_2020}. To be useful here an
optimiser needs to be able to handle index functions which are possibly
not very smooth. In addition, because data structures might be
relatively fine, the optimiser needs to be able to find maxima that
occur with a small squint angle, that can only be seen from very close
by. One last aspect that is useful is for an optimiser to return local
maxima in addition to global because data can contain many different and
interesting features.

\section{The jellyfish optimiser and properties of PP indexes {[}Nicolas
and Jessica{]}}\label{sec-theory}

JSO mimics the natural movements of jellyfish, which include passive and
active motions driven by ocean currents and their swimming patterns,
respectively. In the context of optimization, these movements are
abstracted to explore the search space in a way that balances
exploration (searching new areas) and exploitation (focusing on
promising areas). The algorithm aims to find the optimal solution by
adapting the behaviour of jellyfish to navigate towards the best
solution over iterations \citep{chou_novel_2021}.

To solve the optimisation problem embedded in the PP guided tour, a
starting projection, an index function and the maximum number of
iterations are provided as input. Then, the current projection is
evaluated by the index function. The projection is then moved in a
direction determined by a random factor, influenced by how far along we
are in the optimization process. Occasionally, completely new directions
may be taken like a jellyfish might with ocean currents. A new
projection is accepted if it is an improvement compared to the current
one, rejected otherwise. This process continues and iteratively improves
the projection, until the pre-specified maximum number of trials is
reached.

\begin{tcolorbox}[enhanced jigsaw, toptitle=1mm, bottomtitle=1mm, bottomrule=.15mm, rightrule=.15mm, colbacktitle=quarto-callout-note-color!10!white, titlerule=0mm, leftrule=.75mm, coltitle=black, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, opacityback=0, colback=white, left=2mm, toprule=.15mm, arc=.35mm, colframe=quarto-callout-note-color-frame]

\textbf{Input}: \texttt{current\_projections}, \texttt{index\_function},
\texttt{trial\_id}, \texttt{max\_trial}

\textbf{Output}: \texttt{optimized\_projection}

\textbf{Initialize} \texttt{current\_best} as the projection with the
best index value from \texttt{current\_projections}, and
\texttt{current\_idx} as the array of index function values for each
projection in \texttt{current\_projections}

\textbf{for} each trial\_id in 1 to max\_tries \textbf{do}

\begin{quote}
Calculate \(c_t\) based on the current\_idx and max\_trial
\end{quote}

\begin{quote}
\textbf{if} \(c_t\) is greater than or equal to \(0.5\) \textbf{then}

\begin{quote}
Define trend based on the current\_best and current\_projections
\end{quote}

\begin{quote}
Update each projection towards the trend using a random factor and
orthonormalisation
\end{quote}

\textbf{else}

\begin{quote}
\textbf{if} a random number is greater than \(1 - c_t\) \textbf{then}

\begin{quote}
Slightly adjust each projection with a small random factor (passive)
\end{quote}

\textbf{else}

\begin{quote}
For each projection, compare with a random jelly and adjust towards or
away from it (active)
\end{quote}
\end{quote}

Update the orientation of each projection to maintain consistency

Evaluate the new projections using the index function
\end{quote}

\begin{quote}
\textbf{if} any new projection is worse than the current, revert to the
\texttt{current\_projections} for that case

\begin{quote}
Determine the projection with the best index value as the new
current\_best
\end{quote}
\end{quote}

\begin{quote}
\textbf{if} trial\_id \(\geq\) max\_trial, print the last best
projection \textbf{exit}
\end{quote}

\textbf{return} the set of projections with the updated current\_best as
the optimized\_projection

\end{tcolorbox}

The JSO implementation involves several key parameters that control its
search process in optimization problems. These parameters are designed
to guide the exploration and exploitation phases of the algorithm. While
the specific implementation details can vary depending on the version of
the algorithm or its application, we focus on two main parameters that
are most relevant to our application: the number of jellyfish and the
maximum number of trials.

\citet{laa_using_2020} has proposed five criteria for assessing
projection pursuit indexes (smoothness, squintability, flexibility,
rotation invariance, and speed). Since not all the properties affects
the execution of the optimisation, we choose to focus here on the first
two properties, namely \emph{smoothness} (Section~\ref{sec-smoothness})
and \emph{squintability} (Section~\ref{sec-squintability}), and propose
two metrics to evaluate them.

\subsection{Smoothness}\label{sec-smoothness}

This subsection proposes a metric for the smoothness of a projection
pursuit index.

If a PP index function is evaluated at some random points, as is done at
the initialization stage of the jellyfish search optimization, then
these random index values can be interpreted as a random field, indexed
by a space parameter, namely the random projection angle. This analogy
suggests to use this random training sample to fit a spatial model, a
simple one being a (spatial) Gaussian process.

The distribution of a Gaussian process is fully determined by its mean
function and covariance function. The smoothness property comes into
play in the definition of the covariance function: if a PP index is very
smooth, then two close projection angles should produce close index
values (strong correlation); by contrast, if a PP index is not very
smooth, then two close projection angles might give very different index
values (fast decay of correlations with respect to distance between
angles).

Popular covariance functions are parametric positive semi-definite
functions, some of which have a parameter to capture the smoothness of
the Gaussian field. In particular, consider the Matérn class of
covariance functions, defined by

\[
K(u):=\frac{(\sqrt{2\nu}u)^{\nu}}{\Gamma(\nu)2^{\nu-1}}\mathcal{K}_{\nu}(\sqrt{2\nu}u)\ ,
\]

where \(\nu>0\) is the smoothness parameter and where
\(\mathcal{K}_\nu\) is the modified Bessel function. The Matérn
covariance function can be expressed analytically when \(\nu\) is a
half-integer, the most popular values in the literature being \(1/2\),
\(3/2\) and \(5/2\). The parameter \(\nu\), called \emph{smoothness
parameter}, controls the decay of the covariance function. As such, it
is an appropriate measure of smoothness of a random field.

In our context, we suggest to use this parameter \(\nu\) as a measure of
the smoothness of the index function by fitting a Gaussian process prior
with Matérn covariance on a dataset generated by random evaluations of
the index function, as done at the initialization stage of the jellyfish
search optimization. There exist several R packages, such as GpGp
\citep{guinness2021gpgp} or ExaGeoStatR \citep{abdulah2023large}, to fit
the hyperparameters of a GP covariance function on data. In this
project, we make use of the GpGp package.

After fitting the smoothness parameter \(\nu>0\), its value can be
interpreted as follows: the higher \(\nu\), the smoother the index
function.

\subsection{Squintability}\label{sec-squintability}

Squintability has been a useful concept. Here, it is defined
mathematically and two approaches to compute this metric numerically,
are described.

In the existing literature, it is commonly understood that a large
squint angle implies that the objective function value is close to
optimal even when we are not very close to the perfect view to see the
structure. A small squint angle means that the PP index value improves
substantially only when we are very close to the perfect view. As such,
low squintability implies rapid improvement in the index value when near
the perfect view. For PP, a small squint angle is considered to be
undesirable because it means that the optimiser needs to be very close
to be able to ``see'' the optima. Thus, it could be difficult for the
optimiser to find the optima. We now propose a mathematical formulation
of this intuition.

Let \(A \in \mathbf{R}^{p \times 2}\) be the 2D projection basis and
\(A^*\) the optimal basis achieving the maximum index value for a given
projection pursuit problem defined by the shape to find, data dimension,
and the index function \(f(.)\). The projection distance \(d(A, A^*)\),
is defined as the Frobenius norm \(\lVert . \rVert _F\), between the
squared difference of these two bases:

\begin{equation}\phantomsection\label{eq-proj-dist}{d(A, A^*) = \lVert M \rVert _F = \sum_{ij} M_{ij}^2, \text{ where } M = AA^\prime - A^*A^{*\prime}\ .}\end{equation}

The index value \(f(XA)\) is defined as described in the introduction.
We now focus our attention to how the index value \(f(XA)\) changes with
respect to the projection distance \(d(A, A^*)\), over the course of the
JSO.

It is expected that for a PP index with high squint angle, the
optimization (\ref{eq-optimization}) should make substantial progress
early on. Conversely, for a PP index with low squint angle, it might
take a long while for the optimization to make substantial progress, as
the candidate projections would need to be very close to the optimal one
for the structure of the index function to be visible enough to be
amenable to efficient optimization. This observation suggests that the
extreme values of \(f'\) (the ones for which \(f''=0\), assuming that
\(f\) is twice differentiable), and the projection distances for which
these values are attained, should play an important role in the
mathematical definition of squintability. Noticing that \(f\) is
expected to be a decreasing function, we propose the following
squintability metric:

\begin{equation}\phantomsection\label{eq-squintability}{
\varsigma:=-4\min_{x}f'(x)\times\underset{x}{\arg\min}f'(x)\ .
}\end{equation}

On the one hand, \(-\min_{x}f'(x)\) represents the largest value of the
gradient of \(-f\). On the other hand, \(\underset{x}{\arg\min}f'(x)\)
represents the projection distance at which this largest gradient is
attained. Based on the above discussion, it is expected that these two
values should be both high in the case of high squintability (fast
increase in \(f\) early on), and both low in the case of low
squintability (any substantial increase in \(f\) happens very late,
close to the optimal angle). This suggest that their product
(\ref{eq-squintability}) should provide a sensible measure of
squintability. The multiplicative constant \(4\), which can be deemed
arbitrary, does not change the interpretation of the squintability
metric \(\varsigma\); it is here to adjust the range of values of
\(\varsigma\) and simplify the explicit formula for \(\varsigma\)
obtained later on.

To compute the squintability metric (\ref{eq-squintability}) in
practice, several approaches are possible. The first one is to propose a
parametric model for \(f\), and use it to obtain an explicit formula for
\(\varsigma\). Numerical experiments suggest a scaled sigmoid shape as
described below. Define

\begin{equation}\phantomsection\label{eq-logistic}{
\ell(x):=\frac{1}{1+\exp(\theta_{3}(x-\theta_{2}))}\ ,
}\end{equation}

which is a decreasing logistic function depending on two parameters
\(\theta_2\) and \(\theta_3\), such that
\(\ell(\theta_{2})=\frac{1}{2}\). Then, define

\begin{equation}\phantomsection\label{eq-parametric}{
f(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(x_{\max})}{\ell(0)-\ell(x_{\max})}+\theta_{4}\ ,
}\end{equation}

which depends on three additional parameters, \(\theta_1\),
\(\theta_2\), and \(x_{\max}\), such that \(f(0)=\theta_1\) and
\(f(x_{\max})=\theta_4\). Under the parametric model
(\ref{eq-parametric}), the squintability metric (\ref{eq-squintability})
can be shown to be equal to

\begin{equation}\phantomsection\label{eq-squintability-parametric}{
\varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}\ .
}\end{equation}

In practice, the parameters of this model (\ref{eq-parametric}) can be
estimated numerically, for example by non-linear least squares, and then
used to evaluate \(\varsigma\) as in equation
(\ref{eq-squintability-parametric}).

Alternatively, one can estimate (\ref{eq-squintability}) in a
nonparametric way, for example by fitting \(f\) using kernel regression,
then numerically estimate the angle at which \(-f'\) attains its highest
value.

\section{Simulation details}\label{sec-sim-deets}

The JSO performance is compared with the existing Creeping Random Search
(CRS) currently used in guided tour, and explores how JSO behaves with
different hyper-parameters combinations. A second simulation studies the
effect of index properties (smoothness and squintability), along with
JSO hyper-parameters, on the success rate of the JSO performance.

\subsection{Performance of JSO in pipe-finding
problems}\label{sec-app-1}

The performance of JSO is investigated in two folds: 1) compare it with
a commonly used optimiser: the Csreeping Random Search (CRS)
\citep{RJ-2021-105, laa_using_2020}, and 2) examine its success rate
under two hyper-parameters (number of jellyfishes and maximum number of
tries). The projection pursuit problem used is finding the pipe shape
using the holes index, investigated by \citet{laa_using_2020}. Fifty
simulations are conducted with both JSO and CRS, in four dimensions
(\(d = 6, 8, 10, 12\)). JSO uses 100 jellyfishes and 100 maximum number
of tries, while the CRS allows a maximum of 1000 samples at each
iteration before the algorithm terminates. The results of the simulation
are collected using the data structure proposed in \citet{RJ-2021-105}
is used here for assessing JSO. A sample for the sine-wave structure in
6D data using a spline index function (\texttt{splines2d}) is printed
below:

\begin{verbatim}
Rows: 10
Columns: 12
$ index     <chr> "splines2d", "splines2d", "splines2d", "splines2d", "splines~
$ d         <dbl> 6, 6, 6, 6, 6, 6, 6, 6, 6, 6
$ n_jellies <dbl> 50, 50, 50, 50, 50, 50, 50, 50, 50, 50
$ max_tries <dbl> 50, 50, 50, 50, 50, 50, 50, 50, 50, 50
$ sim       <int> 1, 1, 1, 1, 1, 1, 1, 1, 1, 1
$ seed      <int> 3462, 3462, 3462, 3462, 3462, 3462, 3462, 3462, 3462, 3462
$ basis     <list> <<matrix[6 x 2]>>, <<matrix[6 x 2]>>, <<matrix[6 x 2]>>, <<m~
$ index_val <dbl> 0.036664532, 0.005121912, 0.058683104, 0.056527991, 0.015734~
$ method    <chr> "search_jellyfish", "search_jellyfish", "search_jellyfish", ~
$ tries     <dbl> 1, 1, 1, 1, 1, 1, 1, 1, 1, 1
$ loop      <dbl> 1, 2, 3, 4, 5, 6, 7, 8, 9, 10
$ time      <drtn> 4.362133 mins, 4.362133 mins, 4.362133 mins, 4.362133 mins, ~
\end{verbatim}

In this simulation, 50 jellyfishes (\texttt{n\_jellies}) and a maximum
of 50 tries (\texttt{max\_tries}) are used. The full data contains 50
simulations, indexed by the variable \texttt{sim}, with the associated
\texttt{seed}s. Recorded in columns \texttt{basis}, \texttt{index\_val},
\texttt{tries}, \texttt{loop} and \texttt{time}, are the sampled
projection basis, index value, iteration ID, jellyfish ID, and
computational time, respectively.

In the second experiment, fifty simulations are conducted at each
hyper-parameter combination below to analyse its effects on the JSO
success rate: number of jellyfish: 20, 50, and 100, and maximum number
of tries: 50 and 100). The success rate is defined as the proportion of
simulations that achieves a final index value within 0.05 of the best
index value found among all 50 simulations.
Figure~\ref{fig-success-rate} illustrates the calculation based on the
projection found by the above simulation. Here, the optimal index value
is 0.999. Out of the 50 simulations, 44 achieved an index value within
{[}0.949, 0.999{]}, resulting in a success rate of 0.88.

\begin{figure}

\centering{

\includegraphics[width=4.49in,height=\textheight]{figures/success-rate.png}

}

\caption{\label{fig-success-rate}Illustration of success rate
calculation: The best projections identified by the JSO in 50
simulations are sorted by their index values. The highest index value
found across all simulation is 0.999. Out of the 50 simulations, 44
achieved an index value within 0.05 of the best, resulting in a success
rate of 0.88 (44/50).}

\end{figure}%

\subsection{Factors affecting JSO success rate: index properties and
jellyfish hyper-parameters}\label{sec-app-2}

In addition to the pipe-finding problem, a new shape, sine wave, is
investigated in 6D and 8D spaces with six indexes considered:
\texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
\texttt{spline}, and \texttt{stringy}. Combining with two JSO
hyper-parameters, a total of 52 cases is produced, comprising of 24
pipe-finding cases and 28 sine-wave finding cases. For each case, JSO is
run fifty times and the success rate is calculated.

Smoothness and squintability are computed for each case, following the
procedures outlined in Section~\ref{sec-smoothness} and
Section~\ref{sec-squintability} and illustrated in
Figure~\ref{fig-smoothness} and Figure~\ref{fig-squintability}. For
smoothness, three hundred random bases are simulated. Index values and
projection distance (to the optimal basis) are calculated for each
random basis before fitting into a Gaussian process model to obtain the
smoothness measure. For squintability, fifty random bases are sampled
and interpolated to the optimal basis with a step size of 0.005. Index
values and projection distances are calculated for these interpolated
bases and the index values are averaged with a bin width of 0.005. The
squintability measure is then calculated from fitting a 4-parameter
scaled logistic function of index value against projection distance,
estimated by non-linear least squares.

Combining with the success rate calculated for each case, the dataset is
fitted using a generalised linear model with a binomial family and a
logit link function with three pre-processing steps are applied: 1)
scaling the JSO hyper-parameters by a factor of 10 for interpretation,
2) creating a new binary variable \texttt{long\_time} to indicate cases
with an average run time over 30 seconds, and 3) encoding the success
rate for the stringy case as 0, because upon visual inspection of the
final projection, none of the 50 simulations identified the sine-wave
shape.

\begin{figure}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{figures/smoothness.png}

}

\caption{\label{fig-smoothness}Illustration of steps to calculate
smoothness. For a given projection pursuit problem defined by the shape
to find, data dimension and the index function, sample random bases
given the orthonormality contraint. Calculate the projection distance
and the index value for each random basis. Fit a Gaussian process model
to the index values against the projection distances to obtain the
smoothness measure.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{figures/squintability.png}

}

\caption{\label{fig-squintability}Illustration of steps to calculate
squintability. For a given projection pursuit problem defined by the
shape to find, data dimension and the index function, sample random
bases given the orthonormality and projection distance contraint.
Interpolate the sampled bases to the optimal basis and calculate the
projection distance and the index value for each interpolated basis. Bin
the index value by the projection distance, then claculate the average
index value for each bin. Fit the scaled sigmoid function from Equation
\ref{eq-squintability} to the binned index values against the projection
distance using non-linear least square. Use Equation \ref{eq-parametric}
to calculate the squintability measure from the estimated \(\theta_1\)
to \(\theta_4\).}

\end{figure}%

\section{Results {[}Di and Sherry{]}}\label{sec-sim-res}

The results from the two simulations described in
Section~\ref{sec-sim-deets} are analysed based on the final projections
across two optimisers (JSO and CRS) and the success rate across JSO
hyper-parameters. Smoothness and squintability are calculated across a
collection of pipe-finding and sine-wave finding problems to construct
the relationship between success rate, JSO hyper-parameters, and index
properties.

\subsection{Performance of JSO in pipe-finding
problems}\label{sec-app-1}

The final projections found by the two optimisers (JSO and CRS) are
presented in Figure~\ref{fig-proj}, broken down by 10th quantile,
faceted by the data dimension. The JSO uses 100 jellyfishes and a
maximum number of tries of 100. The CRS uses a maximum of 1000 tries in
each step of random sampling step before the algorithm terminates. In
the 6-dimensional data scenario, JSO consistently identifies a clear
pipe shape. The CRS also finds the pipe shape but with a wide rim,
suggesting a further polish search may be required. With increasing
dimensions, JSO may not always identify the pipe shape due to random
sampling, but it still presents a pipe shape in over 50\% of cases. When
compared to the CRS, JSO reaches higher index values and clearer pipe
shapes across all quantiles. This suggests JSO can be more effective
when searching higher dimension spaces, which may be preferred by users.

The success rate calculated for each hyper-parameter combination is
presented in Figure~\ref{fig-proportion}. As the number of jellyfishes
and maximum tries increase, the success rate also increases. For simpler
problems (6 dimensions), small parameter values (20 jellyfishes and a
maximum of 50 tries) can achieve a high success rate. However, for
higher-dimensional problems (8, 10, and 12 dimensions), larger parameter
values (i.e.~100 jellyfishes and a maximum of 100 tries) are necessary.
Increasing both parameters enhances the performance of JSO, but it also
extends the computational time required for the optimisation, which can
be computationally intensive when evaluating the index function (such as
scagnostic indexes) multiple times across numerous iterations.

\begin{figure}

\centering{

\includegraphics{optim_files/figure-pdf/fig-proj-1.pdf}

}

\caption{\label{fig-proj}Projections found by the JSO and CRS at each
10th quantile across 50 simulations. The projection pursuit problem is
to find the pipe shape using the holes index in the 6, 8, 10, and
12-dimensional spaces. The JSO uses 100 jellyfishes and a maximum number
of tries of 100. The CRS uses a maximum of 1000 tries in each step of
random sampling step before the algorithm terminates. In the 6-D data
space, JSO always finds a clear pipe shape while the CRS also finds the
pipe shape but with a wide rim. At higher data dimensions, JSO finds a
higher index value and a clearer pipe shape across all the quantiles
than the CRS}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{optim_files/figure-pdf/fig-proportion-1.pdf}

}

\caption{\label{fig-proportion}Proportion of simulations reaches
near-optimal index values in the pipe-finding problem using the holes
index. The proportion is calculated based on the number of simulations,
out of 50, that achieve an index value within 0.05 of the
best-performing simulation. As the dimensionality increases, the
proportion of simulations reaching the optimal index value increases.}

\end{figure}%

\subsection{Factors affecting JSO success rate: index properties and
jellyfish hyper-parameters}\label{sec-app-2}

The index properties, including smoothness and squintability, offer
numerical metrics to characterise the complexity of projection pursuit
optimisation problems. Table~\ref{tbl-smoothness-squintability} presents
the parameters estimated from the Gaussian process (variance, range,
smooth, and nugget) and the scaled logistic function (\(\theta_1\) to
\(\theta_4\)) for calculating both metrics for each case considered in
the simulation. The column ``smooth'' is used as the smoothness metric
and the column ``squint'' is calculated as equation
\ref{eq-squintability-parametric} as the squintability metric.
Table~\ref{tbl-mod-output} presents the results of fitting a generalised
linear model with a binomial family and a logit link function to the
data in Table~\ref{tbl-mod-data}, where all the three components
(success rate, JS hyper-parameters, and index properties) are combined.
The model suggests that JSO success rate is positively associated with
the two hyper-parameters, as well as with the index properties:
smoothness and squintability. Specifically, using 10 more jellyfishes
and 10 more tries increases the odd ratio of success by 24.11\% and
11.93\%, respectively. However, being flagged with long runtime and an
increase of data dimension reduce the success rate by 41.72\% and
53.36\%, respectively. The variable \texttt{squintability} and
\texttt{dimension} are significant, suggesting their importance relative
to JSO hyper-parameters in the optimisation success. In defining a
projection pursuit problem, factors such as the shape-to-find, the index
function used, and the data dimension, determine the properties like
smoothness and squintability. These metrics can then be compared across
different problems to understand their relative complexity. The results
suggest that squintability has a more significant influence than
smoothness on the success rate of JSO. Once the characteristics of the
projection pursuit problem are fully understood, increasing JSO
hyper-parameters can enhance the search effectiveness, however, it is
also important to consider the resulting increase in computational
complexity.

\begin{table}

\caption{\label{tbl-smoothness-squintability}Parameters estimated from
the Gaussian process (including variance, range, smoothness, and nugget)
and scaled logistic function (\(\theta_1\) to \(\theta_4\)) for the
pipe-finding and sine-wave finding problems. The column ``smooth'' and
``squint'' represent the smoothness and squintability measures.}

\centering{

\centering\begingroup\fontsize{10}{12}\selectfont

\begin{tabular}{|>{}lll>{}r|rrr>{}r|rrrr>{}r|}
\toprule
  & shape & index & d & variance & range & smooth & nugget & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\theta_4$ & squint\\
\midrule
1 & pipe & holes & 6 & 0.00 & 0.37 & 2.36 & 0.22 & 1.02 & 0.86 & 3.37 & 0.02 & 3.05\\
2 & pipe & holes & 8 & 0.00 & 0.18 & 2.19 & 0.82 & 1.01 & 0.87 & 3.26 & 0.03 & 2.96\\
3 & pipe & holes & 10 & 0.00 & 0.11 & 2.19 & 3.03 & 1.02 & 0.88 & 3.15 & 0.02 & 2.95\\
4 & pipe & holes & 12 & 0.00 & 0.15 & 2.29 & 1.58 & 1.01 & 0.88 & 3.34 & 0.00 & 3.12\\
5 & sine & MIC & 6 & 0.02 & 0.09 & 2.46 & 0.08 & 0.89 & 0.57 & 1.62 & -0.02 & 1.26\\
6 & sine & MIC & 8 & 0.02 & 0.08 & 2.64 & 0.08 & 0.93 & 0.33 & 1.31 & -0.03 & 0.77\\
7 & sine & TIC & 6 & 0.12 & 0.11 & 2.44 & 0.09 & 0.95 & 0.54 & 1.72 & -0.03 & 1.32\\
8 & sine & TIC & 8 & 0.12 & 0.10 & 2.47 & 0.09 & 0.95 & 0.56 & 1.72 & -0.03 & 1.37\\
9 & sine & dcor2d & 6 & 0.03 & 0.17 & 2.66 & 0.11 & 0.95 & 1.04 & 2.74 & -0.02 & 2.95\\
10 & sine & loess2d & 6 & 0.08 & 0.34 & 1.99 & 0.31 & 1.02 & 1.04 & 2.65 & 0.08 & 2.76\\
11 & sine & splines2d & 6 & 0.04 & 0.24 & 2.54 & 0.10 & 1.01 & 1.05 & 2.73 & -0.01 & 3.12\\
12 & sine & stringy & 6 & 0.00 & 0.01 & 1.54 & 38.17 & 1.05 & 0.01 & 254.73 & 0.05 & 2.96\\
\bottomrule
\end{tabular}
\endgroup{}

}

\end{table}%

\begin{table}

\caption{\label{tbl-mod-data}The first 7 rows of the datasets processed
for modelling.}

\centering{

\centering\begingroup\fontsize{10}{12}\selectfont

\begin{tabular}{lrrrrrrr}
\toprule
Index & D & Smoothness & Squintability & N. jellyfish & Max. tries & Success rate & Time (sec)\\
\midrule
MIC & 6 & 2.46 & 1.26 & 20 & 50 & 0.12 & 2.48\\
MIC & 6 & 2.46 & 1.26 & 20 & 100 & 0.24 & 8.95\\
MIC & 6 & 2.46 & 1.26 & 50 & 50 & 0.52 & 5.65\\
MIC & 6 & 2.46 & 1.26 & 50 & 100 & 0.64 & 13.22\\
MIC & 6 & 2.46 & 1.26 & 100 & 50 & 0.76 & 19.45\\
MIC & 8 & 2.64 & 0.77 & 20 & 50 & 0.08 & 2.57\\
MIC & 8 & 2.64 & 0.77 & 20 & 100 & 0.08 & 4.96\\
\bottomrule
\end{tabular}
\endgroup{}

}

\end{table}%

\begin{table}

\caption{\label{tbl-mod-output}Model estimates of proportion of
jellyfish success on index properties and jellyfish hyper-parameters
from the generalised linear model with a binomial family and a logit
link function. The variable smoothness, squintability, number of
jellyfish and maximum number of tries are positively associated with JSO
success rate while data dimension and being flagged as long runtime are
negatively associated with the success rate. The variable squintability
and dimension are significant, suggesting their importance relative to
jellyfish hyper-parameters in the optimisation success.}

\centering{

\centering\begingroup\fontsize{10}{12}\selectfont

\begin{tabular}{|>{}lrrr>{}r|}
\toprule
term & estimate & std.error & statistic & p.value\\
\midrule
Intercept & -4.52 & 5.33 & -0.85 & 0.40\\
Smoothness & 1.19 & 1.92 & 0.62 & 0.53\\
Squintability & 2.06 & 0.68 & 3.01 & 0.00\\
Dimension (d) & -0.63 & 0.26 & -2.46 & 0.01\\
Long time & -0.87 & 1.29 & -0.68 & 0.50\\
N. jellyfish & 0.22 & 0.13 & 1.70 & 0.09\\
Max. tries & 0.11 & 0.15 & 0.75 & 0.45\\
\bottomrule
\end{tabular}
\endgroup{}

}

\end{table}%

\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}

This paper has presented new metrics to mathematically define desirable
features of PP indexes, squintability and smoothness, and used these to
assess the performance of the new jellyfish search optimiser. The
metrics will be generally useful for characterising PP indexes, and halp
with developing nex indexes.

In the comparison of the JSO against the currently used CRS, as expected
the JSO vastly out-performs CRS, and provides a high probability of
finding the global optima. The JSO obtains the maximum more cleanly,
with a slightly higher index value, and plot of the projected data
showing the structure more clearly.

The JSO performance is affected by the hyper-parameters, with a higher
chance of reaching the global optima when more jellies are used and the
maximum number of tries is increased. However, it comes at a
computational cost, as expected. The performance declines if the
projection dimension increases and if the PP index has low
squintability. The higher the squintability the better chance the JSO
can find the optima. However, interestingly smoothness does not affect
the JSO performance.

The new JSO is integrated with the current implementation of the
projection pursuit guided tour in the \texttt{tourr} package, and can be
further examined using PP optimisation diagnostics in the \texttt{ferrn}
package. Using the JSO is a little different than the current CRS,
because it runs many paths. The recommended approach is to conduct the
optimisation off-line, and then use the guided tour to follow selected
jellies. The tools for this are available in the \texttt{tourr} package,
too.

\renewcommand\refname{References}
  \bibliography{bibliography.bib}

\end{document}