### 2024-07-01 optim.qmd
small editsg

---
title: Studying the Performance of the Jellyfish Search Optimiser for the Application of Projection Pursuit
header-includes:
   - \usepackage{algorithm}
author:
  - name: H. Sherry Zhang
    email: huize.zhang@austin.utexas.edu
    affiliations: 
        - id: 1
          name: University of Texas at Austin
          department: Department of Statistics and Data Sciences
          city: Austin
          country: United States
          postal-code: 78751
    attributes:
        corresponding: true
  - name: Dianne Cook
    email: dicook@monash.edu
    affiliations:
        - id: 2
          name: Monash University
          department: Department of Econometrics and Business Statistics
          city: Melbourne
          country: Australia
          postal-code: 3800
  - name: Nicolas  Langrené
    email: nicolaslangrene@uic.edu.cn
    affiliations:
        - id: 3
          name: Guangdong Provincial Key Laboratory of Interdisciplinary Research and Application for Data Science, BNU-HKBU United International College
          department: Department of Mathematical Sciences
          city: Zhuhai
          country: China
          postal-code: 519087
  - name: Jessica Wai Yin Leung
    email: Jessica.Leung@monash.edu
    affiliations:
        - id: 2
          name: Monash University
          department: Department of Econometrics and Business Statistics
          city: Melbourne
          country: Australia
          postal-code: 3800
abstract: |
  The projection pursuit (PP) guided tour interactively optimises a criteria function known as the PP index, to explore high-dimensional data by revealing interesting projections. The optimisation in PP can be non-trivial, involving non-smooth functions and optima with a small "squint angle", detectable only from close proximity. To address these challenges, this study investigates the performance of a recently introduced swarm-based algorithm, Jellyfish Search Optimiser (JSO), for optimising PP indexes. The performance of JSO for visualising data is evaluated across various hyper-parameter settings and compared with existing optimisers. Additionally, this work proposes novel methods to quantify two properties of the PP index -- smoothness and squintability -- that capture the complexities inherent in PP optimisation problems. These two metrics are evaluated along with JSO hyper-parameters to determine their effects on JSO success rate. Our numerical results confirm the positive impact of these metrics on the JSO success rate, with squintability being the most significant. The JSO algorithm has been implemented in the `tourr` package and functions to calculate smoothness and squintability of indexes are available in the `ferrn` package. 
keywords: 
  - projection pursuit
  - jellyfish search optimiser (JSO)
  - optimisation
  - grand tour
  - high-dimensional data
  - exploratory data analysis
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal of Multivariate Analysis
      formatting: preprint
      model: 3p
      cite-style: numbername
editor: 
  markdown: 
    wrap: 72
crossref: 
  eq-prefix: ""
---

```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(ggplot2)
library(patchwork)
library(ggh4x)
library(broom)
library(kableExtra)
library(ferrn)
load(here::here("data/sim_pipe_run_best.rda"))
load(here::here("data/sim_sine_6d_spline_head.rda"))
load(here::here("data/sim_sine_6d_spline_best.rda"))
load(here::here("data/sim_sine_6d_spline_projdist.rda"))
load(here::here("data/pipe_better.rda"))
load(here::here("data/pipe_jellyfish.rda"))
load(here::here("data/smoothness.rda"))
load(here::here("data/squintability.rda"))
load(here::here("data/sq_basis_dist_idx.rda"))
load(here::here("data/sim_df.rda"))
```

# Introduction

Projection Pursuit (PP) (@kr69, @FT74) is a dimension reduction
technique aimed at identifying informative linear projections of data.
The method involves optimising a objective function known as the PP
index (e.g. @hall1989polynomial, @cook1993projection,
@lee2010projection, @Loperfido2018, @Loperfido2020), which defines the
criteria for what constitutes interesting or informative projections.
Given high-dimensional data $X \in \mathbf{R}^{n\times p}$ and an index
function $f(\cdot)$, PP finds the orthonormal basis
$A \in \mathbf{R}^{p \times d}$ that maximises the index value of the
projection, $Y = XA$:

$$
\underset{A}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I
$$ {#eq-optimization}

The index functions can be non-linear and non-convex, hence an effective
and efficient optimisation procedure is essential to explore the data
landscape and achieving a globally optimal viewpoint of the data.

Significant efforts have been devoted to enhancing the functionality of
PP, improving optimisation (e.g. @posse1995, @marie-sainte2010,
@grochowski2011) and extending its use to visually explore
high-dimensional data. @cook1995grand introduced the PP guided tour,
which monitors the optimisation visually to see the data projections
leading in and out of the optima. An implementation is available in the
R [@R] package `tourr` [@tourr]. @RJ-2021-105 illustrated how to
diagnose optimisation processes, particularly focusing on the guided
tour, and revealed a need for improved optimisation. While improving
the quality of the optimisation solutions in the tour is essential, it
is also important to be able to view the data projections as the
optimisation progresses. Integrating the guided tour with a global
optimisation algorithm that is efficient in finding the global optimal
and enables viewing of the projected data during the exploration process
is a goal.

Here, the potential for a Jellyfish Search Optimiser (JSO) to be
integrated with the projection pursuit guided tour is explored. JSO
[@chou_novel_2021] -[@rajwar_exhaustive_2023], inspired by the search
behaviour of jellyfish in the ocean, is a swarm-based metaheuristic
designed to solve global optimisation problems. Compared to traditional
methods, JSO demonstrated stronger search ability and faster
convergence, and requires fewer tuning parameters. These practical
advantages make JSO a promising candidate for enhancing PP optimisation.
<!--The study reported here explores JSO's potential in enhancing projection pursuit and examines the characteristics
and advantages of such implementation.-->

The primary goal of the study reported here is to investigate the
performance of JSO in PP optimisation for the guided tour. It is of
interest to assess how quickly and closely the optimiser reaches a gobal
optima, for various PP indexes that may have differing complexities. To
observe the performance of JSO with different types of PP indexes,
metrics are introduced to capture specific properties of the index
including squintability (based on @barnett1981interpreting's squint
angle) and smoothness as introduced in @laa_using_2020. Here we
mathematically define metrics for squintability and smoothness, which is
new for the field. A series of simulation experiments using specially
structured datasets and various PP indexes are conducted to assess the
behaviour of JSO under different settings and its sensitivity to
hyper-parameter choices (number of jellyfishes and maximum number of tries).
The relationship between the JSO performance, hyper-parameter tuning and
properties of PP indexes is analysed to provide guidance for
practitioners using projection pursuit, and developers of new indexes.

The paper is structured as follows. @sec-background introduces the
background of PP guided tour and reviews existing optimisers and index
functions in the literature. @sec-theory details JSO and introduces
metrics that measure different properties of PP indexes. @sec-sim-deets
details two simulation experiments to assess JSO's performance: one
comparing JSO's performance improvements relative to an existing
optimiser, Creeping Random Search (CRS), and the other studying the
impact of different PP index properties on optimisation performance.
@sec-sim-res presents the results. @sec-conclusion summarises the work
and provides suggestions for future directions.

# Projection pursuit, tours, index functions and optimisation {#sec-background}

<!-- Need to assume that the special issue will have the general introduction to PP, so the focus here will be on background needed for this paper, tours, guidance using PP, optimisation, visualisation, ... -->

A tour on high-dimensional data is constructed by geodesically
interpolating between pairs of planes. Any plane is described by an
orthonormal basis, $A_t$, where $t$ represents time in the sequence. The
term "geodesic" refers to maintaining the orthonormality constraint so
that each view shown is correctly a projection of the data. The PP
guided tour operates by geodesically interpolating to target planes
(projections) which have high PP index values, as provided by the
optimiser. The geodesic interpolation means that the viewer sees a
continuous sequence of projections of the data, so they can watch
patterns of interest forming as the function is optimised. There are
five (unsophisticated) optimisation methods implemented in the `tourr`
package:

-   `search_geodesic()`: provides a pseudo-derivative optimisation. It
    searches locally for the best direction, based on differencing the
    index values for very close projections. Then it follows the
    direction along the geodesic path between planes, stopping when the
    next index value fails to increase.
-   `search_better()`: is a brute-force optimisation searching randomly
    for projections with higher index values.
-   `search_better_random()`: is essentially simulated annealing
    [@Bertsimas93] where the search space is reduced as the optimisation
    progresses.
-   `search_posse()`: implements the algorithm described in @posse95.
-   `search_polish()`: is a very localised search, to take tiny steps to
    get closer to the local maximum.

There are several PP index functions available: `holes()` and `cmass()`
[@cook1993projection]; `lda_pp()` [@lee2005projection]; `pda_pp()`
[@lee2010projection]; `dcor2d()` and `splines2d()` [@Grimm2016];
`norm_bin()` and `norm_kol()` [@huber85]; `slice_index()`
[@Laa:2020wkm]. Most are relatively simply defined, for any projection
dimension, and implemented because they are relatively easy to optimise.
A goal is to be able to incorporate more complex PP indexes, for example
based on scagnostics (@scag, @WW08).

An initial investigation of PP indexes, and the potential for
scagnostics is described in @laa_using_2020. To be useful here an
optimiser needs to be able to handle index functions which are possibly
not very smooth. In addition, because data structures might be
relatively fine, the optimiser needs to be able to find maxima that
occur with a small squint angle, that can only be seen from very close
by. One last aspect that is useful is for an optimiser to return local
maxima in addition to global because data can contain many different and
interesting features.

# The jellyfish optimiser and properties of PP indexes {#sec-theory}

JSO mimics the natural movements of jellyfish, which include passive and
active motions driven by ocean currents and their swimming patterns,
respectively. In the context of optimization, these movements are
abstracted to explore the search space in a way that balances
exploration (searching new areas) and exploitation (focusing on
promising areas). The algorithm aims to find the optimal solution by
adapting the behaviour of jellyfish to navigate towards the best
solution over iterations [@chou_novel_2021].

To solve the optimisation problem embedded in the PP guided tour, a
starting projection, an index function and the maximum number of
iterations are provided as input. Then, the current projection is
evaluated by the index function. The projection is then moved in a
direction determined by a random factor, influenced by how far along we
are in the optimization process. Occasionally, completely new directions
may be taken like a jellyfish might with ocean currents. A new
projection is accepted if it is an improvement compared to the current
one, rejected otherwise. This process continues and iteratively improves
the projection, until the pre-specified maximum number of trials is
reached.

::: {.callout-note icon="false"}
## Algorithm: Jellyfish Optimizer Pseudo Code

**Input**: `current_projections`, `index_function`, `trial_id`,
`max_trial`

**Output**: `optimized_projection`

**Initialize** `current_best` as the projection with the best index
value from `current_projections`, and `current_idx` as the array of
index function values for each projection in `current_projections`

**for** each trial_id in 1 to max_tries **do**

> Calculate $c_t$ based on the current_idx and max_trial

> **if** $c_t$ is greater than or equal to $0.5$ **then**
>
> > Define trend based on the current_best and current_projections
>
> > Update each projection towards the trend using a random factor and
> > orthonormalisation
>
> **else**
>
> > **if** a random number is greater than $1 - c_t$ **then**
> >
> > > Slightly adjust each projection with a small random factor
> > > (passive)
> >
> > **else**
> >
> > > For each projection, compare with a random jelly and adjust
> > > towards or away from it (active)
>
> Update the orientation of each projection to maintain consistency
>
> Evaluate the new projections using the index function

> **if** any new projection is worse than the current, revert to the
> `current_projections` for that case
>
> > Determine the projection with the best index value as the new
> > current_best

> **if** trial_id $\geq$ max_trial, print the last best projection
> **exit**

**return** the set of projections with the updated current_best as the
optimized_projection
:::

<!-- Var names in the function and in the text needs to be consistent??? -->

The JSO implementation involves several key parameters that control its
search process in optimization problems. These parameters are designed
to guide the exploration and exploitation phases of the algorithm. While
the specific implementation details can vary depending on the version of
the algorithm or its application, the focus is on two main parameters that
are most relevant to our application: the number of jellyfish and the
maximum number of trials.

@laa_using_2020 has proposed five criteria for assessing projection
pursuit indexes (smoothness, squintability, flexibility, rotation
invariance, and speed). Since not all the properties affects the
execution of the optimisation, the focus here is on the first two
properties, namely *smoothness* (@sec-smoothness) and *squintability*
(@sec-squintability), and propose two metrics to evaluate them.

## Smoothness {#sec-smoothness}

```{=html}
<!--
An intuitive way to measure smoothness of a function would be to count
how many continuous derivatives exist. 
To help define smoothness in our
context, we can make use of the Sobolev spaces: functions $f$ in the
Sobolev space $W^{p,\infty}$ have all derivatives of order less than $p$
continuous. Smoothness would then be the highest $p$ such that the index
function belongs to $W^{p,\infty}$. We can relax this and consider
$W^{p,q}$ Sobolev spaces.

Consider the following definition of partial derivative. Let $U$ be an
open subset of $\mathbf{R}^n$ and $f: U \rightarrow \mathbf{R}$. The
partial derivative of function $f$ at the point
$\mathbf{a} = (a_1, \ldots, a_n) \in U$ with respective to the $i$-th
variable $x_i$ is defined as

$$
\frac{\delta}{\delta x_i} f(\mathbf{a}) =  \lim_{h \rightarrow 0} \frac{f(a_1, \ldots, a_{i-1}, a_i +h, a_{i+1}, \ldots, a_n) - f(a_1 , \ldots, a_i, \ldots, a_n)}{h} = \lim_{h\rightarrow 0} \frac{f(\mathbf{a} + h\mathbf{e_i}) - f(\mathbf{a})}{h}.
$$ where $\mathbf{e_i}$ is the unit vector of the $i$-th variable $x_i$.

If the derivative is not well-defined, we propose to approximate the
derivative using the following expression:

$$\frac{1}{n_i}\sum_{i} \frac{\|f(\mathbf{a} + h_i\mathbf{e_i}) - f(\mathbf{a})\|^p}{h_i}$$

where $p \in [0,1]$. The choice of $p$ reflects the penalty behaviour.
For the application in this paper, we choose $p = 1$.

-   $h_i$ is an neighbourhood area around $a$. We can insert different
    values of $h$ to approximate the above quantity and observe how this
    quantity changes.
-->
```
This subsection proposes a metric for the smoothness of a projection
pursuit index.

If a PP index function is evaluated at some random points, as is done at
the initialization stage of the jellyfish search optimization, then
these random index values can be interpreted as a random field, indexed
by a space parameter, namely the random projection angle. This analogy
suggests to use this random training sample to fit a spatial model, a
simple one being a (spatial) Gaussian process.

The distribution of a Gaussian process is fully determined by its mean
function and covariance function. The smoothness property comes into
play in the definition of the covariance function: if a PP index is very
smooth, then two close projection angles should produce close index
values (strong correlation); by contrast, if a PP index is not very
smooth, then two close projection angles might give very different index
values (fast decay of correlations with respect to distance between
angles).

Popular covariance functions are parametric positive semi-definite
functions, some of which have a parameter to capture the smoothness of
the Gaussian field. In particular, consider the Matérn class of
covariance functions, defined by

$$
K(u):=\frac{(\sqrt{2\nu}u)^{\nu}}{\Gamma(\nu)2^{\nu-1}}\mathcal{K}_{\nu}(\sqrt{2\nu}u)\ ,
$$

where $\nu>0$ is the smoothness parameter and where $\mathcal{K}_\nu$ is
the modified Bessel function. The Matérn covariance function can be
expressed analytically when $\nu$ is a half-integer, the most popular
values in the literature being $1/2$, $3/2$ and $5/2$. The parameter
$\nu$, called *smoothness parameter*, controls the decay of the
covariance function. As such, it is an appropriate measure of smoothness
of a random field.

In this context, the parameter $\nu$ is suggested as a measure of
the smoothness of the index function by fitting a Gaussian process prior
with Matérn covariance on a dataset generated by random evaluations of
the index function, as done at the initialization stage of the jellyfish
search optimization. There exist several R packages, such as GpGp
[@guinness2021gpgp] or ExaGeoStatR [@abdulah2023large], to fit the
hyperparameters of a GP covariance function on data. In this project, we
make use of the GpGp package.

After fitting the smoothness parameter $\nu>0$, its value can be
interpreted as follows: the higher $\nu$, the smoother the index
function.

## Squintability {#sec-squintability}

Squintability has been a useful concept. Here, it is defined
mathematically and two approaches to compute this metric numerically,
are described.

In the existing literature, it is commonly understood that a large
squint angle implies that the objective function value is close to
optimal even when the perfect view to see the structure is far away.
A small squint angle means that the PP index value improves
substantially only when the perfect view is close by. As such,
low squintability implies rapid improvement in the index value when near
the perfect view. For PP, a small squint angle is considered to be
undesirable because it means that the optimiser needs to be very close
to be able to "see" the optima. Thus, it could be difficult for the
optimiser to find the optima. The mathematical formulation
of this intuition is proposed below:

Let $A \in \mathbf{R}^{p \times 2}$ be the 2D projection basis and $A^*$
the optimal basis achieving the maximum index value for a given
projection pursuit problem defined by the shape to find, data dimension,
and the index function $f(.)$. The projection distance $d(A, A^*)$, is
defined as the Frobenius norm $\lVert . \rVert _F$, between the squared
difference of these two bases:

$$d(A, A^*) = \lVert M \rVert _F = \sum_{ij} M_{ij}^2, \text{ where } M = AA^\prime - A^*A^{*\prime}\ .$$ {#eq-proj-dist}

The index value $f(XA)$ is defined as described in the introduction. We
now focus our attention to how the index value $f(XA)$ changes with
respect to the projection distance $d(A, A^*)$, over the course of the
JSO.

```{=html}
<!--
\[We generate random points that is beyond 1.5 projection distance and
interpolate. Then we fit a kernel or use nonlinear least squares.\]
-->
```
It is expected that for a PP index with high squint angle, the
optimization (@eq-optimization) should make substantial progress early
on. Conversely, for a PP index with low squint angle, it might take a
long while for the optimization to make substantial progress, as the
candidate projections would need to be very close to the optimal one for
the structure of the index function to be visible enough to be amenable
to efficient optimization. This observation suggests that the extreme
values of $f'$ (the ones for which $f''=0$, assuming that $f$ is twice
differentiable), and the projection distances for which these values are
attained, should play an important role in the mathematical definition
of squintability. Noticing that $f$ is expected to be a decreasing
function, the squintability metric is proposed as follows:

$$
\varsigma:=-4\min_{x}f'(x)\times\underset{x}{\arg\min}f'(x)\ .
$$ {#eq-squintability}

On the one hand, $-\min_{x}f'(x)$ represents the largest value of the
gradient of $-f$. On the other hand, $\underset{x}{\arg\min}f'(x)$
represents the projection distance at which this largest gradient is
attained. Based on the above discussion, it is expected that these two
values should be both high in the case of high squintability (fast
increase in $f$ early on), and both low in the case of low squintability
(any substantial increase in $f$ happens very late, close to the optimal
angle). This suggest that their product (@eq-squintability) should
provide a sensible measure of squintability. The multiplicative constant
$4$, which can be deemed arbitrary, does not change the interpretation
of the squintability metric $\varsigma$; it is here to adjust the range
of values of $\varsigma$ and simplify the explicit formula for
$\varsigma$ obtained later on.

To compute the squintability metric (@eq-squintability) in practice,
several approaches are possible. The first one is to propose a
parametric model for $f$, and use it to obtain an explicit formula for
$\varsigma$. Numerical experiments suggest a scaled sigmoid shape as
described below. Define

$$
\ell(x):=\frac{1}{1+\exp(\theta_{3}(x-\theta_{2}))}\ ,
$$ {#eq-logistic}

which is a decreasing logistic function depending on two parameters
$\theta_2$ and $\theta_3$, such that $\ell(\theta_{2})=\frac{1}{2}$.
Then, define

$$
f(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(x_{\max})}{\ell(0)-\ell(x_{\max})}+\theta_{4}\ ,
$$ {#eq-parametric}

which depends on three additional parameters, $\theta_1$, $\theta_2$,
and $x_{\max}$, such that $f(0)=\theta_1$ and $f(x_{\max})=\theta_4$.
Under the parametric model (@eq-parametric), the squintability metric
(@eq-squintability) can be shown to be equal to

$$
\varsigma=\frac{(\theta_{1}-\theta_{4})\theta_{2}\theta_{3}}{\ell(0)-\ell(x_{\max})}\ .
$$ {#eq-squintability-parametric}

In practice, the parameters of this model (@eq-parametric) can be
estimated numerically, for example by non-linear least squares, and then
used to evaluate $\varsigma$ as in equation
(@eq-squintability-parametric).

Alternatively, one can estimate (@eq-squintability) in a nonparametric
way, for example by fitting $f$ using kernel regression, then
numerically estimate the angle at which $-f'$ attains its highest value.

<!-- \textcolor{orange}{See whether further details are needed for the nonparametric approach.} -->

<!--To the best of our knowledge, -->

<!-- ## Speed -->

<!-- The speed of optimizing an index function can be measured empirically by -->

<!-- recording the duration of the optimisation. Alternatively, one can also -->

<!-- gauge the speed in terms of computational complexity of evaluating the -->

<!-- index function (for instance, in big O notation, with respect to the -->

<!-- sample size) of computing the index function. -->

# Simulation details {#sec-sim-deets}

The JSO performance is compared with the existing Creeping Random Search
(CRS) [@RJ-2021-105; @laa_using_2020] currently used in guided tour, and
explores how JSO behaves with different hyper-parameter (number of
jellyfishes and maximum tries) and data dimension combinations. A second
simulation studies the effect of index properties (smoothness and
squintability), along with JSO hyper-parameters, and data dimension, on
the success rate of the JSO performance.

## Performance of JSO relative to CRS {#sec-app-1}

The performance of JSO is investigated in comparison with the existing
optimiser, CRS, with respect to two hyper-parameters (number of
jellyfishes and maximum number of tries). The performance is measured by
success rate, defined as the proportion of simulations that achieves a
final index value within 0.05 of the best index value found among all 50
simulations. This comparison is based on projection pursuit to find the
pipe shape investigated by @laa_using_2020 using two simple and
well-behaved indexes, the `holes` index, and the `splines2d` index.
Fifty simulations are conducted with both JSO and CRS, in four different
data dimensions ($d = 6, 8, 10, 12$). JSO uses 100 jellyfishes with 100
maximum number of tries, while the CRS allows a maximum of 1000 samples
at each iteration before the algorithm terminates. The different numbers
accounts for the multiple paths of JSO to enable fairer comparison with
CRS. The results of the simulation are collected using the data
structure proposed in @RJ-2021-105 is used here for assessing JSO, where
the design parameters are stored along with index value, projection
basis, random seed, and computation time.
<!--A sample for the sine-wave structure in 6D data using a spline index function (`splines2d`) is printed below:

```{r}
dplyr::glimpse(sim_sine_6d_spline_head |> select(-id, -info)) 
```
-->

In this simulation, 50 jellyfishes (`n_jellies`) and a maximum of 50 tries
(`max_tries`) are used. The full data contains 50 simulations, indexed
by the variable `sim`, with the associated `seed`s. Recorded in columns
`basis`, `index_val`, `tries`, `loop` and `time`, are the projection
basis, index value, iteration ID, jellyfish ID, and computational time,
respectively.
<!-- The basis column records every visited basis before -->
<!-- comparing with the current (last) basis. If a visited basis has a -->
<!-- smaller index value than the current (last) one, JSO will keep the -->
<!-- current basis for the next iteration in the search but still records the -->
<!-- visited inferior basis.  -->

In the second experiment, 50 simulations are conducted at each
hyper-parameter combination to analyse its effect on the JSO success
rate: number of jellyfishes (20, 50, and 100), and maximum number of tries
(50 and 100). @fig-success-rate illustrates the calculation based on the
projection found by the above simulation. Here, the optimal index value
is
`r sim_sine_6d_spline_best |> pull(index_val) |> max() |> round(3) -> aa; aa`.
Out of the 50 simulations,
`r sim_sine_6d_spline_best |> filter(index_val >= aa - 0.05) |> nrow() -> bb; bb`
achieved an index value within \[`r aa - 0.05`, `r aa`\], resulting in a
success rate of
`r sim_sine_6d_spline_best |> mutate(a = index_val >= aa - 0.05) |> pull(a) -> cc; round(sum(cc)/length(cc), 3)`.

<!-- @fig-projection shows the best projection found in each simulation, -->

<!-- sorted by the index value. Most of the projections show a clear -->

<!-- sine-wave structure, while the last six show only a straight line -->

<!-- with clusters at the end. When local optima exist in the projection -->

<!-- pursuit problem, plotting the projections found in each simulation -->

<!-- or by each jellyfish can help identify these local optima. -->

```{r}
#| label: fig-success-rate
#| fig.cap: "Illustration of success rate calculation: The best projections identified by the JSO in 50 simulations are sorted by their index values. The highest index value found across all simulation is 0.999. Out of the 50 simulations, 44 achieved an index value within 0.05 of the best, resulting in a success rate of 0.88 (44/50)."
# proj_dt <- map_dfr(sim_sine_6d_spline_best |> pull(basis), 
#               ~as_tibble(as.matrix(sine1000) %*% .x), .id = "sim") |> 
#   mutate(sim = as.numeric(sim)) |> 
#   left_join(sim_sine_6d_spline_best |> select(sim, index_val), by = "sim")
# 
# idx_val_sorted <- proj_dt |> pull(index_val) |> unique() |> sort()
# proj_dt |> 
#   #filter(index_val %in% c(idx_val_sorted[1:12], idx_val_sorted[39:50])) |>
#   ggplot(aes(x = V1, y = V2)) + 
#   geom_point(size = 0.5) + 
#   geom_text(aes(label = round(index_val, 3)), x = 0, y = 3.7, 
#             size = 8) +
#   xlim(-4, 4) + ylim(-4, 4) +
#   facet_wrap(~fct_reorder(as.factor(sim), -index_val), ncol = 10) + 
#   theme_bw() +
#   theme(aspect.ratio = 1, axis.ticks = element_blank(),
#         axis.text = element_blank(), axis.title = element_blank(),
#         panel.grid = element_blank(), strip.background = element_blank(),
#         strip.text = element_blank())
knitr::include_graphics(here::here("figures", "success-rate.png"))
```

## Factors affecting JSO success rate: index properties and jellyfish hyper-parameters {#sec-app-2}

This simulation uses two different data shapes, pipe and a sine wave,
investigated in 6D and 8D spaces with six different PP indexes:
`dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`. The two
JSO hyper-parameters are varied. A total of `r nrow(sim_df)`
combinations result, comprising of
`r nrow(filter(sim_df, index == "holes"))` computed on the pipe data and
`r nrow(filter(sim_df, index != "holes"))` on the sine-wave data. For
each case, JSO is run 50 times to calculate the success rate.

Smoothness and squintability are computed for each run, following the
procedures outlined in @sec-smoothness and @sec-squintability and as
illustrated in @fig-smoothness and @fig-squintability.

To compute smoothness, 300 random bases are simulated. Index values and
projection distance (to the optimal basis) are calculated for each
random basis before fitting the Gaussian process model to obtain the
smoothness measure for the index.

To compute squintability, 50 random bases are sampled and interpolated
to the optimal basis with a step size of 0.005. Index values and
projection distances are calculated for these interpolated bases and the
index values are averaged with a bin width of 0.005. The squintability
measure is then calculated by fitting a four-parameter scaled logistic
function of index value against projection distance, estimated by
non-linear least squares.

A generalised linear model is fitted to the resulting data, using a
binomial family and a logit link function. The data is pre-processed
by 1) scaling the JSO hyper-parameters by a factor of 10 for
interpretation, 2) creating a new binary variable `long_time` to
indicate cases with an average run time over 30 seconds, and 3)
re-coding the success rate for the `stringy` index as 0, because none of
the 50 simulations correctly identified the sine-wave shape.

<!-- **Projection distance**: The projection distance between projection -->

<!-- bases and the optimal best basis (or the empirical best basis if -->

<!-- the optimal is unknown) measures how close a visited basis is to  -->

<!-- the best basis and it is calculated as the Frobenius norm in Equation -->

<!-- @eq-proj-dist. Plotting projection distances against index values for the  -->

<!-- bases visited reveals how the index value changes as the basis -->

<!-- approaches the optimal basis, helping to calculate the squintability -->

<!-- measure. -->

```{r}
#| label: fig-smoothness
#| out.width: 100%
#| fig.cap: "Illustration of steps to calculate smoothness. For a given projection pursuit problem defined by the shape to find, data dimension and the index function, sample random bases given the orthonormality contraint. Calculate the projection distance and the index value for each random basis. Fit a Gaussian process model to the index values against the projection distances to obtain the smoothness measure. "
# basis_smoothness <- sample_bases(idx = "splines2d", n_basis = 1000)
# basis_smoothness |>
#    ggplot(aes(x = dist, y = index)) +
#    geom_point() + 
#    labs(x = "Projection distance", y = "Index value") + 
#    theme_bw() +   
#   theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30))
knitr::include_graphics(here::here("figures", "smoothness.png"))
```

```{r}
#| label: fig-squintability
#| out.width: 100%
#| fig.cap: "Illustration of steps to calculate squintability. For a given projection pursuit problem defined by the shape to find, data dimension and the index function, sample random bases given the orthonormality and projection distance contraint. Interpolate the sampled bases to the optimal basis and calculate the projection distance and the index value for each interpolated basis. Bin the index value by the projection distance, then claculate the average index value for each bin. Fit the scaled sigmoid function from Equation @eq-squintability to the binned index values against the projection distance using non-linear least square. Use Equation @eq-parametric to calculate the squintability measure from the estimated $\\theta_1$ to $\\theta_4$. "
# p1 <- sim_sine_6d_spline_projdist |> 
#   ggplot(aes(x = proj_dist, y = index_val)) +
#   geom_point(size = 0.1, alpha = 0.5) + 
#   labs(x = "Projection distance", y = "Index value") + 
#   theme_bw() + 
#   theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30))
# 
# 
# dist_bin <- ceiling(sim_sine_6d_spline_projdist[["proj_dist"]] / 0.1) * 0.1
# dt <- sim_sine_6d_spline_projdist |>
#   dplyr::bind_cols(dist_bin = dist_bin) |>
#   dplyr::group_by(dist_bin) |>
#   dplyr::summarise(index_val = mean(index_val, na.rm = TRUE))
# 
# p2 <-  dt |> 
#   ggplot(aes(x = dist_bin, y = index_val)) + 
#   geom_line() +
#   geom_point() + 
#   labs(x = "Projection distance", y = "Index value") + 
#   theme_bw() + 
#   theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30))
knitr::include_graphics(here::here("figures", "squintability.png"))
```

# Results {#sec-sim-res}

The results from the two simulations described in @sec-sim-deets are
analysed based on the final projections across two optimisers (JSO and
CRS) and the success rate across JSO hyper-parameters. Smoothness and
squintability are calculated across a collection of pipe-finding and
sine-wave finding problems to construct the relationship between success
rate, JSO hyper-parameters, and index properties.

The final projections found by the two optimisers (JSO and CRS) are
presented in @fig-proj, broken down by 10th quantile, faceted by the
data dimension. The JSO uses 100 jellyfishes and a maximum number of
tries of 100. The CRS uses a maximum of 1000 tries in each step of
random sampling step before the algorithm terminates. In the
6-dimensional data scenario, JSO consistently identifies a clear pipe
shape. The CRS also finds the pipe shape but with a wide rim, suggesting
a further polish search may be required. With increasing dimensions, JSO
may not always identify the pipe shape due to random sampling, but it
still presents a pipe shape in over 50% of cases. When compared to the
CRS, JSO reaches higher index values and clearer pipe shapes across all
quantiles. This suggests JSO can be more effective when searching higher
dimension spaces, which may be preferred by users.

The success rate calculated for each hyper-parameter combination is
presented in @fig-proportion. As the number of jellyfishes and maximum
tries increase, the success rate also increases. For simpler problems (6
dimensions), small parameter values (20 jellyfishes and a maximum of 50
tries) can achieve a high success rate. However, for higher-dimensional
problems (8, 10, and 12 dimensions), larger parameter values (i.e. 100
jellyfishes and a maximum of 100 tries) are necessary. Increasing both
parameters enhances the performance of JSO, but it also extends the
computational time required for the optimisation, which can be
computationally intensive when evaluating the index function (such as
scagnostic indexes) multiple times across numerous iterations.

```{r}
#| fig.height = 7.5,
#| fig.width = 10,
#| label = fig-proj,
#| fig.align = "center",
#| fig.cap = "Projections found by the JSO and CRS at each 10th quantile across 50 simulations. The projection pursuit problem is to find the pipe shape using the holes index in the 6, 8, 10, and 12-dimensional spaces. The JSO uses 100 jellyfishes and a maximum number of tries of 100. The CRS uses a maximum of 1000 tries in each step of random sampling step before the algorithm terminates. In the 6-D data space, JSO always finds a clear pipe shape while the CRS also finds the pipe shape but with a wide rim. At higher data dimensions, JSO finds a higher index value and a clearer pipe shape across all the quantiles than the CRS"

aaa <- map_dfr(
  list(pipe1000, pipe1000_8d, pipe1000_10d, pipe1000_12d), 
  ~{
    d_dim = dim(.x)[2]
    pipe_jellyfish |> 
      filter(d == d_dim) |> 
      arrange(index_val) |>
      filter(row_number() %in% c(1, seq(5, 50, 5))) |> 
      mutate(id = factor(
        paste0(c(0, seq(5, 50, 5)) / 50 * 100, "th"),
        levels = paste0(c(0, seq(5, 50, 5)) / 50 * 100, "th"))) |> 
      compute_projection(data = .x, col = c("d", "index_val", "id"))
  }
)

p1 <- aaa |> 
  ggplot() + 
  geom_point(aes(x = V1, y = V2), size = 0.05) + 
  geom_text(data = aaa |> select(-V1, -V2) |> unique(), 
            aes(label = round(index_val,3)), x =0, y = 3.7, 
            size = 3) + 
  facet_grid(d ~ id) + 
  xlim(-4, 4) + ylim(-4, 4) +
  theme_bw() +
  theme(aspect.ratio = 1, axis.ticks = element_blank(),
        axis.text = element_blank(), axis.title = element_blank(),
        panel.grid = element_blank())

bbb <- pipe_better |> 
  group_by(dim) |> 
  arrange(index_val) |> 
  filter(row_number() %in% c(1, seq(5, 50, 5))) |>  
  mutate(id = factor(
    paste0(c(0, seq(5, 50, 5)) / 50 * 100, "th"),
    levels = paste0(c(0, seq(5, 50, 5)) / 50 * 100, "th")), 
    d= 4 + 2 * as.numeric(dim)) |> 
  unnest(proj)

p2 <- bbb |> 
  ggplot() + 
  geom_point(aes(x = V1, y = V2), size = 0.05) + 
  geom_text(data = bbb |> select(-V1, -V2) |> unique(), 
            aes(label = round(index_val,3)), x =0, y = 3.7, 
            size = 3) + 
  facet_grid(d ~ id) + 
  xlim(-4, 4) + ylim(-4, 4) +
  theme_bw() +
  theme(aspect.ratio = 1, axis.ticks = element_blank(),
        axis.text = element_blank(), axis.title = element_blank(),
        panel.grid = element_blank())

(p1 + ggtitle("a. JSO"))/(p2 + ggtitle("b. CRS"))
```

```{r}
#| label = fig-proportion,
#| fig.width = 6,
#| fig.height = 2.5,
#| fig.align = "center",
#| fig.cap = "Proportion of simulations reaches near-optimal index values in the pipe-finding problem using the holes index. The proportion is calculated based on the number of simulations, out of 50, that achieve an index value within 0.05 of the best-performing simulation. As the dimensionality increases, the proportion of simulations reaching the optimal index value increases."
pipe_sim_best <- sim_pipe_run_best |> 
  group_by(n_jellies, max_tries, d) |>
  summarise(I_best = max(I_max))

pipe_res <- sim_pipe_run_best |> 
  left_join(pipe_sim_best) |> 
  mutate(diff = abs(I_max - I_best) < 0.05) |> 
  group_by(n_jellies, max_tries, d) |> 
  summarise(proportion = sum(diff) / n()) |> 
  mutate(d = as.factor(d), n_jellies = as.factor(n_jellies))

pipe_res |> 
  rename(`max # of tries` = max_tries) |>
  ggplot(aes(x = n_jellies, y = proportion, 
             group = interaction(d, `max # of tries`),
             color = d)) + 
  geom_line() + 
  scale_color_brewer(palette = "Dark2", name = "dimension") +
  facet_wrap(vars(`max # of tries`), 
             labeller = "label_both") + 
  theme_bw() + 
  xlab("Number of jellyfish") + 
  ylab("Proportion of success")

```

```{r}
sim_df2 <- sim_df |>   
  mutate(n_jellies = n_jellies/10,
         max_tries = max_tries/10,
         long_time = ifelse(time > 20, 1, 0),
         P_J = ifelse(index == "stringy", 0, P_J))

mod1 <- glm(P_J ~ smoothness + squintability + d + long_time + n_jellies + max_tries, data = sim_df2, family = binomial)
```

<!-- The low -->

<!-- squint value of MIC and TIC index is due to its convex shape of the -->

<!-- index value against the projection distance, as shown in -->

<!-- @fig-idx-proj-dist. Comparing to other concave shapes, the maximum first -->

<!-- derivative happens at a smaller projection distance ($\theta_2$), -->

<!-- requiring the optimiser to get closer to the optimal to see a -->

<!-- significant change in the index value, hence a small squintability -->

<!-- measure.  -->

The index properties, including smoothness and squintability, offer
numerical metrics to characterise the complexity of projection pursuit
optimisation problems. @tbl-smoothness-squintability presents the
parameters estimated from the Gaussian process (variance, range, smooth,
and nugget) and the scaled logistic function ($\theta_1$ to $\theta_4$)
for calculating both metrics for each case considered in the simulation.
The column "smooth" is used as the smoothness metric and the column
"squint" is calculated as equation @eq-squintability-parametric as the
squintability metric. @tbl-mod-output presents the results of fitting a
generalised linear model with a binomial family and a logit link
function to the data in @tbl-mod-data, where all the three components
(success rate, JS hyper-parameters, and index properties) are combined.
The model suggests that JSO success rate is positively associated with
the two hyper-parameters, as well as with the index properties:
smoothness and squintability. Specifically, using 10 more jellyfishes
and 10 more tries increases the odd ratio of success by
`r round(100 * (exp(broom::tidy(mod1)$estimate[6])-1), 2)`% and
`r round(100 * (exp(broom::tidy(mod1)$estimate[7])-1), 2)`%,
respectively. However, being flagged with long runtime and an increase
of data dimension reduce the success rate by
`r round(exp(broom::tidy(mod1)$estimate[5]) * 100, 2)`% and
`r round(exp(broom::tidy(mod1)$estimate[4]) * 100, 2)`%, respectively.
The variable `squintability` and `dimension` are significant, suggesting
their importance relative to JSO hyper-parameters in the optimisation
success. In defining a projection pursuit problem, factors such as the
shape-to-find, the index function used, and the data dimension,
determine the properties like smoothness and squintability. These
metrics can then be compared across different problems to understand
their relative complexity. The results suggest that squintability has a
more significant influence than smoothness on the success rate of JSO.
Once the characteristics of the projection pursuit problem are fully
understood, increasing JSO hyper-parameters can enhance the search
effectiveness, however, it is also important to consider the resulting
increase in computational complexity.

```{r fig-idx-proj-dist, eval = FALSE}
#| fig.width = 9,
#| fig.height = 5,
#| fig.cap = "Index values versus projection distance for the 12 pipe/sine-wave finding problem, after the binning procedure for calculating the squintability measure. The index values, averaged at bin width of 0.005, are scaled from 0 to 1 for comparison (holes, TIC, and stringy). The `MIC` and `TIC` index curves are convex while others are concave. The stringy curve shows an instantaneous jump to the optimum when aproaching the best basis." 
sq_basis_dist_idx |>
  ggplot(aes(x = dist, y = index, group = index_name)) +
  geom_line() + 
  facet_nested_wrap(~index_name + n, nrow = 3, labeller = label_both) + 
  xlab("projection distance") + ylab("index value") +
  theme_bw()
```

```{r}
#| label: tbl-smoothness-squintability
#| tbl-cap: Parameters estimated from the Gaussian process (including variance, range, smoothness, and nugget) and scaled logistic function ($\theta_1$ to $\theta_4$) for the pipe-finding and sine-wave finding problems.  The column "smooth" and "squint" represent the smoothness and squintability measures. 
dt <- tibble(shape = c(rep("pipe", 4), rep("sine", 8)), smoothness) |>
  left_join(squintability) |> 
  rename(d = n, smooth = smoothness) |> 
  mutate(index = ifelse(index == "dcor2d_2", "dcor2d", index)) 
dt |> 
  knitr::kable(
    digits = 2, format = "latex", 
    col.names = c(
      "shape", "index", "d",
      "variance", "range", "smooth", "nugget",
      "$\\theta_1$", "$\\theta_2$", "$\\theta_3$", "$\\theta_4$", "squint"), linesep = "",
    booktabs = T, escape = FALSE, row.names = TRUE) |> 
  column_spec(1, border_left = TRUE) |> 
  column_spec(c(4, 8, 13), border_right = TRUE) |> 
  kable_styling(font_size = 10) 
```

```{r}
#| label: tbl-mod-data
#| tbl-cap: The first 7 rows of the datasets processed for modelling. 
sim_df |> 
  dplyr::select(index, d, smoothness, 
    squintability, n_jellies, max_tries, 
    -I_max, P_J, time) |> 
  mutate(time = as.numeric(time)) |>
  arrange(index, d, n_jellies, max_tries) |> 
  head(7) |> 
  kable(digits = 2, format = "latex", 
        col.names = c("index", "d", 
          "smoothness", "squintability", 
          "n. jellyfish", 
          "max. tries", 
          "success rate", "time (sec)"),
        linesep = "",
        booktabs = T) |> 
  kable_styling(#latex_options = "hold_position",
    font_size = 10)
```

```{r}
#| label: tbl-mod-output
#| tbl-cap: Model estimates of proportion of jellyfish success on index properties and jellyfish hyper-parameters from the generalised linear model with a binomial family and a logit link function. The variable smoothness, squintability, number of jellyfish and maximum number of tries are positively associated with JSO success rate while data dimension and being flagged as long runtime are negatively associated with the success rate. The variable squintability and dimension are significant, suggesting their importance relative to jellyfish hyper-parameters in the optimisation success.
broom::tidy(mod1) |> 
  mutate(term = c("Intercept", "Smoothness",
   "Squintability", "Dimension (d)", 
   "Long time", "N. jellyfish", 
   "Max. tries")) |>
  kable(digits = 2, format = "latex", 
    linesep = "", booktabs = T) |> 
  column_spec(c(1,5), border_right = TRUE) |> 
  column_spec(1, border_left = TRUE) |> 
  kable_styling(font_size = 10) 
```

# Conclusion {#sec-conclusion}

This paper has presented new metrics to mathematically define desirable
features of PP indexes, squintability and smoothness, and used these to
assess the performance of the new jellyfish search optimiser. The
metrics will be generally useful for characterising PP indexes, and help
with developing new indexes.

In the comparison of the JSO against the currently used CRS, as expected
the JSO vastly out-performs CRS, and provides a high probability of
finding the global optima. The JSO obtains the maximum more cleanly,
with a slightly higher index value, and plot of the projected data
showing the structure more clearly.

The JSO performance is affected by the hyper-parameters, with a higher
chance of reaching the global optima when more jellyfishes are used and the
maximum number of tries is increased. However, it comes at a
computational cost, as expected. The performance declines if the
projection dimension increases and if the PP index has low
squintability. The higher the squintability the better chance the JSO
can find the optima. However, interestingly smoothness does not affect
the JSO performance.

The new JSO is integrated with the current implementation of the
projection pursuit guided tour in the `tourr` package, and can be
further examined using PP optimisation diagnostics in the `ferrn`
package. Using the JSO is a little different than the current CRS,
because it runs many paths. The recommended approach is to conduct the
optimisation off-line, and then use the guided tour to follow selected
jellyfish. The tools for this are available in the `tourr` package, too.

```{=html}
<!--
- Summarise the result of the simulation: JSO immense improvement on current methods, ...
- JSO for a guided tour, what changes are needed? Do we follow one tentacle, do we perform off-line and then pull tentacles out to watch, how does it integrate with ferrn

Main findings:

  - JSO tends to obtian a clearer projection and higher index value when searching higher dimension spaces than CRS (and other current optimisers) - Figure 4

  - Increasing JS hyper-parameter enhances its performance, but the computational time required for the optimisation can quickly become intensive, especially when evaluating the index function (such as scagnostic indexes) multiple times across numerous iterations. - Figure 5

  - Smoothness and squintability can be calculated and compared across different problems to understand their relative complexity. Once a projection pursuit problem is fully characterised (by the shape-to-find, the index function used, and the data dimension), increasing JSO hyper-parameters can enhance the search effectiveness, however, the resulting increase in computational complexity will also need to be considered. - modelling
-->
```
# Acknowledgement

The article is created using Quarto [@Allaire_Quarto_2022] in R [@R].
The source code for reproducing the work reported in this paper can be
found at: <https://github.com/huizezhang-sherry/paper-optim>. Nicolas
Langrené acknowledges the partial support of the Guangdong Provincial
Key Laboratory IRADS (2022B1212010006, R0400001-22) and the UIC Start-up
Research Fund UICR0700041-22.

# References