### 2025-02-12 jso.tex
edits in sec 6

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
]{interact}

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{anyfontsize}
\usepackage{orcidlink}
\usepackage{algorithm}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\theoremstyle{plain}
\newtheorem{defn}{\protect\definitionname}
\newtheorem{prop}{\protect\propositionname}
\providecommand{\definitionname}{Definition}
\providecommand{\propositionname}{Proposition}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Squintability and Other Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimisation Choices},
  pdfauthor={H. Sherry Zhang; Dianne Cook; Nicolas Langrené; Jessica Wai Yin Leung},
  pdfkeywords={projection pursuit, jellyfish search optimiser
(JSO), optimisation, grand tour, high-dimensional data, exploratory data
analysis},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Squintability and Other Metrics for Assessing Projection Pursuit
Indexes, and Guiding Optimisation Choices}
\author{H. Sherry Zhang$\textsuperscript{1}$, Dianne
Cook$\textsuperscript{2}$, Nicolas
Langrené$\textsuperscript{3}$, Jessica Wai Yin
Leung$\textsuperscript{2}$}

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@austin.utexas.edu}{\nolinkurl{huize.zhang@austin.utexas.edu}}. Dianne
Cook. Email: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}. Nicolas
Langrené. Email: \href{mailto:nicolaslangrene@uic.edu.cn}{\nolinkurl{nicolaslangrene@uic.edu.cn}}. Jessica
Wai Yin
Leung. Email: \href{mailto:Jessica.Leung@monash.edu}{\nolinkurl{Jessica.Leung@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Statistics and Data
Sciences, University of Texas at Austin, Austin, United
States\\ \textsuperscript{2} Department of Econometrics and Business
Statistics, Monash
University, Melbourne, Australia\\ \textsuperscript{3} Department of
Mathematical Sciences, Guangdong Provincial/Zhuhai Key Laboratory of
Interdisciplinary Research and Application for Data Science,
BNU-HKBU~United~International~College, Zhuhai, China
\begin{abstract}
The projection pursuit (PP) guided tour interactively optimises a
criterion function known as the PP index, to explore high-dimensional
data by revealing interesting projections. Optimisation of some PP
indexes can be non-trivial, if they are non-smooth functions, or the
optimum has a small ``squint angle'', detectable only from close
proximity. To address these challenges, this study investigates the
performance of a recently introduced swarm-based algorithm, Jellyfish
Search Optimiser (JSO), for optimising PP indexes. The performance of
JSO for visualising data is evaluated across various hyper-parameter
settings and compared with existing optimisers. Additionally, methods
for calculating the smoothness and squintability properties of the PP
index are proposed. They are used to assess the optimiser performance in
the presence of PP index complexities. A simulation study illustrates
the use of these performance metrics to compare the JSO with existing
optimisation methods available for the guided tour. The JSO algorithm
has been implemented in the R package, \texttt{tourr}, and functions to
calculate smoothness and squintability are available in the
\texttt{ferrn} package.
\end{abstract}
\begin{keywords}
\def\sep{;\ }
projection pursuit\sep jellyfish search optimiser
(JSO)\sep optimisation\sep grand tour\sep high-dimensional data\sep 
exploratory data analysis
\end{keywords}

\setstretch{2}
\section{Introduction}\label{introduction}

Projection pursuit (PP) (Kruskal 1969; Friedman and Tukey 1974; Huber
1985) is a dimension reduction technique aimed at identifying
informative linear projections of data. This is useful for exploring
high-dimensional data, and creating plots of the data that reveal the
main features to use for publication. The method involves optimising an
objective function known as the PP index (e.g., Hall 1989; Cook, Buja,
and Cabrera 1993; Lee and Cook 2010; Loperfido 2018, 2020), which
defines the criterion for what constitutes interesting or informative
projections. Let \(X \in \mathbb{R}^{n\times p}\) be the data matrix,
\(A \in\mathbb{R}^{p \times d}\) be an orthonormal matrix, where \(A\)
belongs to the Stiefel manifold \(\mathcal{A} = V_d(\mathbb{R}^p)\). The
projection \(Y = XA\) is a linear transformation that maps data from a
\(p\)-dimensional space into a \(d\)-dimensional space. The index
function \(f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}\) is a scalar
function that measures an interesting aspect of the projected data, such
as deviation from normality, presence of clusters, or non-linear
structure. For a fixed sample of data, PP finds the orthonormal basis
\(A\) that maximises the index value of the projection, \(Y = XA\):

\begin{equation}\phantomsection\label{eq-optimization}{
\underset{A \in \mathcal{A}}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I_d
}\end{equation}

It is interesting to note that when using PP visually, one cares less
about \(A\) than the plane described by \(A\), because the orientation
in the plane is irrelevant. The space of planes belongs to a Grassmann
manifold. This is usually how the projection pursuit guided tour (PPGT)
(Cook et al. 1995) operates, when using geodesic interpolation between
starting and target planes. It interpolates plane to plane, removing
irrelevant within plane spin, and is agnostic to the basis (\(A\)) used
to define the plane. Thus, indexes which are used for the PPGT should be
rotationally invariant.

Index functions are quite varied in form, partially depending on the
data that is being projected. Figure~\ref{fig-example-functions} shows
two examples. Huber plots (Huber 1990) of 2D data sets are in (a) and
(c), showing the PP index values for all 1D projections of the 2D data
in polar coordinates, which reveals the form of these functions. The
dashed circle is a baseline set at the average value, and the straight
line marks the optimal projection. Plots (b) and (d) show the respective
best projections of the data as histograms. Indexes like the holes,
central mass and skewness (Cook, Buja, and Cabrera 1993) are generally
smooth for most data sets, but capture only large patterns. Many indexes
are noisy and non-convex, requiring an effective and efficient
optimisation procedure to explore the data landscape and achieve a
globally optimal viewpoint of the data. The skewness index computed for
trimodal data, in (a), is smooth with a large squint angle but has three
modes, and thus is not convex. The binned normality index (a simple
version of a non-normality index as described in Huber 1985) computed on
the famous RANDU data, in (c), is noisier and has a very small squint
angle. The discreteness cannot be seen unless the optimiser is very
close to the optimal projection.

\begin{figure}

\centering{

\includegraphics[width=0.8\textwidth,height=\textheight]{jso_files/figure-pdf/fig-example-functions-1.pdf}

}

\caption{\label{fig-example-functions}Examples of PP indexes with large
(top row) and small (bottom row) squint angles, shown with a Huber plot,
and histogram of the projected data corresponding to the optimal
projection. A Huber plot shows the PP index values for all 1D data
projections in polar coordinates.}

\end{figure}%

Optimisation of PP is often discussed when new indexes are proposed
(Posse 1995; Marie-Sainte, Berro, and Ruiz-Gazen 2010; Grochowski and
Duch 2011). Cook et al. (1995) tied the optimisation more closely to the
index, when they introduced the PPGT, which monitors the optimisation
visually so that the user can see the projected data leading in and out
of the optimum. An implementation is available in the \texttt{tourr}
package (Wickham et al. 2011) in R (R Core Team 2023). Zhang et al.
(2021) illustrated how to diagnose optimisation processes, particularly
focusing on the guided tour, and revealed a need for improved
optimisation. While improving the quality of the optimisation solutions
in the tour is essential, it is also important to be able to view the
data projections as the optimisation progresses. Integrating the guided
tour with a global optimisation algorithm that is efficient in finding
the global optimal and enables viewing of the projected data during the
exploration process is a goal.

Here, the potential for a Jellyfish Search Optimiser (JSO) Rajwar, Deep,
and Das (2023) for the PPGT is explored. JSO, inspired by the search
behaviour of jellyfish in the ocean, is a swarm-based metaheuristic
designed to solve global optimisation problems. Compared to traditional
methods, JSO has demonstrated stronger search ability and faster
convergence, and requires fewer tuning parameters. These practicalities
make JSO a promising candidate for enhancing PP optimisation.

The primary goal of the study reported here is to investigate the
performance of JSO in PP optimisation for the guided tour. It is of
interest to assess how quickly and closely the optimiser reaches a
global optimum, for various PP indexes that may have differing
complexities. To observe the performance of JSO with different types of
PP indexes, metrics are introduced to capture specific properties of the
index including squintability (based on Tukey and Tukey 1981's squint
angle) and smoothness. Here, we mathematically define metrics for
squintability and smoothness, which is a new contribution for PP
research. A series of simulation experiments using various datasets and
PP indexes are conducted to assess JSO's behaviour and its sensitivity
to hyper-parameter choices (number of jellyfish and maximum number of
tries). The relationship between the JSO performance, hyper-parameter
choices and properties of PP indexes (smoothness and squintability) is
analysed to provide guidance on selecting optimisers for practitioners
using projection pursuit. Additionally, this work should guide the
design of new PP indexes and facilitate better optimization for PP.

The paper is structured as follows. Section~\ref{sec-background}
introduces the background of the PPGT, reviews existing optimisers and
index functions in the literature. Section~\ref{sec-PP-properties}
introduces the metrics that measure two properties of PP indexes,
smoothness and squintability, and Section~\ref{sec-JSO} describes the
new JSO to be used for the PPGT. Section~\ref{sec-sim-deets} outlines
two simulation experiments to assess JSO's performance: one comparing
JSO's performance improvements relative to an existing optimiser,
Creeping Random Search (CRS), and the other studying the impact of PP
index properties on optimisation performance, and
Section~\ref{sec-sim-res} presents the results.
Section~\ref{sec-discussion} discusses the implementation of JSO in the
\texttt{tourr} package and the PP property calculation in the
\texttt{ferrn} package. Section~\ref{sec-conclusion} summarises the work
and provides suggestions for future directions.

\section{Projection pursuit, tours, index functions and
optimisation}\label{sec-background}

A tour on high-dimensional data is constructed by geodesically
interpolating between pairs of planes. Any plane is described by an
orthonormal basis, \(A_t\), where \(t\) represents time in the sequence.
The term ``geodesic'' refers to maintaining the orthonormality
constraint so that each view shown is correctly a projection of the
data. The PP guided tour operates by geodesically interpolating to
target planes (projections) which have high PP index values, as provided
by the optimiser. The geodesic interpolation means that the viewer sees
a continuous sequence of projections of the data, so they can watch
patterns of interest forming as the function is optimised. There are
five optimisation methods implemented in the \texttt{tourr} package:

\begin{itemize}
\tightlist
\item
  a pseudo-derivative, that searches locally for the best direction,
  based on differencing the index values for very close projections.
\item
  a brute-force optimisation (CRS).
\item
  a modified brute force algorithm described in Posse (1995).
\item
  an essentially simulated annealing (Bertsimas and Tsitsiklis 1993)
  where the search space is reduced during the optimisation.
\item
  a very localised search, to take tiny steps to get closer to the local
  maximum.
\end{itemize}

There are numerous PP index functions available: introduced in Huber
(1985), Cook, Buja, and Cabrera (1993), Lee et al. (2005), Lee and Cook
(2010), Grimm (2016), Ursula Laa and Valencia (2022). Most are
relatively simply defined, for any projection dimension, and implemented
because they are relatively easy to optimise. A goal is to develop PP
indexes based on scagnostics Leland Wilkinson and Wills (2008), but the
blockage is their optimisation as these tend to be noisy, with
potentially small squint angles.

An initial investigation of PP indexes, and the potential for
scagnostics is described in Laa and Cook (2020). To be useful here an
optimiser needs to be able to handle index functions that are possibly
not very smooth. In addition, because data structures might be
relatively fine, the optimiser needs to be able to find maxima that
occur with a small squint angle, that can only be seen from very close
by. One last aspect that is useful is for an optimiser to return local
maxima in addition to the global one because data can contain many
different and interesting features.

\section{Properties of PP indexes}\label{sec-PP-properties}

Laa and Cook (2020) has proposed five criteria for assessing projection
pursuit indexes (smoothness, squintability, flexibility, rotation
invariance, and speed). Since not all index properties affect the
optimisation process, the focus here is on the first two properties,
\emph{smoothness} (Section~\ref{sec-smoothness}) and
\emph{squintability} (Section~\ref{sec-squintability}), for which
metrics are proposed to quantify them.

\subsection{Smoothness}\label{sec-smoothness}

This subsection proposes a metric for the smoothness of a projection
pursuit index.

A classical way to describe the smoothness of a function is to identify
how many continuous derivatives of the function exist. This can be
characterized by Sobolev spaces (Adams and Fournier 2003).

\begin{defn}[Sobolev space]\label{def:sobolev_space}
The Sobolev space $W^{k,p}(\mathbb{R})$ for $1\leq p\leq \infty$ is the set of all functions $f$ in $L^p(\mathbb{R})$ for which all weak derivatives $f^{(\ell)}$ of order $\ell\leq k$ exist and have a finite $L^p$ norm.
\end{defn}

The Sobolev index \(k\) in Definition \ref{def:sobolev_space} can be
used to characterize the smoothness of a function: if \(f\in W^{k,p}\),
then the higher \(k\), the smoother \(f\). While this Sobolev index
\(k\) is a useful measure of smoothness, it can be difficult to compute
or even estimate in practice.

To obtain a computable estimator of the smoothness of the index function
\(f\), we propose an approach based on random fields. If a PP index
function \(f\) is evaluated at some random bases, as is done at the
initialization stage of JSO, then these random index values can be
interpreted as a random field, indexed by a space parameter, namely the
random projection basis. This analogy suggests to use this random
training sample to fit a spatial model. We propose to use a Gaussian
process equipped with a Matérn covariance function, due to the
connections between this model and Sobolev spaces, see for example Porcu
et al. (2024).

The distribution of a Gaussian process is fully determined by its mean
and covariance function. The smoothness property comes into play in the
definition of the covariance function: if a PP index is very smooth,
then two close projection bases should produce close index values
(strong correlation); by contrast, if a PP index is not very smooth,
then two close projection bases might give very different index values
(fast decay of correlations with respect to distance between bases).
Popular covariance functions are parametric positive semi-definite
functions. In particular, the Matérn class of covariance functions has a
dedicated parameter to capture the smoothness of the Gaussian field.

\begin{defn}[Matérn covariance function]\label{def:matern}
The Matérn covariance function $K$ is defined by
\begin{equation}
K(u)=K_{\nu,\eta,\ell}(u):=\eta^2\frac{\left(\sqrt{2\nu}\frac{\left\Vert u\right\Vert}{\ell}\right)^{\nu}}{\Gamma(\nu)2^{\nu-1}}\mathcal{K}_{\nu}\left(\sqrt{2\nu}\frac{\left\Vert u\right\Vert}{\ell}\right)\ ,\label{eq:matern}
\end{equation}
where $\left\Vert u\right\Vert$ is the Euclidean norm of $u\in\mathbb{R}^{p{\times}d}$, $\nu>0$ is the smoothness parameter, $\eta$ is the outputscale, $\ell$ is the lengthscale, and $\mathcal{K}_\nu$ is
the modified Bessel function [DLMF 10.25].
\end{defn}

The Matérn covariance function can be expressed analytically when
\(\nu\) is a half-integer, the most popular values in the literature
being \(\frac{1}{2}\), \(\frac{3}{2}\) and \(\frac{5}{2}\) (Rasmussen
and Williams 2006). The parameter \(\nu\), called \emph{smoothness
parameter}, controls the decay of the covariance function. As such, it
is an appropriate measure of smoothness of a random field, as shown by
the simulations on Figure~\ref{fig-matern-1d} and
Figure~\ref{fig-matern-2d}. For example, Karvonen (2023) showed that if
a function \(f\) has a Sobolev index of \(k\), then the smoothness
parameter estimate \(\nu\) in \eqref{eq:matern} cannot be asymptotically
less than \(k\). See the survey Porcu et al. (2024) for additional
results on the connection between the Matérn model and Sobolev spaces.
An interesting result is that the asymptotic case
\(\nu\rightarrow\infty\) coincides with the Gaussian kernel:
\(K_\infty(u)=\exp(-{\left\Vert u\right\Vert}^{2}/2)\).

\begin{figure}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{figures/matern_simulation_1d.png}

}

\caption{\label{fig-matern-1d}Five random simulations from a Gaussian
Process defined on \(\mathbb{R}\) with zero mean and Matérn-\(\nu\)
covariance function, with \(\nu=1\) (left), \(\nu=2\) (middle), and
\(\nu=4\) (right), showing that higher values of \(\nu\) produce
smoother curves.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{figures/matern_simulation_2d.png}

}

\caption{\label{fig-matern-2d}One random simulation from a Gaussian
Process defined on \(\mathbb{R}^2\) with zero mean and Matérn-\(\nu\)
covariance function, with \(\nu=1\) (left), \(\nu=2\) (middle), and
\(\nu=4\) (right), showing that higher values of \(\nu\) produce
smoother surfaces.}

\end{figure}%

In view of these results, the parameter \(\nu\) is suggested as a
measure of the smoothness of the PP index function by fitting a Gaussian
process prior with Matérn covariance on a dataset generated by random
evaluations of the index function, as done at the initialization stage
of the jellyfish search optimization. There exist several R packages,
such as \texttt{GpGp} (Guinness, Katzfuss, and Fahmy 2021) or
\texttt{ExaGeoStatR} (Abdulah et al. 2023), to fit the hyper-parameters
of a GP covariance function on data, which is usually done by maximum
likelihood estimation. In this project, the \texttt{GpGp} package is
used.

\begin{defn}
Let $\mathbf{A}=[A_1, \ldots, A_N] \in (\mathbb{R}^{p \times d})^N$ be d-dimensional projection bases, let $\mathbf{y}=[f(XA_1),\ldots,f(XA_N)]$ be the corresponding PP index values, and let $\mathbf{K}=[K_\theta(A_{i}-A_{j})]_{1\leq i,j\leq N}\in\mathbb{R}^{N\times N}$ be the Matérn covariance matrix evaluated at the input bases, where the vector $\theta$ contains all the parameters of the multivariate Matérn covariance function $K$ (smoothness, outputscale, lengthscales). The log-likelihood of the parameters $\theta$ is defined by 
\begin{equation}
\mathcal{L}(\theta)=\log p(\mathbf{y}\left|\mathbf{A},\theta\right.)=-\frac{1}{2}\mathbf{y}^{\top}(\mathbf{K}+\sigma^{2}\mathbf{I})^{-1}\mathbf{y}-\frac{1}{2}\mathrm{\log}(\det(\mathbf{K}+\sigma^{2}\mathbf{I}))-\frac{N}{2}\log(2\pi)\, \label{eq:gp_log_likelihood}
\end{equation}
where the nugget parameter $\sigma$ is the standard deviation of the intrinsic noise of the Gaussian process.
The optimal parameters (including smoothness) are obtained by maximum log-likelihood
\begin{equation}
\theta^* = \underset{\theta}{\max}\mathcal{L}(\theta)
\end{equation}
The resulting optimal smoothness parameter $\nu$ is chosen as our smoothness metric.
\end{defn}

The value of the optimal smoothness parameter \(\nu>0\) can be naturally
interpreted as follows: the higher \(\nu\), the smoother the index
function.

\subsection{Squintability}\label{sec-squintability}

This section defines the projection distance and introduces the
squintability metric, followed by a description of two numerical
approaches for computing squintability.

\begin{defn}[projection distance]\label{def:proj-dist}
Let $A \in \mathbb{R}^{p \times d}$ be a $d$-dimensional orthonormal matrix, and let $A^*$ be the optimal matrix that achieves the maximum index value for a given data. The projection distance between $A$ 
and $A^*$, $r(A, A^*)$, is defined by
$r(A, A^*) = \lVert AA^\prime - A^*A^{*\prime}\,\rVert _F$
where $\lVert . \rVert _F$ denotes the Frobenius norm, given by
$\lVert M \rVert _F = \sqrt{\sum_{ij} M_{ij}^2}$. 
\end{defn}
\begin{defn}[squint angle]\label{def:squint-angle}

Let $A$ and $B$ be two $d$-dimensional orthonormal matrices in $\mathbb{R}^p$. The squint angle $\theta$ between the subspace spanned by $A$ and $B$ is defined as the smallest principal angle between these subspaces: $\theta = \min_{i \in \{1, \cdots, d\}} \arccos(\tau_i)$, where $\tau_i$ are the singular values of the matrix $M = A^T B$ obtained from its singular value decomposition.
\end{defn}

Squintability can be defined based on how the index value \(f(XA)\)
changes with respect to the projection distance \(r(A, A^*)\), over the
course of the JSO. Suppose that the optimization starts with an
arbitrary candidate matrix \(A_0\) at a distance \(r_0:=r(A_0, A^*)\)
from the optimal one. A possible way to define squintability is as
follows:

\begin{defn}[squintability]\label{def:squintability}
Let $g: \mathbb{R} \mapsto  \mathbb{R}$ be a decreasing function that maps the projection distance $r(A, A^*)$ to the index value $f(XA)$, such that $g(r) = g(r(A, A^*)) = f(XA)$.  The squintability of an index function $f$ is defined by

\begin{equation}
\varsigma(f) = \frac{g(r_{0}/2)-g(r_{0})}{g(0)-g(r_{0})} \in [0,1]
\label{eq-squintability}
\end{equation}

\end{defn}

Remark that the amount by which the index value can improve from the
starting matrix \(A_0\) to the optimal one \(A^*\) is given by
\(g(0)-g(r_{0})\). As a result, equation~\eqref{eq-squintability}
represent the proportion of this maximum improvement which has been
achieved by the time the distance \(r_0\) to the optimal matrix has been
reduced by half (\(r_0/2\)).

It is expected that this value should be high in the case of high
squintability (fast increase in \(g\) early on), and low in the case of
low squintability (any substantial increase in \(g\) happens very late,
close to the optimal angle). This suggests that this half-point
improvement metric should provide a sensible measure of squintability.

From Tukey and Tukey (1981) and Laa and Cook (2020), a large squint
angle implies that the objective function value is close to optimal even
when the perfect view to see the structure is far away. A small squint
angle means that the PP index value improves substantially only when the
perfect view is close by. As such, low squintability implies rapid
improvement in the index value when near the perfect view. For PP, a
small squint angle is considered to be undesirable because it means that
the optimiser needs to be very close to be able to ``see'' the optimum.
Thus, it could be difficult for the optimiser to find the optimum.

It is expected that for a PP index with high squint angle, the
optimization (\ref{eq-optimization}) should make substantial progress
early on. Conversely, for a PP index with low squint angle, it might
take a long while for the optimization to make substantial progress, as
the candidate projections would need to be very close to the optimal one
for the structure of the index function to be visible enough to be
amenable to efficient optimization. This observation suggests that the
half-point index value improvement percentage provides an appropriate
mathematical definition of squintability, which matches the intuition
behind this concept, while being amenable to numerical computation.

To compute the squintability metric \eqref{eq-squintability} in
practice, several approaches are possible. The first one is to propose a
parametric model for \(g\), and use it to obtain an explicit formula for
\(\varsigma\). Numerical experiments suggest a scaled sigmoid shape as
described below. Define
\begin{equation}\phantomsection\label{eq-logistic}{
\ell(x):=\frac{1}{1+\exp(\theta_{3}(x-\theta_{2}))}\ ,
}\end{equation}

which is a decreasing logistic function depending on two parameters
\(\theta_2\) and \(\theta_3\), such that
\(\ell(\theta_{2})=\frac{1}{2}\). Then, define
\begin{equation}\phantomsection\label{eq-parametric}{
g(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(r_0)}{\ell(0)-\ell(r_0)}+\theta_{4}\ ,
}\end{equation}

which depends on three additional parameters, \(\theta_1\),
\(\theta_2\), and \(r_0\), such that \(g(0)=\theta_1\) and
\(g(r_0)=\theta_4\). Under the parametric model (\ref{eq-parametric}),
the squintability metric \eqref{eq-squintability} can be shown to be
equal to
\begin{equation}\phantomsection\label{eq-squintability-parametric}{
\varsigma=\frac{g\left(r_{0}/2\right)-\theta_{4}}{\theta_{1}-\theta_{4}}=\frac{\ell(r_0/2)-\ell(r_0)}{\ell(0)-\ell(r_0)}\ .
}\end{equation}

In practice, the parameters of this model (\ref{eq-parametric}) can be
estimated numerically, for example by non-linear least squares, and then
used to evaluate \(\varsigma\) as in equation
(\ref{eq-squintability-parametric}).

Alternatively, one can estimate \eqref{eq-squintability} in a
nonparametric way, for example by fitting \(g\) using kernel regression,
then numerically estimate \(\varsigma\) from its definition
\eqref{eq-squintability}.

\section{The jellyfish optimiser}\label{sec-JSO}

The Jellyfish Search Optimiser (JSO) mimics the natural movements of
jellyfish, which include passive and active motions driven by ocean
currents and their swimming patterns, respectively. In the context of
optimization, these movements are abstracted to explore the search
space, aiming to balance exploration (searching new areas) and
exploitation (focusing on promising areas). The algorithm aims to find
the optimal solution by adapting the behaviour of jellyfish to navigate
towards the best solution over iterations (Chou and Truong 2021).

To solve the optimisation problem embedded in the PP guided tour, a
starting projection, an index function, the number of jellyfish, and the
maximum number of trials (tries) are provided as input. Then, the
current projection is evaluated by the index function. The projection is
then moved in a direction determined by a random factor, influenced by
how far along we are in the optimisation process. Occasionally,
completely new directions may be taken like a jellyfish might with ocean
currents. A new projection is accepted if it is an improvement compared
to the current one, rejected otherwise. This process continues and
iteratively improves the projection, until the pre-specified maximum
number of trials is reached.

\begin{tcolorbox}[enhanced jigsaw, bottomrule=.15mm, rightrule=.15mm, left=2mm, colbacktitle=quarto-callout-note-color!10!white, colframe=quarto-callout-note-color-frame, leftrule=.75mm, toptitle=1mm, colback=white, titlerule=0mm, breakable, opacityback=0, opacitybacktitle=0.6, bottomtitle=1mm, toprule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, coltitle=black]

\textbf{Input}: \texttt{current\_projections}, \texttt{index\_function},
\texttt{trial\_id}, \texttt{max\_trial}

\textbf{Output}: \texttt{optimised\_projection}

\textbf{Initialize} \texttt{current\_best} as the projection with the
best index value from \texttt{current\_projections}, and
\texttt{current\_idx} as the array of index values for each projection
in \texttt{current\_projections}

\textbf{for} each \texttt{trial\_id} in 1 to \texttt{max\_tries}
\textbf{do}

\begin{quote}
Calculate the time control value, \(c_t\), based on
\texttt{current\_idx} and \texttt{max\_trial}
\end{quote}

\begin{quote}
\textbf{if} \(c_t\) is greater than or equal to \(0.5\) \textbf{then}

\begin{quote}
Define trend based on the \texttt{current\_best} and
\texttt{current\_projections}
\end{quote}

\begin{quote}
Update each projection towards the trend using a random factor and
orthonormalisation
\end{quote}

\textbf{else}

\begin{quote}
\textbf{if} a random number is greater than \(1 - c_t\) \textbf{then}

\begin{quote}
Slightly adjust each projection with a small random factor (passive)
\end{quote}

\textbf{else}

\begin{quote}
For each projection, compare with a random jellyfish and adjust towards
or away from it (active)
\end{quote}
\end{quote}

Update the orientation of each projection to maintain consistency

Evaluate the new projections using the index function
\end{quote}

\begin{quote}
\textbf{if} any new projection is worse than the current, revert to the
\texttt{current\_projections} for that case

\begin{quote}
Determine the projection with the best index value as the new
\texttt{current\_best}
\end{quote}
\end{quote}

\begin{quote}
\textbf{if} \texttt{trial\_id} \(\ge\) \texttt{max\_trial}, print the
last best projection \textbf{exit}
\end{quote}

\textbf{return} the set of projections with the updated
\texttt{current\_best} as the \texttt{optimised\_projection}

\end{tcolorbox}

The JSO implementation involves several key parameters that control its
search process in optimization problems. These parameters are designed
to guide the exploration and exploitation phases of the algorithm. While
the specific implementation details can vary depending on the version of
the algorithm or its application, the focus is on two main parameters
that are most relevant to our application: the number of jellyfish and
the maximum number of tries.

\section{Assessing optimisers: two simulation
studies}\label{sec-sim-deets}

This section describes the details of two simulation studies. The first
simulation compares the JSO performance with an existing optimiser,
Creeping Random Search (CRS) (Zhang et al. 2021; Laa and Cook 2020) to
explore JSO's behaviour under different data dimensions. The second
simulation examines various factors affecting the performance of JSO.
The JSO performance is evaluated under different hyper-parameters
combinations (number of jellyfish and the maximum number of tries). The
effect of index properties (smoothness and squintability) along with JSO
hyper-parameters and data dimension variations is captured and studied
using logistic regression.

The performance of JSO is measured by the success rate, defined as the
proportion of simulations that achieves a final index value within 0.05
of the best index value found among all 50 simulations (see
Figure~\ref{fig-success-rate} for an illustration). This comparison is
based on projection pursuit to find the pipe shape investigated by Laa
and Cook (2020) using the \texttt{holes} index.

\subsection{Performance of JSO relative to CRS}\label{sec-app-1}

The JSO performance is compared to the existing optimizer, CRS, in a
pipe-finding problem using the holes index in different dimensional
spaces. Fifty simulations are conducted in four data dimensions
(\(d = 6, 8, 10, 12\)) with each optimiser. JSO uses 100 jellyfish with
a maximum of 100 tries, while the CRS allows a maximum of 1000 samples
at each iteration before the algorithm terminates. The different numbers
account for the multiple paths of JSO to enable fairer comparison with
CRS. The results of the simulation are collected using the data
structure proposed in Zhang et al. (2021) for assessing JSO, where the
design parameters are stored along with index value, projection basis,
random seed, and computation time.

\begin{figure}

\centering{

\includegraphics[width=4.52in,height=\textheight]{figures/success-rate.png}

}

\caption{\label{fig-success-rate}Illustration of success rate
calculation: Final projections based on projection pursuit to find the
pipe shape in 8D data using the holes index, optimised by CRS, in 50
simulations. The 50 final projections are sorted by their index values.
The highest index value found across all simulations is 0.969. Out of
the 50 simulations, 43 achieved an index value within 0.05 of the best,
resulting in a success rate of 0.86 (43/50).}

\end{figure}%

\subsection{Factors affecting JSO success rate: JSO hyper-parameters and
index properties}\label{sec-app-2}

To examine the performance of JSO across various hyper-parameter
combinations, the pipe-finding problem using the holes index is repeated
for fifty times in each hyper-parameter combination: 20, 50, and 100
jellyfish and a maximum of 50 and 100 attempts. The success rate is
calculated for each combination.

In addition, to assess the performance of JSO across various PP
problems, two different data shapes, pipe and a sine wave, are
considered in various dimensional spaces using six different PP indexes:
\texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
\texttt{spline}, and \texttt{stringy}, under varied JSO
hyper-parameters. This results in a total of 76 scenarios, comprising of
30 computed on the pipe data and 46 on the sine-wave data. A list of
combinations of the data shape, dimensions and PP indexes considered are
provided in Table~\ref{tbl-smoothness-squintability}. Again, JSO is run
50 times for each scenario to calculate the success rate for each
hyper-parameter combination.

\begin{figure}

\centering{

\includegraphics{jso_files/figure-pdf/fig-plot-data-1.pdf}

}

\caption{\label{fig-plot-data}The sine and pipe structures in the
example data, along with index values for these projections.}

\end{figure}%

Smoothness and squintability are computed following the procedures
outlined in Section~\ref{sec-smoothness} and
Section~\ref{sec-squintability} and as illustrated in
Figure~\ref{fig-smoothness} and Figure~\ref{fig-squintability}. To
compute smoothness, 500 random bases are simulated. Index values are
calculated for each random basis, followed by fitting a Gaussian process
model to obtain the smoothness measure for the index.

To compute squintability, 50 random bases are sampled and interpolated
to the optimal basis with a step size of 0.005. Index values and
projection distances are calculated for these interpolated bases. A
binning procedure is applied to average the index values within each
0.005 projection distance bin. The four-parameter scaled logistic
function (\ref{eq-parametric}) is fitted to the index values against
projection distances, estimated by non-linear least squares to obtain
the squintability measure as Equation
(\ref{eq-squintability-parametric}).

To construct a relationship among success rate, smoothness,
squintability, and JSO hyper-parameters, a generalised linear model is
fitted using a binomial family and a logit link function. The data is
pre-processed by 1) scaling the JSO hyper-parameters by a factor of 10
for interpretation, 2) creating a new binary variable
\texttt{long\_time} to indicate cases with an average run time over 20
seconds, and 3) re-coding the success rate for the \texttt{stringy}
index as 0, because none of the 50 simulations correctly identified the
sine-wave shape.

\begin{figure}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{figures/smoothness.png}

}

\caption{\label{fig-smoothness}Steps for calculating smoothness in a
projection pursuit problem. Given the target shape, data dimension and
the index function: 1) sample random bases given the orthonormality
contraint and calculate their corresponding index values, and 2) fit a
Gaussian process model of index values against the sampled bases to
obtain the smoothness measure.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{figures/squintability.png}

}

\caption{\label{fig-squintability}Steps for calculating squintability in
a projection pursuit problem. Given the target shape, data dimension and
the index function: 1) sample random bases given the orthonormality and
projection distance contraint and calculates their corresponding index
values, 2) interpolate the sampled bases to the optimal basis and
calculate the projection distance and the index value. 3) bin the index
values by projection distances through averaging index value, 4) fit the
scaled sigmoid function in equation \eqref{eq-squintability} to the
binned index values against projection distances using non-linear least
square to obtian the squintability measure using equation
(\ref{eq-squintability-parametric}).}

\end{figure}%

\section{Results}\label{sec-sim-res}

\subsection{Performance of JSO relative to
CRS}\label{performance-of-jso-relative-to-crs}

Performance of JSO is assessed based on the final projections across the
two optimisers: JSO and CRS. The final projections found by the two
optimisers are presented in Figure~\ref{fig-proj}, broken down by 10th
quantile, faceted by the data dimensions. In the 6-dimensional data
scenario, JSO consistently identifies a clear pipe shape. The CRS also
finds the pipe shape but with a wide rim, suggesting a further polish
search may be required. With increasing dimensions, JSO may not always
identify the pipe shape due to random sampling, but it still finds the
pipe shape in over 50\% of cases. Compared to CRS, JSO achieves higher
index values and clearer pipe shapes across all quantiles in data of 8,
10, 12 dimensions, suggesting its advantage in exploring
high-dimensional spaces.

\subsection{Factors affecting JSO success rate: JSO hyper-parameters and
index
properties}\label{factors-affecting-jso-success-rate-jso-hyper-parameters-and-index-properties}

The effect of JSO hyper--parameters (number of jellyfish and the maximum
number of tries) on the success rate is presented in
Figure~\ref{fig-proportion}. The uncertainty is quantified through 500
bootstrap replicates for each case. As the number of jellyfish and
maximum tries increase, the success rate also increases. For problems
with relatively lower dimensional search spaces (4 and 6 dimensions),
small parameter values (20 jellyfish and a maximum of 50 tries) are
sufficient for space exploration and thus can already achieve a high
success rate. Higher parameter values (i.e.~100 jellyfish and a maximum
of 100 tries) enhances exploration of the search space, which is
particularly beneficial in higher-dimensional problems (8, 10, and 12
dimensions) where a more complex search is necessary. However, in
lower-dimensional problems, the benefit may be less pronounced, as the
space is alredy relatively easy to navigate. While increasing both
parameters enhances the performance of JSO, it also extends the
computational time required for the optimisation, which can be
computationally intensive when evaluating the index function (such as
scagnostic indexes) multiple times across numerous iterations.

Smoothness and squintability are calculated across a collection of
pipe-finding and sine-wave finding problems to construct the
relationship between success rate, JSO hyper-parameters, and index
properties. Table~\ref{tbl-smoothness-squintability} presents the
parameters estimated from the Gaussian process (outputscale \(\eta\),
lengthscale \(\ell\), smoothness \(\nu\), and nugget \(\sigma\)) and
from the scaled logistic function (\(\theta_1\) to \(\theta_4\)) for
calculating smoothness and squintability in each PP problem considered.
The column \(\nu\) is used as the smoothness metric and the column
\(\varsigma\) is calculated as equation
(\ref{eq-squintability-parametric}) as the squintability metric.

\begin{figure}

\centering{

\includegraphics{jso_files/figure-pdf/fig-proj-1.pdf}

}

\caption{\label{fig-proj}Projections found by the JSO and CRS at each
10th quantile across 50 simulations. The projection pursuit problem is
to find the pipe shape using the holes index in the 6, 8, 10, and
12-dimensional spaces. The JSO uses 100 jellyfish and a maximum number
of tries of 100. The CRS uses a maximum of 1000 tries in each step of
random sampling step before the algorithm terminates. In the 6-D data
space, JSO always finds a clear pipe shape while the CRS also finds the
pipe shape but with a wide rim. At higher data dimensions, JSO finds a
higher index value and a clearer pipe shape across all the quantiles
than the CRS}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{jso_files/figure-pdf/fig-proportion-1.pdf}

}

\caption{\label{fig-proportion}The proportion of simulations that reach
near-optimal index values in the pipe-finding problem using the holes
index. The proportion is calculated based on the number of simulations,
out of 50, that achieve an index value within 0.05 of the
best-performing simulation. To quantify uncertainty, 500 bootstrap
samples are generated. The thin segments represent the proportion for
each of the 500 bootstrap replicates, while the thicker segments
represent the mean of these bootstrap samples, connnected by lines. As
the dimensionality increases, the proportion of simulations reaching the
optimal index value decreases.}

\end{figure}%

\begin{table}

\caption{\label{tbl-smoothness-squintability}Parameters estimated from
the Gaussian process (outputscale \(\eta\), lengthscale \(\ell\),
smoothness \(\nu\), and nugget \(\sigma\)) and scaled logistic function
(\(\theta_1\) to \(\theta_4\) and \(\varsigma\)) for the pipe-finding
and sine-wave finding problems. The columns \(\nu\) and \(\varsigma\)
represent the smoothness and squintability measures respectively.}

\centering{

\centering\begingroup\fontsize{10}{12}\selectfont

\begin{tabular}{|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr}
\toprule
shape & index & d & $\eta$ & $\ell$ & $\nu$ & $\sigma$ & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\theta_4$ & $\varsigma$\\
\midrule
pipe & holes & 4 & 0.03 & 0.36 & 2.99 & 0.00 & 1.00 & 0.78 & 4.50 & 0.03 & 0.31\\
pipe & holes & 6 & 0.03 & 0.38 & 3.09 & 0.00 & 1.02 & 0.86 & 3.37 & 0.02 & 0.41\\
pipe & holes & 8 & 0.02 & 0.33 & 2.90 & 0.00 & 1.01 & 0.87 & 3.26 & 0.03 & 0.41\\
pipe & holes & 10 & 0.02 & 0.34 & 2.88 & 0.00 & 1.02 & 0.88 & 3.15 & 0.02 & 0.43\\
sine & holes & 12 & 0.02 & 0.43 & 2.72 & 0.00 & 1.01 & 0.88 & 3.34 & 0.00 & 0.41\\
sine & MIC & 4 & 0.14 & 0.27 & 3.56 & 0.00 & 0.94 & 0.15 & 1.14 & -0.04 & 0.39\\
sine & MIC & 6 & 0.06 & 0.31 & 3.08 & 0.00 & 0.89 & 0.57 & 1.62 & -0.03 & 0.39\\
sine & MIC & 8 & 0.03 & 0.32 & 2.75 & 0.00 & 0.93 & 0.33 & 1.31 & -0.04 & 0.38\\
sine & TIC & 4 & 0.40 & 0.27 & 3.49 & 0.00 & 0.97 & 0.46 & 1.55 & -0.04 & 0.38\\
sine & TIC & 6 & 0.16 & 0.31 & 3.08 & 0.00 & 0.95 & 0.54 & 1.72 & -0.03 & 0.37\\
sine & TIC & 8 & 0.07 & 0.30 & 2.72 & 0.00 & 0.95 & 0.56 & 1.72 & -0.03 & 0.38\\
sine & dcor & 4 & 0.22 & 0.30 & 3.63 & 0.00 & 0.96 & 1.07 & 2.67 & -0.03 & 0.56\\
sine & dcor & 6 & 0.08 & 0.29 & 3.13 & 0.00 & 0.95 & 1.04 & 2.74 & -0.03 & 0.53\\
sine & loess & 4 & 0.34 & 0.50 & 2.90 & 0.00 & 1.02 & 1.10 & 2.50 & 0.05 & 0.57\\
sine & loess & 6 & 0.24 & 0.46 & 3.56 & 0.00 & 1.02 & 1.04 & 2.65 & 0.07 & 0.53\\
sine & loess & 8 & 0.19 & 0.46 & 3.36 & 0.00 & 1.02 & 1.04 & 2.62 & 0.08 & 0.53\\
sine & skinny & 6 & 0.05 & 0.19 & 2.24 & 0.01 & 0.83 & 0.00 & 2.98 & 0.05 & 0.10\\
sine & splines & 4 & 0.24 & 0.32 & 3.55 & 0.00 & 1.01 & 1.09 & 2.65 & -0.03 & 0.57\\
sine & splines & 6 & 0.10 & 0.32 & 3.11 & 0.00 & 1.01 & 1.05 & 2.73 & -0.02 & 0.54\\
sine & splines & 8 & 0.05 & 0.29 & 2.74 & 0.00 & 1.02 & 1.06 & 2.68 & -0.02 & 0.54\\
sine & stringy2 & 4 & 0.02 & 0.30 & 2.76 & 1.01 & 0.64 & 0.05 & 7.89 & 0.05 & 0.00\\
sine & stringy2 & 6 & 0.02 & 0.07 & 1.51 & 0.05 & 0.65 & 0.05 & 7.78 & 0.06 & 0.00\\
\bottomrule
\end{tabular}
\endgroup{}

}

\end{table}%

Table~\ref{tbl-mod-output} presents the results from fitting a logistic
regression model using the proportion of success as the response and
smoothness, squintability, dimension, number of jellyfish and maximum
number of tries, as predictors. The fit suggests that the JSO success
rate is only affected by squintability, dimension and number of jellies.
As expected, the JSO success rate is higher with higher squintability
and/or more jellyfish, and lower when dimension is higher
(higher-dimensional optimization is more difficult). Interestingly, the
JSO is unaffected by the smoothness of the index function. This is
consistent with the way random search algorithms jump from value to
value without taking local regularity into account, as opposed to
gradient-based optimizers for example. Allowing for more tries also does
not affect success rate significantly. A unit increase in squintability
increases the success rate by 24\%. As dimension increases by one the
success rate almost halves. Increasing the number of jellies by 10,
increases the success rate by 32\%.

\begin{table}

\caption{\label{tbl-mod-output}Jellyfish success rate relative to index
properties and jellyfish hyper-parameters. This is the summary from a
logistic regression fit to smoothness, squintability, dimension, number
of jellyfish and maximum number of tries. Interestingly, squintability
and dimension strongly affect jellyfish optimisation success. The number
of jellies marginally affects success, but index smoothness, and
increasing the number of tries do not.}

\centering{

\fontsize{12.0pt}{14.4pt}\selectfont
\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lcc}
\toprule
\textbf{Characteristic} & \textbf{OR} \textbf{(95\% CI)}\textsuperscript{\textit{1}} & \textbf{p-value} \\ 
\midrule\addlinespace[2.5pt]
Smoothness & 1.00 (0.81 to 1.20) & 0.961 \\ 
Squintability & 1.24 (1.05 to 1.58) & {\bfseries 0.033} \\ 
Dimension & 0.57 (0.33 to 0.87) & {\bfseries 0.022} \\ 
Number of jellyfish & 1.32 (1.06 to 1.71) & {\bfseries 0.022} \\ 
Maximum number of tries & 1.06 (0.85 to 1.35) & 0.603 \\ 
\bottomrule
\end{tabular*}
\begin{minipage}{\linewidth}
\textsuperscript{\textit{1}}OR = Odds Ratio, CI = Confidence Interval\\
\end{minipage}

}

\end{table}%

\section{Practicalities}\label{sec-discussion}

The simulation studies have compared the JSO with CRS and shown how
smoothness and squintability affect the JSO success rate. Here we
explain how they can be used for new index and optimiser development.
Using the JSO optimiser for PPGT requires a change in user behaviour.
For all the existing optimisers a single optimisation path is followed.
The JSO has multiple optimisation paths, and needs additional tools to
incorporate into the PPGT, as explained here.

\subsection{Using the JSO in a PPGT}\label{using-the-jso-in-a-ppgt}

To use JSO for PPGT in the \texttt{tourr} package, specify
\texttt{search\_f\ =\ search\_jellyfish} in the guided tour. Unlike
existing optimisers, the animation function \texttt{animate\_*()} won't
directly render the tour path for JSO due to the generation of multiple
tour paths. To visualise the path visited by each individual jellyfish,
assign the animation to an object
(\texttt{res\ \textless{}-\ animate\_xy(...)}). This will save the bases
visited by JSO, along with relevant metadata, as a tibble data object
(see Zhang et al. 2021 for more details on the data object). The bases
visited for individual jellyfish can then be extracted and viewed using
the \texttt{planned\_tour()} function.

\subsection{Computing the index properties for your new
index}\label{computing-the-index-properties-for-your-new-index}

The \texttt{ferrn} package (Zhang et al. 2021) provides functionality
for computing the smoothness and squintability metrics. Both metrics
require sampling random bases using the \texttt{sample\_bases()}
function, followed by the metric calculation with
\texttt{calc\_smoothness()} or \texttt{calc\_squintability()}.

\subsubsection{Smoothness}\label{smoothness}

To sample bases for calculating smoothness, the following arguments are
required: the index function, the dataset, and the number of random
bases to sample. The output of \texttt{sample\_bases()} is a data frame
with a list-column of sampled basis matrix and index value.
Parallellasation is available to speed up the index value computation
through the \texttt{parallel\ =\ TRUE} argument.

The \texttt{calc\_smoothness()} function takes the output from
\texttt{sample\_bases()} and fits a Gaussian process model to the index
values against the sampled basis, as the location, to obtain the
smoothness metric. Starting parameters and additional Gaussian process
arguments can be specified and details of the fit can be accessed
through the \texttt{fit\_res} attribute of the output.

\subsubsection{Squintability}\label{squintability}

Bases sampling for squintability includes an additional step of
interpolating between the sampled bases and the best basis. This step is
performed when the arguments \texttt{step\_size} and
\texttt{min\_proj\_dist} in the \texttt{sample\_bases()} function are
set to non-NA numerical values. Given the projection distance typically
ranging from 0 to 2, it is recommended to set \texttt{step\_size} to 0.1
or lower, and \texttt{min\_proj\_dist} to be at least 0.5 to ensure a
meaningful interpolation length.

The \texttt{calc\_squintability()} function computes squintability using
two methods: 1) parametrically, by fitting a scaled sigmoid function
through non-linear least square (\texttt{method\ =\ "nls"}), and 2)
non-parametrically, using kernel smoothing (\texttt{method\ =\ "ks"}). A
\texttt{bin\_width} argument is required to average the index values
over the projection distance before the fit. For the parametric case,
the output provides the estimated parameters for the scaled sigmoid
function (\(\theta_1\) to \(\theta_4\)) and the calculated squintability
metric as Equation (\ref{eq-squintability-parametric}). For the
non-parametric case, it shows the maximum gradient attained
(\texttt{max\_d}), the corresponding projection distance
(\texttt{max\_dist}), and the squintability metric as their products.

\section{Conclusion}\label{sec-conclusion}

This paper has presented new metrics to mathematically define desirable
features of PP indexes, squintability and smoothness, and used these to
assess the performance of the new jellyfish search optimiser. The
metrics will be generally useful for characterising PP indexes, and help
with developing new indexes.

In the comparison of the JSO against the currently used CRS, as expected
the JSO vastly outperforms CRS, and provides a high probability of
finding the global optimum. The JSO obtains the maximum more cleanly,
with a slightly higher index value, and plot of the projected data
showing the structure more clearly.

The JSO performance is affected by the hyper-parameters, with a higher
chance of reaching the global optimum when more jellyfish are used and
the maximum number of tries is increased. However, it comes at a
computational cost, as expected. The performance declines if the
projection dimension increases and if the PP index has low
squintability. The higher the squintability the better chance the JSO
can find the optimum. However, interestingly smoothness does not affect
the JSO performance.

One future direction of this work is to test the JSO, along with its
variations and other swarm-based optimisers, on a broader range of
indexes, for example, scagnostic indexes.

\section{Acknowledgement}\label{acknowledgement}

The article has been created using Quarto (Allaire et al. 2022) in R (R
Core Team 2023). The source code for reproducing the work reported in
this paper can be found at:
\url{https://github.com/huizezhang-sherry/paper-jso}. The simulation
data produced in Section 5 can be found at
\url{https://figshare.com/articles/dataset/Simulated_raw_data/26039506}.
Nicolas Langrené acknowledges the partial support of the Guangdong
Provincial/Zhuhai Key Laboratory IRADS (2022B1212010006) and the UIC
Start-up Research Fund UICR0700041-22.

\section*{Supplementary materials}\label{supplementary-materials}
\addcontentsline{toc}{section}{Supplementary materials}

The supplementary materials available at
\url{https://github.com/huizezhang-sherry/paper-jso} include: 1) details
of the indexes used in the simulation study, 2) the script to get
started with using JSO in a PPGT and calculating smoothness and
squintability, as explained in Section~\ref{sec-discussion}, and (3) the
full code to reproduce the plots and summaries in this paper.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-abdulah2023large}
Abdulah, Sameh, Yuxiao Li, Jian Cao, Hatem Ltaief, David Keyes, Marc
Genton, and Ying Sun. 2023. {``Large-Scale Environmental Data Science
with {ExaGeoStatR}.''} \emph{Environmetrics} 34 (1): e2770.
\url{https://doi.org/10.1002/env.2770}.

\bibitem[\citeproctext]{ref-adams2003sobolev}
Adams, Robert, and John Fournier. 2003. \emph{Sobolev Spaces}. Vol. 140.
Pure and Applied Mathematics. Elsevier.

\bibitem[\citeproctext]{ref-Allaire_Quarto_2022}
Allaire, J. J., Charles Teague, Carlos Scheidegger, Yihui Xie, and
Christophe Dervieux. 2022. \emph{{Quarto}} (version 1.2).
\url{https://doi.org/10.5281/zenodo.5960048}.

\bibitem[\citeproctext]{ref-Bertsimas93}
Bertsimas, Dimitris, and John Tsitsiklis. 1993. {``{Simulated
Annealing}.''} \emph{Statistical Science} 8 (1): 10--15.
\url{https://doi.org/10.1214/ss/1177011077}.

\bibitem[\citeproctext]{ref-chou_novel_2021}
Chou, Jui-Sheng, and Dinh-Nhat Truong. 2021. {``A Novel Metaheuristic
Optimizer Inspired by Behavior of Jellyfish in Ocean.''} \emph{Applied
Mathematics and Computation} 389 (January): 125535.
\url{https://doi.org/10.1016/j.amc.2020.125535}.

\bibitem[\citeproctext]{ref-cook1993projection}
Cook, D., A. Buja, and J. Cabrera. 1993. {``Projection Pursuit Indexes
Based on Orthonormal Function Expansions.''} \emph{Journal of
Computational and Graphical Statistics} 2 (3): 225--50.
\url{https://doi.org/10.2307/1390644}.

\bibitem[\citeproctext]{ref-cook1995grand}
Cook, D., A. Buja, J. Cabrera, and C. Hurley. 1995. {``Grand Tour and
Projection Pursuit.''} \emph{Journal of Computational and Graphical
Statistics} 4 (3): 155--72.
\url{https://doi.org/10.1080/10618600.1995.10474674}.

\bibitem[\citeproctext]{ref-DLMF}
DLMF. 2024. {``{NIST Digital Library of Mathematical Functions}.''}
\url{https://dlmf.nist.gov/10.25}.

\bibitem[\citeproctext]{ref-FT74}
Friedman, J. H., and J. W. Tukey. 1974. {``A Projection Pursuit
Algorithm for Exploratory Data Analysis.''} \emph{IEEE Transactions on
Computers} C-23 (9): 881--90.
\url{https://doi.org/10.1109/T-C.1974.224051}.

\bibitem[\citeproctext]{ref-Grimm2016}
Grimm, Katrin. 2016. {``Kennzahlenbasierte Grafikauswahl.''} Doctoral
thesis, Universit{ä}t Augsburg.

\bibitem[\citeproctext]{ref-grochowski2011}
Grochowski, Marek, and Włodzisław Duch. 2011. {``Fast Projection Pursuit
Based on Quality of Projected Clusters.''} In \emph{Adaptive and Natural
Computing Algorithms}, edited by Andrej Dobnikar, Uroš Lotrič, and
Branko Šter, 89--97. Berlin, Heidelberg: Springer Berlin Heidelberg.
\url{https://doi.org/10.1007/978-3-642-20267-4_10}.

\bibitem[\citeproctext]{ref-guinness2021gpgp}
Guinness, Joseph, Matthias Katzfuss, and Youssef Fahmy. 2021. {``{GpGp}:
Fast {G}aussian Process Computation Using {V}ecchia's Approximation.''}
R package.

\bibitem[\citeproctext]{ref-hall1989polynomial}
Hall, Peter. 1989. {``On Polynomial-Based Projection Indices for
Exploratory Projection Pursuit.''} \emph{The Annals of Statistics} 17
(2): 589--605. \url{https://doi.org/10.1214/aos/1176347127}.

\bibitem[\citeproctext]{ref-huber85}
Huber, Peter J. 1985. {``Projection Pursuit.''} \emph{The Annals of
Statistics} 13 (2): 435--75.
\url{https://doi.org/10.1214/aos/1176349519}.

\bibitem[\citeproctext]{ref-huberplot}
---------. 1990. {``Data Analysis and Projection Pursuit.''} Technical
Report PJH-90-1. Dept. of Mathematics, Massachusetts Institute of
Technology.

\bibitem[\citeproctext]{ref-karvonen2023asymptotic}
Karvonen, Toni. 2023. {``Asymptotic Bounds for Smoothness Parameter
Estimates in {G}aussian Process Interpolation.''} \emph{SIAM/ASA Journal
on Uncertainty Quantification} 11 (4): 1225--57.
\url{https://doi.org/10.1137/22M149288X}.

\bibitem[\citeproctext]{ref-kr69}
Kruskal, J. B. 1969. {``Toward a Practical Method Which Helps Uncover
the Structure of a Set of Observations by Finding the Line
Transformation Which Optimizes a New {`Index of Condensation'}.''} In
\emph{Statistical Computation}, edited by R. C. Milton and J. A. Nelder,
427--40. New York: Academic Press.
\url{https://doi.org/10.1016/B978-0-12-498150-8.50024-0}.

\bibitem[\citeproctext]{ref-laa_using_2020}
Laa, Ursula, and Dianne Cook. 2020. {``Using Tours to Visually
Investigate Properties of New Projection Pursuit Indexes with
Application to Problems in Physics.''} \emph{Computational Statistics}
35 (3): 1171--1205. \url{https://doi.org/10.1007/s00180-020-00954-8}.

\bibitem[\citeproctext]{ref-lee2010projection}
Lee, Eun-Kyung, and Dianne Cook. 2010. {``A Projection Pursuit Index for
Large {\(p\)} Small {\(n\)} Data.''} \emph{Statistics and Computing} 20
(3): 381--92. \url{https://doi.org/10.1007/s11222-009-9131-1}.

\bibitem[\citeproctext]{ref-lee2005projection}
Lee, Eun-Kyung, Dianne Cook, Sigbert Klinke, and Thomas Lumley. 2005.
{``Projection Pursuit for Exploratory Supervised Classification.''}
\emph{Journal of Computational and Graphical Statistics} 14 (4):
831--46. \url{https://doi.org/10.1198/106186005X77702}.

\bibitem[\citeproctext]{ref-Loperfido2018}
Loperfido, Nicola. 2018. {``Skewness-Based Projection Pursuit: A
Computational Approach.''} \emph{Computational Statistics and Data
Analysis} 120 (C): 42--57.
\url{https://doi.org/10.1016/j.csda.2017.11.001}.

\bibitem[\citeproctext]{ref-Loperfido2020}
---------. 2020. {``Kurtosis-Based Projection Pursuit for Outlier
Detection in Financial Time Series.''} \emph{The European Journal of
Finance} 26 (2-3): 142--64.
\url{https://doi.org/10.1080/1351847X.2019.1647864}.

\bibitem[\citeproctext]{ref-marie-sainte2010}
Marie-Sainte, Souad Larabi, Alain Berro, and Anne Ruiz-Gazen. 2010.
{``An Efficient Optimization Method for Revealing Local Optima of
Projection Pursuit Indices.''} In \emph{Swarm Intelligence}, edited by
Marco Dorigo, Mauro Birattari, Gianni A. Di Caro, René Doursat, Andries
P Engelbrecht, Dario Floreano, Luca Maria Gambardella, et al., 60--71.
Berlin, Heidelberg: Springer Berlin Heidelberg.
\url{https://doi.org/10.1007/978-3-642-15461-4_6}.

\bibitem[\citeproctext]{ref-porcu2024matern}
Porcu, Emilio, Moreno Bevilacqua, Robert Schaback, and Chris Oates.
2024. {``The {M}atérn Model: A Journey Through Statistics, Numerical
Analysis and Machine Learning.''} \emph{Statistical Science} 39 (3):
469--92. \url{https://doi.org/10.1214/24-STS923}.

\bibitem[\citeproctext]{ref-posse95}
Posse, Christian. 1995. {``Projection Pursuit Exploratory Data
Analysis.''} \emph{Computational Statistics and Data Analysis} 20 (6):
669--87. \url{https://doi.org/10.1016/0167-9473(95)00002-8}.

\bibitem[\citeproctext]{ref-R}
R Core Team. 2023. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\bibitem[\citeproctext]{ref-rajwar_exhaustive_2023}
Rajwar, Kanchan, Kusum Deep, and Swagatam Das. 2023. {``An Exhaustive
Review of the Metaheuristic Algorithms for Search and Optimization:
Taxonomy, Applications, and Open Challenges.''} \emph{Artificial
Intelligence Review}, 1--71.
\url{https://doi.org/10.1007/s10462-023-10470-y}.

\bibitem[\citeproctext]{ref-rasmussen2006gaussian}
Rasmussen, Carl Edward, and Christopher K. I. Williams. 2006.
\emph{Gaussian Processes for Machine Learning}. The MIT Press.

\bibitem[\citeproctext]{ref-barnett1981interpreting}
Tukey, P. A., and J. W. Tukey. 1981. \emph{Graphical Display of Data in
Three and Higher Dimensions}. Wiley Series in Probability and
Mathematical Statistics: Applied Probability and Statistics. Wiley.
\url{https://books.google.com.au/books?id=WBzvAAAAMAAJ}.

\bibitem[\citeproctext]{ref-Laa:2020wkm}
Ursula Laa, Andreas Buja, Dianne Cook, and German Valencia. 2022.
{``Hole or Grain? A Section Pursuit Index for Finding Hidden Structure
in Multiple Dimensions.''} \emph{Journal of Computational and Graphical
Statistics} 31 (3): 739--52.
\url{https://doi.org/10.1080/10618600.2022.2035230}.

\bibitem[\citeproctext]{ref-tourr}
Wickham, H., D. Cook, H. Hofmann, and A. Buja. 2011. {``{tourr}: An {R}
Package for Exploring Multivariate Data with Projections.''}
\emph{Journal of Statistical Software} 40 (2): 1--18.
\url{https://doi.org/10.18637/jss.v040.i02}.

\bibitem[\citeproctext]{ref-scag}
Wilkinson, L., A. Anand, and R. Grossman. 2005. {``Graph-Theoretic
Scagnostics.''} In \emph{IEEE Symposium on Information Visualization,
2005. INFOVIS 2005.}, 157--64.
\url{https://doi.org/10.1109/INFVIS.2005.1532142}.

\bibitem[\citeproctext]{ref-WW08}
Wilkinson, Leland, and Graham Wills. 2008. {``Scagnostics
Distributions.''} \emph{Journal of Computational and Graphical
Statistics} 17 (2): 473--91.
\url{https://doi.org/10.1198/106186008X320465}.

\bibitem[\citeproctext]{ref-RJ-2021-105}
Zhang, H. Sherry, Dianne Cook, Ursula Laa, Nicolas Langrené, and
Patricia Menéndez. 2021. {``Visual Diagnostics for Constrained
Optimisation with Application to Guided Tours.''} \emph{The R Journal}
13: 624--41. \url{https://doi.org/10.32614/RJ-2021-105}.

\end{CSLReferences}

\end{document}