### 2025-03-03 jso.qmd
proofread before Sec 3.2

---
title: Squintability and Other Metrics for Assessing Projection Pursuit Indexes, and Guiding optimization Choices
author:
  - name: H. Sherry Zhang
    email: huize.zhang@austin.utexas.edu
    affiliations: 
        - id: 1
          name: University of Texas at Austin
          department: Department of Statistics and Data Sciences
          city: Austin
          country: United States
          postal-code: 78751
    attributes:
        corresponding: true
  - name: Dianne Cook
    email: dicook@monash.edu
    affiliations:
        - id: 2
          name: Monash University
          department: Department of Econometrics and Business Statistics
          city: Melbourne
          country: Australia
          postal-code: 3800
  - name: Nicolas  Langrené
    email: nicolaslangrene@uic.edu.cn
    affiliations:
        - id: 3
          name: Guangdong Provincial/Zhuhai Key Laboratory of Interdisciplinary Research and Application for Data Science, BNU-HKBU United International College
          department: Department of Mathematical Sciences
          city: Zhuhai
          country: China
          postal-code: 519087
  - name: Jessica Wai Yin Leung
    email: Jessica.Leung@monash.edu
    affiliations:
        - id: 2
          name: Monash University
          department: Department of Econometrics and Business Statistics
          city: Melbourne
          country: Australia
          postal-code: 3800
abstract: |
  The projection pursuit (PP) guided tour optimizes a criterion function, known as the PP index, to gradually reveal projections of interest from high-dimensional data through animation. Optimization of some PP indexes can be non-trivial, if they are non-smooth functions, or when the optimum has a small "squint angle", detectable only from close proximity. Here, measures for calculating the smoothness and squintability properties of the PP index are defined. These are used to investigate the performance of a recently introduced swarm-based algorithm, Jellyfish Search Optimizer (JSO), for optimizing PP indexes. The performance of JSO in detecting the target pattern (pipe shape) is compared with existing optimizers in PP. Additionally, JSO's performance on detecting the sine-wave shape is evaluated using different PP indexes (hence different smoothness and squintability) across various data dimensions (d = 4, 6, 8, 10, 12) and JSO hyper-parameters. The JSO algorithm has been implemented in the R package, `tourr`, and functions to calculate smoothness and squintability measures are implemented in the `ferrn` package.
keywords: 
  - Projection Pursuit Guided Tour (PPGT)
  - Jellyfish Search Optimizer (JSO)
date: last-modified
bibliography: bibliography.bib
format:
  tandf-pdf:
    keep-tex: true
    fontsize: 12pt
    linestretch: 2
editor: 
  markdown: 
    wrap: 72
crossref: 
  eq-prefix: ""
editor_options: 
  chunk_output_type: console
header-includes:
   - \usepackage{algorithm}
   - \usepackage{float} 
   - \usepackage{amsmath}
   - \usepackage{amsthm}
   - \usepackage{amssymb}
   - \theoremstyle{plain}
   - \newtheorem{defn}{\protect\definitionname}
   - \newtheorem{prop}{\protect\propositionname}
   - \providecommand{\definitionname}{Definition}
   - \providecommand{\propositionname}{Proposition}
nocite: |
   @DLMF
---

```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(patchwork)
library(ggh4x)
library(broom)
library(kableExtra)
library(tourr)
library(ferrn)
library(gtsummary)
load(here::here("data/sim_pipe_run_best.rda"))
load(here::here("data/sim_sine_6d_spline_head.rda"))
load(here::here("data/sim_sine_6d_spline_best.rda"))
load(here::here("data/sim_sine_6d_spline_projdist.rda"))
load(here::here("data/pipe_better.rda"))
load(here::here("data/pipe_jellyfish.rda"))
load(here::here("data/smoothness.rda"))
load(here::here("data/squintability.rda"))
load(here::here("data/sq_basis_df_6d_TIC_spline_skinny.rda"))
load(here::here("data/sim_df.rda"))
```

# Introduction

Projection pursuit (PP) [@kr69; @FT74; @huber85] is a dimension
reduction technique aimed at identifying informative linear projections
of data. This is useful for exploring high-dimensional data, and
creating plots of the data that reveal the main features to use for
publication. The method involves optimizing an objective function known
as the PP index [e.g., @hall1989polynomial; @cook1993projection;
@lee2010projection; @Loperfido2018; @Loperfido2020], which defines the
criterion for what constitutes interesting or informative projections.
Let $X \in \mathbb{R}^{n\times p}$ be the data matrix,
$A \in\mathbb{R}^{p \times d}$ be an orthonormal matrix in the Stiefel 
manifold $\mathcal{A} = V_d(\mathbb{R}^p)$. The
projection $Y = XA$ is a linear transformation of the $p$-dimensional data to
a $d$-dimensional space. The index function
$f(XA): \mathbb{R}^{n \times d} \to \mathbb{R}$ defines a statistics to measure
an pattern of interest in the projected data, such as
deviation from normality, presence of clusters, or non-linear structure.
For a fixed sample of data, PP finds the orthonormal matrix $A$ (also called 
*projection basis* in PP literature) that maximizes the index value of the 
projection:

$$
\underset{A \in \mathcal{A}}{\max } \quad f(XA) \quad \text{subject to} \quad A'A = I_d
$$ {#eq-optimization}

where $I_d$ is the identity matrix of dimension $d$.  
It is interesting to note that when using PP visually, one cares less
about $A$ than the plane described by $A$, because the orientation in
the plane is irrelevant. The space of planes belongs to a Grassmann
manifold. This is usually how the projection pursuit guided tour (PPGT)
[@cook1995grand] operates, when using geodesic interpolation between
starting and target planes. It interpolates plane to plane, removing
irrelevant within plane spin, and is agnostic to the basis ($A$) used to
define the plane. Thus, indexes which are used for the PPGT should be
rotationally invariant.

Index functions are quite varied in form, partially depending on the
data that is being projected. @fig-example-functions shows two examples.
Huber plots [@huberplot] of 2D data sets are in (a) and (c), showing the
PP index values for all 1D projections of the 2D data in polar
coordinates, which reveals the form of these functions. The dashed
circle is a baseline set at the average value, and the straight line
marks the optimal projection. Plots (b) and (d) show the respective best
projections of the data as histograms. Indexes like the holes, central
mass and skewness [@cook1993projection] are generally smooth for most
data sets, but capture only large patterns. Many indexes are noisy and
non-convex, requiring an effective and efficient optimization procedure
to explore the data landscape and achieve a globally optimal viewpoint
of the data. The skewness index computed for trimodal data, in (a), is
smooth with a large squint angle but has three modes, and thus is not
convex. The binned normality index [a simple version of a non-normality
index as described in @huber85] computed on the famous RANDU data, in
(c), is noisier and has a very small squint angle. The discreteness
cannot be seen unless the optimizer is very close to the optimal
projection.

```{r}
#| label: fig-example-functions
#| echo: false
#| fig-width: 9
#| fig-height: 6
#| out-width: 80%
#| fig-cap: "Examples of PP indexes with large (top row) and small (bottom row) squint angles, shown with a Huber plot, and histogram of the projected data corresponding to the optimal projection. A Huber plot shows the PP index values for all 1D data projections in polar coordinates. The skewness index on the trimodal data is also smoother than the binned normality index on RANDU."
#proj <- animate_xy(flea[,1:6], guided_tour(lda_pp(flea$species)))
best_proj <- matrix(c(-0.44157005,  0.80959831,
                      0.14622089, -0.08228193,
                      -0.08454314, -0.23234287,
                      0.70367852,  0.28185021,
                      0.27876825,  0.44819691,
                      0.45123453,  0.05896646), ncol=2, byrow=T)

flea2D <- as.matrix(flea[,1:6]) %*% best_proj
colnames(flea2D) <- c("X1", "X2")
flea2D <- as.data.frame(flea2D)
flea_huber <- prep_huber(flea2D, index = skewness())
sqa1 <- ggplot() +
  geom_huber(data = flea_huber$idx_df, aes(x = x, y = y)) +
  geom_point(data = flea2D, aes(x = X1, y = X2)) +
  geom_abline(slope = flea_huber$slope, intercept = 0) + 
  coord_fixed() +
  theme_huber() +
  ggtitle("(a) Skewness index")

sqa2 <- ggplot(flea_huber$proj_df, aes(x=x)) + 
  geom_histogram(binwidth=0.25, colour="white") +
  xlab("") + ylab("") +
  ggtitle("(b) Optimally projected data") +
  theme_bw() +
  theme(axis.text.y = element_blank())

data(randu)
randu_std <- as.data.frame(apply(randu, 2, function(x) (x-mean(x))/sd(x)))
randu_std$yz <- sqrt(35)/6*randu_std$y-randu_std$z/6
randu_df <- randu_std[c(1,4)]
randu_huber <- prep_huber(randu_df, index = norm_bin(nr = nrow(randu_df)))

sqa3 <- ggplot() +
  geom_huber(data = randu_huber$idx_df, aes(x = x, y = y)) +
  geom_point(data = randu_df, aes(x = x, y = yz)) +
  geom_abline(slope = randu_huber$slope, intercept = 0) +
  theme_huber() +
  coord_fixed() + 
  ggtitle("(c) Binned normality index")

sqa4 <- ggplot(randu_huber$proj_df, aes(x = x)) +
  geom_histogram(breaks = seq(-2.2, 2.4, 0.12)) +
  xlab("") + ylab("") +
  theme_bw() +
  theme(axis.text.y = element_blank()) + 
  ggtitle("(d) Optimally projected data")

sqa1 + sqa2 + sqa3 + sqa4 + plot_layout(ncol=2, widths=c(2,2))

```

Optimization of PP is often discussed when new indexes are proposed
[@posse95; @marie-sainte2010; @grochowski2011]. @cook1995grand tied the
optimization more closely to the index, when they introduced the PPGT,
which monitors the optimization visually so that the user can see the
projected data leading in and out of the optimum. An implementation is
available in the `tourr` package [@tourr] in R [@R]. @RJ-2021-105
illustrated how to diagnose optimization processes, particularly
focusing on the guided tour, and revealed a need for improved
optimization. While improving the quality of the optimization solutions
in the tour is essential, it is also important to be able to view the
data projections as the optimization progresses. Integrating the guided
tour with a global optimization algorithm that is efficient in finding
the global optimal and enables viewing of the projected data during the
exploration process is a goal.

In this work, the potential for a Jellyfish Search Optimizer (JSO) [
@chou_novel_2021; @rajwar_exhaustive_2023] for the PPGT is explored.
The JSO, inspired by the search behavior of jellyfish in the ocean, is a
swarm-based metaheuristic designed to solve global optimization
problems. Compared to other metaheuristic methods, JSO has demonstrated stronger
search ability and faster convergence, and requires fewer tuning
parameters. These practicalities make JSO a promising candidate for
enhancing PP optimization.

The primary goal of the study reported here is to investigate the
performance of JSO in PP optimization for the guided tour. It is of
interest to assess how quickly and closely the optimizer reaches a
global optimum, for various PP indexes that may have differing
complexities. To observe the performance of JSO with different types of
PP indexes, metrics are introduced to capture specific properties of the
index including squintability [based on @barnett1981interpreting's
squint angle] and smoothness. We mathematically define metrics for
squintability and smoothness, which is a new contribution for PP
research. A series of simulation experiments are conducted using JSO to detect 
different target patterns (pipe and sine-wave) with different PP indexes 
(holes, MIC, TIC, dcor, loess, splines, skinny, and stringy), as well as JSO's 
hyper-parameter choices (number of jellyfish and maximum number of
tries). This work should facilitate better optimization for PP and
guide the choice of optimizer when designing new PP indexes.

The paper is structured as follows. @sec-background introduces the
background of the PPGT, reviews existing optimizers and index functions
in the literature. @sec-PP-properties introduces the metrics that
measure two properties of PP indexes, smoothness and squintability. 
@sec-JSO describes the new JSO for PP. @sec-sim-deets
outlines two simulation experiments to assess JSO's performance: one
comparing JSO's performance improvements relative to an existing
optimizer, Creeping Random Search (CRS), and the other studying the
impact of PP index properties, controlling for data dimension and JSO 
hyper-parameters, on optimization performance, and
@sec-sim-res presents the results. @sec-discussion discusses the
implementation of JSO in the `tourr` package and the PP property
calculation in the `ferrn` package. @sec-conclusion summarizes the work
and provides suggestions for future directions.

# Projection pursuit, tours, index functions and optimization {#sec-background}

A tour on high-dimensional data is constructed by geodesically
interpolating between pairs of planes. Any plane is described by an
orthonormal basis, $A_t$, where $t$ represents time in the sequence. The
term "geodesic" refers to maintaining the orthonormality constraint so
that each view shown is correctly a projection of the data. The PP
guided tour operates by geodesically interpolating to target planes
(projections) which have high PP index values, as provided by the
optimizer. The geodesic interpolation means that the viewer sees a
continuous sequence of projections of the data, so they can watch
patterns of interest forming as the function is optimized. There are
five optimization methods implemented in the `tourr` package:

-   a pseudo-derivative, that searches locally for the best direction,
    based on differencing the index values for very close projections.
-   a brute-force optimization (CRS).
-   a modified brute force algorithm described in @posse95.
-   an essentially simulated annealing [@Bertsimas93] where the search
    space is reduced during the optimization.
-   a very localized search, to take tiny steps to get closer to the
    local maximum.

```{=html}
<!-- 
-   `search_geodesic()`: provides a pseudo-derivative optimization. It
    searches locally for the best direction, based on differencing the
    index values for very close projections. Then it follows the
    direction along the geodesic path between planes, stopping when the
    next index value fails to increase.
-   `search_better()`: also known as Creeping Random Search (CRS), is a
    brute-force optimization searching randomly for projections with
    higher index values.
-   `search_better_random()`: is essentially simulated annealing
    [@Bertsimas93] where the search space is reduced as the optimization
    progresses.
-   `search_posse()`: implements the algorithm described in @posse95.
-   `search_polish()`: is a very localized search, to take tiny steps to
    get closer to the local maximum.
-->
```
There are numerous PP index functions available: introduced in @huber85,
@cook1993projection, @lee2005projection, @lee2010projection, @Grimm2016,
@Laa:2020wkm. Most are relatively simply defined, for any projection
dimension, and implemented because they are relatively easy to optimize.
A goal is to develop PP indexes based on scagnostics [@scag; @WW08], but
the blockage is their optimization as these tend to be noisy, with
potentially small squint angles.

An initial investigation of PP indexes, and the potential for
scagnostics is described in @laa_using_2020. To be useful here an
optimizer needs to be able to handle index functions that are possibly
not very smooth. In addition, because the target structure in the data might be
relatively fine, the optimizer needs to be able to find maxima that
occur with a small squint angle, that can only be seen from very close
by. One last aspect that is useful is for an optimizer to return local
maxima in addition to the global one because data can contain many
different and interesting features.

# Properties of PP indexes {#sec-PP-properties}

@laa_using_2020 has proposed five criteria for assessing projection
pursuit indexes (smoothness, squintability, flexibility, rotation
invariance, and speed). Since not all index properties affect the
optimization process, the focus of this work is on the first two properties,
*smoothness* (@sec-smoothness) and *squintability* (@sec-squintability),
for which metrics are proposed to quantify them.

## Smoothness {#sec-smoothness}

A classical way to describe the smoothness of a function is to identify
how many continuous derivatives of the function exist. This can be
characterized by Sobolev spaces [@adams2003sobolev].

```{=tex}
\begin{defn}[Sobolev space]\label{def:sobolev_space}
The Sobolev space $W^{k,p}(\mathbb{R})$ for $1\leq p\leq \infty$ is the set of all functions $f$ in $L^p(\mathbb{R})$ for which all weak derivatives $f^{(\ell)}$ of order $\ell\leq k$ exist and have a finite $L^p$ norm.
\end{defn}
```

The Sobolev index $k$ in Definition \ref{def:sobolev_space} can be used
to characterize the smoothness of a function: if $f\in W^{k,p}$, then
the higher $k$, the smoother $f$. While this Sobolev index $k$ is a
useful measure of smoothness, it can be difficult to compute or even
estimate in practice. For example, the loess and splines indexes are not
continuous due to the use of the $\max$ operator in their definitions. 
Scagnostic indexes, such as skinny and stringy, are defined based on
graph elements (area, diameter, perimeter, and length) of alph hull or 
minimal spanning tree).

To obtain a computable estimator of the smoothness of the index function
$f$, we propose an approach based on random fields. If a PP index
function $f$ is evaluated at some random bases, then these random index values 
can be interpreted as a random field, indexed by a space parameter, namely the
random projection basis. This analogy suggests to use this random
training sample to fit a spatial model. We propose to use a Gaussian
process equipped with a Matérn covariance function, due to the
connections between this model and Sobolev spaces, see for example
@porcu2024matern.

The distribution of a Gaussian process is fully determined by its mean
and covariance function. The smoothness property comes into play in the
definition of the covariance function: if a PP index is very smooth,
then two close projection bases should produce close index values
(strong correlation); by contrast, if a PP index is not very smooth,
then two close projection bases might give very different index values
(fast decay of correlations with respect to distance between bases).
Popular covariance functions are parametric positive semi-definite
functions. In particular, the Matérn class of covariance functions has a
dedicated parameter to capture the smoothness of the Gaussian field.

```{=tex}
\begin{defn}[Matérn covariance function]\label{def:matern}
The Matérn covariance function $K$ is defined by
\begin{equation}
K(u)=K_{\nu,\eta,\ell}(u):=\eta^2\frac{\left(\sqrt{2\nu}\frac{\left\Vert u\right\Vert}{\ell}\right)^{\nu}}{\Gamma(\nu)2^{\nu-1}}\mathcal{K}_{\nu}\left(\sqrt{2\nu}\frac{\left\Vert u\right\Vert}{\ell}\right)\ ,\label{eq:matern}
\end{equation}
where $\left\Vert u\right\Vert$ is the Euclidean norm of $u\in\mathbb{R}^{p{\times}d}$, $\nu>0$ is the smoothness parameter, $\eta$ is the outputscale, $\ell$ is the lengthscale, and $\mathcal{K}_\nu$ is
the modified Bessel function [DLMF 10.25].
\end{defn}
```
The Matérn covariance function can be expressed analytically when $\nu$
is a half-integer, the most popular values in the literature being
$\frac{1}{2}$, $\frac{3}{2}$ and $\frac{5}{2}$ [@rasmussen2006gaussian].
The parameter $\nu$, called *smoothness parameter*, controls the decay
of the covariance function. As such, it is an appropriate measure of
smoothness of a random field, as shown by the simulations on
@fig-matern-1d and @fig-matern-2d. For example, @karvonen2023asymptotic
showed that if a function $f$ has a Sobolev index of $k$, then the
smoothness parameter estimate $\nu$ in \eqref{eq:matern} cannot be
asymptotically less than $k$. See the survey @porcu2024matern for
additional results on the connection between the Matérn model and
Sobolev spaces. An interesting result is that the asymptotic case
$\nu\rightarrow\infty$ coincides with the Gaussian kernel:
$K_\infty(u)=\exp(-{\left\Vert u\right\Vert}^{2}/2)$.

```{r}
#| label: fig-matern-1d
#| out.width: 100%
#| fig.cap: |
#|   Five random simulations from a Gaussian Process defined on $\mathbb{R}$ with zero mean and Matérn-$\nu$ covariance function, with $\nu=1$ (left), $\nu=2$ (middle), and $\nu=4$ (right), showing that higher values of $\nu$ produce smoother curves.
knitr::include_graphics(here::here("figures", "matern_simulation_1d.png"))
```

```{r}
#| label: fig-matern-2d
#| out.width: 100%
#| fig.cap: |
#|   One random simulation from a Gaussian Process defined on $\mathbb{R}^2$ with zero mean and Matérn-$\nu$ covariance function, with $\nu=1$ (left), $\nu=2$ (middle), and $\nu=4$ (right), showing that higher values of $\nu$ produce smoother surfaces.
knitr::include_graphics(here::here("figures", "matern_simulation_2d.png"))
```

In view of these results, the parameter $\nu$ is suggested as a measure
of the smoothness of the PP index function by fitting a Gaussian process
prior with Matérn covariance on a dataset generated by random
evaluations of the index function. There exist several R packages,
such as `GpGp` [@guinness2021gpgp] or `ExaGeoStatR` [@abdulah2023large],
to fit the hyper-parameters of a GP covariance function on data, which
is usually done by maximum likelihood estimation. In this project, the
`GpGp` package is used.

```{=tex}
\begin{defn}
Let $\mathbf{A}=[A_1, \ldots, A_N] \in (\mathbb{R}^{p \times d})^N$ be d-dimensional projection bases, $\mathbf{y}=[f(XA_1),\ldots,f(XA_N)]$ be the corresponding PP index values, and $\mathbf{K}=[K_\theta(A_{i}-A_{j})]_{1\leq i,j\leq N}\in\mathbb{R}^{N\times N}$ be the Matérn covariance matrix evaluated at the input bases, where the vector $\theta$ contains all the parameters of the multivariate Matérn covariance function $K$ (smoothness, outputscale, lengthscales). The log-likelihood of the parameters $\theta$ is defined by 
\begin{equation}
\mathcal{L}(\theta)=\log p(\mathbf{y}\left|\mathbf{A},\theta\right.)=-\frac{1}{2}\mathbf{y}^{\top}(\mathbf{K}+\sigma^{2}\mathbf{I})^{-1}\mathbf{y}-\frac{1}{2}\mathrm{\log}(\det(\mathbf{K}+\sigma^{2}\mathbf{I}))-\frac{N}{2}\log(2\pi)\, \label{eq:gp_log_likelihood}
\end{equation}
where the nugget parameter $\sigma$ is the standard deviation of the intrinsic noise of the Gaussian process.
The optimal parameters (including smoothness) are obtained by maximum log-likelihood
\begin{equation}
\theta^* = \underset{\theta}{\max}\mathcal{L}(\theta)
\end{equation}
The resulting optimal smoothness parameter $\nu$ is chosen as our smoothness metric.
\end{defn}
```
The value of the optimal smoothness parameter $\nu>0$ can be naturally
interpreted as follows: the higher $\nu$, the smoother the index
function.

## Squintability {#sec-squintability}

Squintability relates to the notion of squint angle and is defined based on how the index value $f(XA)$ changes with respect to the projection distance over the course of the optimization.

```{=tex}
\begin{defn}[squint angle]\label{def:squint-angle}
Let $A$ and $B$ be two $d$-dimensional orthonormal matrices in $\mathbb{R}^p$. The squint angle $\theta$ between the subspace spanned by $A$ and $B$ is defined as the smallest principal angle between these subspaces: $\theta = \min_{i \in \{1, \cdots, d\}} \arccos(\tau_i)$, where $\tau_i$ are the singular values of the matrix $M = A^T B$ obtained from its singular value decomposition.
\end{defn}

\begin{defn}[projection distance]\label{def:proj-dist}
Let $A \in \mathbb{R}^{p \times d}$ be a $d$-dimensional orthonormal matrix, and let $A^*$ be the optimal matrix that achieves the maximum index value for a given data. The projection distance between $A$ 
and $A^*$, $r(A, A^*)$, is defined by
$r(A, A^*) = \lVert AA^\prime - A^*A^{*\prime}\,\rVert _F$
where $\lVert . \rVert _F$ denotes the Frobenius norm, given by
$\lVert M \rVert _F = \sqrt{\sum_{ij} M_{ij}^2}$. 
\end{defn}
```

```{=tex}
\begin{defn}[squintability]\label{def:squintability}
Let $g: \mathbb{R} \mapsto  \mathbb{R}$ be a decreasing function that maps the projection distance $r(A, A^*)$ to the index value $f(XA)$, such that $g(r) = g(r(A, A^*)) = f(XA)$.  The squintability of an index function $f$ is defined by

\begin{equation}
\varsigma(f) = \frac{g(r_{0}/2)-g(r_{0})}{g(0)-g(r_{0})} \in [0,1]
\label{eq-squintability}
\end{equation}

\end{defn}
```

Remark that the amount by which the index value can improve from the starting matrix $A_0$ to the optimal one $A^*$ is given by $g(0)-g(r_{0})$. As a result, equation \eqref{eq-squintability} represent the proportion of this maximum improvement which has been achieved by the time the distance $r_0$ to the optimal matrix has been reduced by half ($r_0/2$).

It is expected that this value should be high in the case of
high squintability (fast increase in $g$ early on), and low in the
case of low squintability (any substantial increase in $g$ happens very
late, close to the optimal angle). This suggests that this half-point improvement metric should provide a sensible measure of
squintability.

From @barnett1981interpreting and @laa_using_2020, a large squint angle
implies that the objective function value is close to optimal even when
the perfect view to see the structure is far away. A small squint angle
means that the PP index value improves substantially only when the
perfect view is close by. As such, low squintability implies rapid
improvement in the index value when near the perfect view. For PP, a
small squint angle is considered to be undesirable because it means that
the optimizer needs to be very close to be able to "see" the optimum.
Thus, it could be difficult for the optimizer to find the optimum.
<!--The
mathematical formulation of this intuition is proposed below:-->

It is expected that for a PP index with high squint angle, the
optimization (@eq-optimization) should make substantial progress early
on. Conversely, for a PP index with low squint angle, it might take a
long while for the optimization to make substantial progress, as the
candidate projections would need to be very close to the optimal one for
the structure of the index function to be visible enough to be amenable
to efficient optimization. This observation suggests that the half-point index value improvement percentage provides an appropriate mathematical definition of squintability, which matches the intuition behind this concept, while being amenable to numerical computation.

To compute the squintability metric \eqref{eq-squintability} in
practice, several approaches are possible. The first one is to propose a
parametric model for $g$, and use it to obtain an explicit formula for
$\varsigma$. Numerical experiments suggest a scaled sigmoid shape as
described below. Define $$
\ell(x):=\frac{1}{1+\exp(\theta_{3}(x-\theta_{2}))}\ ,
$$ {#eq-logistic}

which is a decreasing logistic function depending on two parameters
$\theta_2$ and $\theta_3$, such that $\ell(\theta_{2})=\frac{1}{2}$.
Then, define $$
g(x)=(\theta_{1}-\theta_{4})\frac{\ell(x)-\ell(r_0)}{\ell(0)-\ell(r_0)}+\theta_{4}\ ,
$$ {#eq-parametric}

which depends on three additional parameters, $\theta_1$, $\theta_2$,
and $r_0$, such that $g(0)=\theta_1$ and $g(r_0)=\theta_4$.
Under the parametric model (@eq-parametric), the squintability metric
\eqref{eq-squintability} can be shown to be equal to $$
\varsigma=\frac{g\left(r_{0}/2\right)-\theta_{4}}{\theta_{1}-\theta_{4}}=\frac{\ell(r_0/2)-\ell(r_0)}{\ell(0)-\ell(r_0)}\ .
$$ {#eq-squintability-parametric}

In practice, the parameters of this model (@eq-parametric) can be
estimated numerically, for example by non-linear least squares, and then
used to evaluate $\varsigma$ as in equation
(@eq-squintability-parametric).

Alternatively, one can estimate \eqref{eq-squintability} in a
nonparametric way, for example by fitting $g$ using kernel regression,
then numerically estimate $\varsigma$ from its definition \eqref{eq-squintability}.

# The jellyfish search optimizer {#sec-JSO}

The Jellyfish Search Optimizer (JSO) mimics the natural movements of
jellyfish, which include passive and active motions driven by ocean
currents and their swimming patterns, respectively. In the context of
optimization, these movements are abstracted to explore the search
space, aiming to balance exploration (searching new areas) and
exploitation (focusing on promising areas). The algorithm aims to find
the optimal solution by adapting the behavior of jellyfish to navigate
towards the best solution over iterations [@chou_novel_2021].

To solve the optimization problem embedded in the PP guided tour, a
starting projection, an index function, the number of jellyfish, and the
maximum number of trials (tries) are provided as input. Then, the
current projection is evaluated by the index function. The projection is
then moved in a direction determined by a random factor, influenced by
how far along we are in the optimization process. Occasionally,
completely new directions may be taken like a jellyfish might with ocean
currents. A new projection is accepted if it is an improvement compared
to the current one, rejected otherwise. This process continues and
iteratively improves the projection, until the pre-specified maximum
number of trials is reached.

::: {.callout-note icon="false"}
## Algorithm: Jellyfish Optimizer Pseudo Code

**Input**: `current_projections`, `index_function`, `trial_id`,
`max_trial`

**Output**: `optimized_projection`

**Initialize** `current_best` as the projection with the best index
value from `current_projections`, and `current_idx` as the array of
index values for each projection in `current_projections`

**for** each `trial_id` in 1 to `max_tries` **do**

> Calculate the time control value, $c_t$, based on `current_idx` and
> `max_trial`

> **if** $c_t$ is greater than or equal to $0.5$ **then**
>
> > Define trend based on the `current_best` and `current_projections`
>
> > Update each projection towards the trend using a random factor and
> > orthonormalisation
>
> **else**
>
> > **if** a random number is greater than $1 - c_t$ **then**
> >
> > > Slightly adjust each projection with a small random factor
> > > (passive)
> >
> > **else**
> >
> > > For each projection, compare with a random jellyfish and adjust
> > > towards or away from it (active)
>
> Update the orientation of each projection to maintain consistency
>
> Evaluate the new projections using the index function

> **if** any new projection is worse than the current, revert to the
> `current_projections` for that case
>
> > Determine the projection with the best index value as the new
> > `current_best`

> **if** `trial_id` $\ge$ `max_trial`, print the last best projection
> **exit**

**return** the set of projections with the updated `current_best` as the
`optimized_projection`
:::

<!-- Var names in the function and in the text needs to be consistent??? -->

The JSO implementation involves several key parameters that control its
search process in optimization problems. These parameters are designed
to guide the exploration and exploitation phases of the algorithm. While
the specific implementation details can vary depending on the version of
the algorithm or its application, the focus is on two main parameters
that are most relevant to our application: the number of jellyfish and
the maximum number of tries.

# Assessing the optimizers {#sec-sim-deets}

This section explains the details of two simulation studies: (1) comparison between the existing creeping random search [@RJ-2021-105; @laa_using_2020] and JSO, and (2) examining the factors affecting the performance of JSO.

## Performance of JSO relative to CRS {#sec-app-1}

The CRS is the main optimization routine currently used for the guided tour. Here the two optimizers are compared on the task of finding the pipe using the `holes` index in data with dimensions, $d = 6, 8, 10, 12$. Fifty simulated data sets are used for each dimension. JSO uses 100 jellyfish with a maximum of 100 tries, while the CRS allows a maximum of 1000 tries at each iteration before the algorithm terminates. These choices enable fair comparison between CRS and JSO, that conforms to how they are used in practice.

The performance of the opt misers is measured by the success rate, which is defined as the proportion of simulations that achieves a final index value within 0.05
of the best index value found among all 50 simulations. @fig-success-rate illustrates why the choice of 0.05 is reasonable: the pipe is not recognizable in the projected data. This is motivated by @laa_using_2020's approach to investigating PP indexes.

The results of the simulation are collected using the data structure used in @RJ-2021-105 for assessing PP optimizers. The design parameters are stored along with index value, projection basis, random seed, and computation time.

```{r}
#| label: fig-success-rate
#| fig.cap: "How success rate is calculated, illustrated using the optimal projections from 50 optimizations of 8D pipe data sorted by index value. The pipe shape is recognisable in the projection index values between 0.933-0.969. Of the 50 simulations, 43 achieved an index value within 0.05 of the best, resulting in a success rate of 0.86."
# pipe_jellyfish_12d <- pipe_better |> 
#   filter(dim == 2) |> 
#   mutate(dim = as.numeric(dim) * 8, sim = as.numeric(sim)) 
# proj_dt <- map_dfr(pipe_jellyfish_12d |> pull(basis),
#               ~as_tibble(pipe1000_8d %*% .x), .id = "sim") |>
#   mutate(sim = as.numeric(sim)) |>
#   left_join(pipe_jellyfish_12d |> select(sim, index_val), by = "sim")
# 
# idx_val_sorted <- proj_dt |> pull(index_val) |> unique() |> sort()
# proj_dt |>
#   ggplot(aes(x = V1, y = V2)) +
#   geom_point(size = 0.5) +
#   geom_text(aes(label = sprintf("%.3f", index_val)), x = 0, y = 3.6,
#             size = 5) +
#   xlim(-4, 4) + ylim(-4, 4) +
#   facet_wrap(~fct_reorder(as.factor(sim), -index_val), ncol = 10) +
#   theme_bw() +
#   theme(aspect.ratio = 1, axis.ticks = element_blank(),
#         axis.text = element_blank(), axis.title = element_blank(),
#         panel.grid = element_blank(), strip.background = element_blank(),
#         strip.text = element_blank()) 
knitr::include_graphics(here::here("figures", "success-rate.png"))
```

## Factors affecting JSO success rate: JSO hyper-parameters and index properties {#sec-app-2}

To examine the performance of JSO relative to hyper-parameter choices, the `holes` index is used to find the 2D pipe in data of dimensions 4-12. Additional variables are Gaussian noise. The levels of hyper-parameters are 20, 50, and 100
jellyfish and a maximum of 50 and 100 attempts. Fifty samples of each data set are generated, and the success rate is calculated in the same manner as described in @sec-app-1.

This set of simulations is expanded to assess the performance of JSO relative to the index properties, with a second data set type and additional PP indexes. The second data set has a sine wave in two dimensions and Gaussian noise in the remaining dimensions. Six additional PP indexes that should be able to extract the sine wave are used: `dcor2d_2`, `loess2d`, `MIC`, `TIC`, `spline`, and `stringy`, following @laa_using_2020. (The supplementary material includes definitions of these PP indexes, and descriptions of the pipe and sine wave data simulation.). The JSO hyper-parameters, number of jellies and maximum tries, are also included. This results in a total of `r nrow(sim_df)` scenarios, comprising of `r nrow(filter(sim_df, index == "holes"))` computed on the pipe data and `r nrow(filter(sim_df, index != "holes"))` on the sine wave data. The choice of 4-12 dimensions covers the range of easy to relatively difficult for the optimizers to detect the structured projections. <!-- A list of combinations of the data shape, dimensions and PP indexes that are considered in this study is provided in @tbl-smoothness-squintability.--> The same approach is used, that 50 samples are generated and the success rate is calculated at each level of the simulation design. A success rate of 0 indicates the optimizer did not come sufficiently close to the optimal projection.

```{r}
#| label: fig-smoothness
#| out.width: 100%
#| fig.cap: "Steps for calculating smoothness in a projection pursuit problem. Given the target shape, data dimension and the index function: 1) sample random bases given the orthonormality contraint and calculate their corresponding index values, and 2) fit a Gaussian process model of index values against the sampled bases to obtain the smoothness measure."
# basis_smoothness <- sample_bases(idx = "splines2d", n_basis = 1000)
# basis_smoothness |>
#    ggplot(aes(x = dist, y = index)) +
#    geom_point() + 
#    labs(x = "Projection distance", y = "Index value") + 
#    theme_bw() +   
#   theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30))
knitr::include_graphics(here::here("figures", "smoothness.png"))
```

@fig-smoothness describes the procedure to compute smoothness. Index values are calculated for each of 500 simulated random bases. A Gaussian process model is fitted to the resulting data to obtain the smoothness measure for the index, as described in @sec-smoothness.

```{r}
#| label: fig-squintability
#| out.width: 100%
#| fig.cap: |
#|   Steps for calculating squintability in a projection pursuit problem. Given the target shape, data dimension and the index function: 1) sample random bases given the orthonormality and projection distance contraint and calculates their corresponding index values, 2) interpolate the sampled bases to the optimal basis and calculate the projection distance and the index value. 3) bin the index values by projection distances through averaging index value, 4) fit the scaled sigmoid function in equation \eqref{eq-squintability} to the binned index values against projection distances using non-linear least square to obtian the squintability measure using equation (@eq-squintability-parametric).
# p1 <- sim_sine_6d_spline_projdist |> 
#   ggplot(aes(x = proj_dist, y = index_val)) +
#   geom_point(size = 0.1, alpha = 0.5) + 
#   labs(x = "Projection distance", y = "Index value") + 
#   theme_bw() + 
#   theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30))
# 
# 
# dist_bin <- ceiling(sim_sine_6d_spline_projdist[["proj_dist"]] / 0.1) * 0.1
# dt <- sim_sine_6d_spline_projdist |>
#   dplyr::bind_cols(dist_bin = dist_bin) |>
#   dplyr::group_by(dist_bin) |>
#   dplyr::summarise(index_val = mean(index_val, na.rm = TRUE))
# 
# p2 <-  dt |> 
#   ggplot(aes(x = dist_bin, y = index_val)) + 
#   geom_line() +
#   geom_point() + 
#   labs(x = "Projection distance", y = "Index value") + 
#   theme_bw() + 
#   theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30))
knitr::include_graphics(here::here("figures", "squintability.png"))
```

@fig-squintability describes the procedure to compute squintability. Index values and
projection distances are computed for the projections along interpolated paths between 50 random bases and the optimal basis. An interpolation step size of 0.005 is used. The projection distances are binned to 0.005 intervals and index values are averaged in each bin. The four-parameter scaled logistic function (@eq-parametric) is fitted to the index values against projection distances, estimated by non-linear least squares to obtain the squintability measure as Equation (@eq-squintability-parametric).

A generalized linear model is fitted using a quasibinomial family and a logit link function to assess the factors affecting success rate.  Predictors are smoothness,
squintability, data dimension and JSO hyper-parameters. <!-- The data is
pre-processed by 1) scaling the JSO hyper-parameters by a factor of 10
for interpretation, 2) creating a new binary variable `long_time` to
indicate cases with an average run time over 20 seconds, and 3)
re-coding the success rate for the `stringy` index as 0, because none of
the 50 simulations correctly identified the sine wave shape.-->

# Results {#sec-sim-res}

This section summarises the findings from the simulations that compare the JSO performance with the existing CRS optimizer, and the relationship between optimization success and hyper-parameter choices and PP index properties.

```{r}
#| fig.height: 7.5
#| fig.width: 10
#| label: fig-proj
#| fig.align: "center"
#| fig.cap: "Visual comparison of JSO and CRS results, using optimal data projections obtained over 50 simulations of the pipe data. Rows correspond to data dimension. Columns correspond to quantiles of the index values, with 0 being minimum, 100 being maximum, and 50 the median. JSO achieves the better views of the pipe generally than CRS. As dimension increases both have more difficulty finding the pipe."

aaa <- map_dfr(
  list(pipe1000, pipe1000_8d, pipe1000_10d, pipe1000_12d), 
  ~{
    d_dim = dim(.x)[2]
    pipe_jellyfish |> 
      filter(d == d_dim) |> 
      arrange(index_val) |>
      filter(row_number() %in% c(1, seq(5, 50, 5))) |> 
      mutate(id = factor(
        paste0(c(0, seq(5, 50, 5)) / 50 * 100),
        levels = c(0, seq(5, 50, 5) / 50 * 100))) |> 
      compute_projection(data = .x, col = c("d", "index_val", "id"))
  }
)

p1 <- aaa |> 
  ggplot() + 
  geom_point(aes(x = V1, y = V2), size = 0.05) + 
  geom_text(data = aaa |> select(-V1, -V2) |> unique(), 
            aes(label = sprintf("%.3f", index_val)),
            x = 0, y = 3.7, 
            size = 3) + 
  facet_grid(d ~ id) + 
  xlim(-4, 4) + ylim(-4, 4) +
  theme_bw() +
  theme(aspect.ratio = 1, axis.ticks = element_blank(),
        axis.text = element_blank(), axis.title = element_blank(),
        panel.grid = element_blank())

bbb <- pipe_better |> 
  group_by(dim) |> 
  arrange(index_val) |> 
  filter(row_number() %in% c(1, seq(5, 50, 5))) |>  
  mutate(id = factor(
    paste0(c(0, seq(5, 50, 5)) / 50 * 100),
        levels = c(0, seq(5, 50, 5) / 50 * 100)), 
    d = 4 + 2 * as.numeric(dim)) |> 
  unnest(proj)

p2 <- bbb |> 
  ggplot() + 
  geom_point(aes(x = V1, y = V2), size = 0.05) + 
  geom_text(data = bbb |> select(-V1, -V2) |> unique(), 
            aes(label = sprintf("%.3f", index_val)), x = 0, y = 3.7, 
            size = 3) + 
  facet_grid(d ~ id) + 
  xlim(-4, 4) + ylim(-4, 4) +
  theme_bw() +
  theme(aspect.ratio = 1, axis.ticks = element_blank(),
        axis.text = element_blank(), axis.title = element_blank(),
        panel.grid = element_blank())

(p1 + ggtitle("a. JSO"))/(p2 + ggtitle("b. CRS"))
```

## Performance of JSO relative to CRS

The performance of JSO is assessed relative to the existing CRS optimizer based on the optimal projection obtained. The final projections for the pipe data sets found by the two optimizers are shown in @fig-proj, with rows corresponding to data dimension. The 50 simulations in each data dimension are sorted by index value, and the projections corresponding to each 10th quantile value are shown, with minimum at left and maximum at right. Index value is printed on the plot. The purpose is to summarize the views of the data resulting from different optimization, and hence compare results between the two optimizers. Generally, the JSO does a more consistent job of finding the pipe structure clearly. As the data dimension increases, both optimizers struggle and less clearly capture the circle shape.

<!-- broken down by 10th quantile,
faceted by the data dimensions. In the 6-dimensional data scenario, JSO
consistently identifies a clear pipe shape. The CRS also finds the pipe
shape but with a wide rim, suggesting a further polish search may be
required. With increasing dimensions, JSO may not always identify the
pipe shape due to random sampling, but it still finds the pipe shape in
over 50% of cases. Compared to CRS, JSO achieves higher index values and
clearer pipe shapes across all quantiles in data of 8, 10, 12
dimensions, suggesting its advantage in exploring high-dimensional
spaces.-->

## Effect of hyper-parameters effect on JSO success rate

The effect of JSO hyper-parameters (number of jellyfish and the maximum
number of tries) on the success rate is summarized in @fig-proportion.
The uncertainty is quantified through 500 bootstrap replicates for each
case. As the number of jellyfish and maximum tries increase, the success
rate also increases. For problems with relatively lower dimensional (4, 6) search spaces, small parameter values (20 jellyfish and a maximum of 50 tries) are sufficient to achieve a high success rate. Many more jellyfish (100 jellyfish and a maximum of 100 tries) are needed to obtain a mediocre success rate for dimensions 10 and 12. This suggests that even JSO will struggle with high dimensions, and require substantial more computational time to have an acceptable success rate, especially for difficult PP indexes based on scagnostics. <!-- dimensions) where a more complex search is necessary. However, in lower-dimensional problems, the benefit may be less pronounced, as the space is already relatively easy to navigate. While increasing both parameters enhances the performance of JSO, it also extends the computational time required for the optimization, which can be
computationally intensive when evaluating the index function (such as
scagnostic indexes) multiple times across numerous iterations.-->

```{r}
#| label: fig-proportion
#| fig.width: 8
#| fig.height: 3.5
#| fig.align: "center"
#| fig.cap: "Summary of the relationship between JSO hyper-parameters on optimization success, using the holes index for pipe data of dimensions 4-12. Bootstrap samples show the variability in success rate. Success rate mostly plateaus by 50 jellies. The JSO has some difficulty finding the pipe when dimension is higher than 8! Maximum number of tries has little effect."
pipe_sim_best <- sim_pipe_run_best |> 
  group_by(n_jellies, max_tries, d) |>
  summarise(I_best = max(I_max))

pipe_res <- sim_pipe_run_best |> 
  left_join(pipe_sim_best) |> 
  mutate(diff = abs(I_max - I_best) < 0.05) |> 
  group_by(n_jellies, max_tries, d) |> 
  summarise(proportion = sum(diff) / n()) |> 
  mutate(d = as.factor(d), n_jellies = as.factor(n_jellies))

res_df <- sim_pipe_run_best |> 
  left_join(pipe_sim_best |> ungroup() |> mutate(case_id = row_number()),
            by = c("n_jellies", "max_tries", "d")) |> 
  ungroup() |> 
  group_split(case_id)

resampling <- function(dt, size = size, replace = TRUE){
  row_id <- sample(1:50, size = size, replace = replace)

  dt |> filter(row_number() %in% row_id) |> 
    mutate(diff = abs(I_max - I_best) < 0.05) |> 
    summarise(proportion = sum(diff) / n()) |> 
    pull(proportion)

}

res_interval_df <- tibble(df = res_df) |> 
  crossing(rep = 1:500) |> 
  rowwise() |> 
  mutate(proportion = resampling(df, size = 50))

res_quantile <- res_interval_df |> 
  unnest(df) |> 
  group_by(n_jellies, max_tries, case_id, d) |> 
  select(-I_max, -id) |> 
  distinct() |> 
  mutate(d = as.factor(d),
         n_jellies = ifelse(d == 10, n_jellies - 1, n_jellies),
         n_jellies = ifelse(d == 12, n_jellies + 1, n_jellies)) |> 
  rename(`max # of tries` = max_tries)

res_quantile |> 
  ungroup() |> 
  ggplot(aes(x = n_jellies - 0.5, y = proportion, 
             group = d, color = d)) + 
  geom_segment(aes(xend = n_jellies + 0.5), size = 0.05) + 
  geom_segment(aes(x = n_jellies - 1, xend = n_jellies + 1), 
               data = res_quantile |> summarise(proportion = mean(proportion)), 
               size = 2) + 
  geom_line(aes(x = n_jellies), 
               data = res_quantile |> summarise(proportion = mean(proportion)), 
            alpha = 0.3) + 
  scale_color_brewer(palette = "Dark2", name = "dimension") +
  scale_x_continuous(breaks = c(20, 50, 100), labels = c("20", "50", "100")) +
  facet_wrap(vars(`max # of tries`), 
             labeller = "label_both") + 
  theme_bw() +
  theme(panel.grid.minor = element_blank()) + 
  xlab("Number of jellyfish") + 
  ylab("Proportion of success")
```

## Effect of index properties on JSO success rate

Smoothness and squintability are calculated across a collection of
pipe and sine wave finding problems to construct the
relationship between success rate, JSO hyper-parameters, and index
properties. @tbl-smoothness-squintability presents values computed for a subset of PP indexes on the sine wave data for dimensions 4-8. (A full table of the parameters
estimated from the Gaussian process (outputscale $\eta$, lengthscale
$\ell$, smoothness $\nu$, and nugget $\sigma$) and from the scaled
logistic function ($\theta_1$ to $\theta_4$) for calculating smoothness
and squintability for all PP indexes and data combinations is included in the supplementary materials). The column $\nu$ is
used as the smoothness metric and the column $\varsigma$ is calculated
as equation (@eq-squintability-parametric) as the squintability metric.

@fig-idx-proj-dist shows a sample of traces from the indexes in @tbl-smoothness-squintability for the 6D sine wave data. The traces the PP index values computed for projections along an interpolated path between a noise projection and the optimal projection (sine wave). (Note index values were self-standardized to range between 0 and 1, for comparison.) The `splines` index is very smooth with a gradual increase in value between the noise projection and the optimal projection. The `TIC` index is slightly less smooth and slightly less gradual change in value. Least desirable is the `skinny` index which is noisier and the optimization needs to be much closer to the optimal angle to have a higher value. This illustrates how the new measures for squintability and smoothness work to describe the PP indexes.

```{r}
#| label: tbl-smoothness-squintability
#| tbl-cap: The columns $\nu$ and $\varsigma$ show the smoothness and squintability measures for three PP indexes for the sine wave data. The splines index has the largest squint angle and is the smoothest. The skinny index has low smoothness and smallest squint angle. As dimension increases the squint angle gets smaller but smoothness doesn't change much.
dt <- tibble(shape = c(rep("pipe", 5), rep("sine", 17)), smoothness) |>
  left_join(squintability) |> 
  rename(d = n, smooth = smoothness) |> 
  mutate(index = ifelse(index == "dcor2d_2", "dcor", index),
         index = ifelse(index == "loess2d", "loess", index),
         index = ifelse(index == "splines2d", "splines", index),
         variance = sqrt(variance)) 
dt |> 
  filter(shape == "sine", index %in% c("TIC", "splines", "skinny")) |>
  select(shape, index, d, smooth, squint) |>
  arrange(desc(squint)) |>
  knitr::kable(
    digits = 4, format = "latex", 
    col.names = c(
      "shape", "index", "d",
      #"$\\eta$", "$\\ell$", 
      "$\\nu$", #"$\\sigma$",
      #"$\\theta_1$", "$\\theta_2$", "$\\theta_3$", "$\\theta_4$", 
      "$\\varsigma$"), linesep = "",
    booktabs = TRUE, escape = FALSE, row.names = FALSE, align = "|llr|rr") #align = "|llr|rrrr|rrrrr") |> 
  #column_spec(1, border_left = TRUE) |> 
  #column_spec(c(4, 8, 13), border_right = TRUE) |> 
  #kable_styling(font_size = 10) 
```

```{r}
#| label: fit-logistic
sim_df2 <- sim_df |>   
  mutate(n_jellies = n_jellies/10, max_tries = max_tries/10) |> 
  rename(proportion = P_J) #|>
  #filter(d < 10)

mod1 <- glm(proportion ~ smoothness + squintability + d + n_jellies + max_tries, data = sim_df2, family = quasibinomial) 
mod2 <- glm(proportion ~ squintability + d + n_jellies + max_tries, data = sim_df2, family = quasibinomial)
mod3 <- glm(proportion ~ smoothness + d + n_jellies + max_tries, data = sim_df2, family = quasibinomial)
mod4 <- glm(proportion ~ squintability*d + n_jellies, data = sim_df2, family = quasibinomial)
mod5 <- glm(proportion ~ smoothness*d + squintability*d + n_jellies, data = sim_df2, family = quasibinomial)
```

```{r}
#| label: fig-idx-proj-dist
#| fig.width: 8
#| fig.height: 4
#| fig.cap: "Five traces of index value against projection distance for each of the three indexes (TIC, spline, and skinny) on 6D sine wave data. The traces of the splines index are smooth, with a gradual change between noise and otimal projection. In contrast, the skinny index has a noisy trace that increases rapidly near the optimum. This illustrates the usefulness of the new measures for squintability and smoothness."
sq_basis_df_6d_TIC_spline_skinny |>
  ggplot(aes(x = dist, y = index, group = interaction(index_name, id), color = index_name)) +
  geom_line() + 
  scale_color_brewer(palette = "Dark2") +
  xlab("Projection distance") + ylab("Index value") +
  theme_bw() + 
  theme(legend.position = "bottom")
```

@tbl-mod-output presents the results from fitting a logistic regression
model using the proportion of success as the response and smoothness,
squintability, dimension, number of jellyfish and maximum
number of tries, as predictors. The fit suggests that the JSO success
rate is only affected by squintability, dimension and number of jellies. As expected, the JSO success rate is higher with higher squintability and/or more jellyfish, and lower when dimension is higher (higher-dimensional optimization is more difficult). Interestingly, the JSO is unaffected by the smoothness of the index function. This is consistent with the way random search algorithms jump from value to value without taking local regularity into account, as opposed to gradient-based optimizers for example. Allowing
for more tries also does not affect success rate significantly. A unit increase in squintability increases the success rate by
`r round(100 * (exp(broom::tidy(mod1)$estimate[3])-1), 0)`%. As dimension
increases by one the success rate almost halves. Increasing the number of
jellies by 10 increases the success rate by
`r round(100 * (exp(broom::tidy(mod1)$estimate[5])-1), 0)`%. @tbl-joint-output shows the fit statistics supporting that the model is reasonably strong.

Note that the addition of interaction terms for squintability and data dimension, and smoothness and data dimension, improves the model a small amount, but does not change the interpretation above.

```{r}
#| tbl-pos: H
#| label: tbl-mod-output
#| tbl-cap: "Jellyfish success rate relative to index properties and jellyfish hyper-parameters. This is the summary from a logistic regression fit to smoothness, squintability, dimension, number of jellyfish and maximum number of tries. Interestingly, squintability and dimension strongly affect jellyfish optimization success. The number of jellies marginally affects success, but index smoothness, and increasing the number of tries do not."
#| fig-pos: t
set_gtsummary_theme(theme_gtsummary_journal("jama"))
mod1 |> tbl_regression(exponentiate = TRUE, 
                       pvalue_fun = label_style_pvalue(digits = 3),
                       label = list(smoothness = "Smoothness", 
                                    squintability = "Squintability",
                                    d = "Dimension", 
                                    #long_time = "Long runtime", 
                                    n_jellies = "Number of jellyfish", 
                                    max_tries = "Maximum number of tries")) |>
  bold_p(t = 0.1)
```

```{r}
#| tbl-pos: H
#| label: tbl-joint-output
summary_table <- #bind_rows(
  glance(mod1) |> #%>% mutate(Model = "Full"),
  #glance(mod2) %>% mutate(Model = "Squintability"),
  #glance(mod3) %>% mutate(Model = "Smoothness")
#  glance(mod4) %>% mutate(Model = "Interaction")
#) |>
  select(#Model, 
    deviance, df.residual, null.deviance, df.null, #logLik, 
         ) |> #, AIC, BIC) %>%
  rename(
    `Null Deviance` = null.deviance,
    `DF Null` = df.null,
    `Deviance` = deviance,
    `DF Residual` = df.residual#,
    #`Log-Likelihood` = logLik,
  ) |>
  mutate(across(where(is.numeric), ~ round(.x, 2)))
summary_table |>
  kable(caption = "Fit statistics for the model.") |>
  kable_styling(font_size = 10) |>
  row_spec(0, bold = TRUE) #, width = "5em")
```

<!--The simulation result suggests that squintability and smoothness are likely dependent. Intuitively, a highly smooth search space exhibits gradual changes, making it easier to perceive the overall structure from a distance, thereby increasing the squint angle. @tbl-joint-output compare the model fits when squintability and smoothness are included separately. Given that the null deviance remains the same across models, the model with squintability yields a smaller residual deviance and lower AIC and BIC values, indicating a better fit. This suggests that squintability captures the relevant variation more effectively than smoothness. While it may seem counterintuitive at first that smoothness is not a significant factor, this outcome is reasonable in this context, as the swarm-based method used in this study, JSO, is designed to perform well even in non-smooth search spaces by leveraging population-based exploration.-->

This study intends to serve as an illustration that the proposed measures effectively captures the key factors influencing the success rate of PP. <!--The interactions between the two variables were not examined in this study due to the limited number of data points, nor was the linearity assumption of the relationship with success rate explicitly tested.-->

# Practicalities {#sec-discussion}

The simulation studies have compared the JSO with CRS and shown how
smoothness and squintability affect the JSO success rate. Here we
explain how they can be used for new index and optimizer development.
Using the JSO optimizer for PPGT requires a change in user behavior.
For all the existing optimizers a single optimization path is followed.
The JSO has multiple optimization paths, and needs additional tools to
incorporate into the PPGT, as explained here.

## Using the JSO in a PPGT

To use JSO for PPGT in the `tourr` package, specify
`search_f = search_jellyfish` in the guided tour. Unlike existing
optimizers, the animation function `animate_*()` won't directly render
the tour path for JSO due to the generation of multiple tour paths. To
visualize the path visited by each individual jellyfish, assign the
animation to an object (`res <- animate_xy(...)`). This will save the
bases visited by JSO, along with relevant metadata, as a tibble data
object [see @RJ-2021-105 for more details on the data object]. The bases
visited for individual jellyfish can then be extracted and viewed using
the `planned_tour()` function.

## Computing the index properties for your new index

The `ferrn` package [@RJ-2021-105] provides functionality for computing
the smoothness and squintability metrics. Both metrics require sampling
random bases using the `sample_bases()` function, followed by the metric
calculation with `calc_smoothness()` or `calc_squintability()`.

### Smoothness

To sample bases for calculating smoothness, the following arguments are
required: the index function, the dataset, and the number of random
bases to sample. The output of `sample_bases()` is a data frame with a
list-column of sampled basis matrix and index value. Particularization is
available to speed up the index value computation through the
`parallel = TRUE` argument.

The `calc_smoothness()` function takes the output from `sample_bases()`
and fits a Gaussian process model to the index values against the
sampled basis, as the location, to obtain the smoothness metric.
Starting parameters and additional Gaussian process arguments can be
specified and details of the fit can be accessed through the `fit_res`
attribute of the output.

### Squintability

Bases sampling for squintability includes an additional step of
interpolating between the sampled bases and the best basis. This step is
performed when the arguments `step_size` and `min_proj_dist` in the
`sample_bases()` function are set to non-NA numerical values. Given the
projection distance typically ranging from 0 to 2, it is recommended to
set `step_size` to 0.1 or lower, and `min_proj_dist` to be at least 0.5
to ensure a meaningful interpolation length.

The `calc_squintability()` function computes squintability using two
methods: 1) parametrically, by fitting a scaled sigmoid function through
non-linear least square (`method = "nls"`), and 2) non-parametrically,
using kernel smoothing (`method = "ks"`). A `bin_width` argument is
required to average the index values over the projection distance before
the fit. For the parametric case, the output provides the estimated
parameters for the scaled sigmoid function ($\theta_1$ to $\theta_4$)
and the calculated squintability metric as Equation
(@eq-squintability-parametric). For the non-parametric case, it shows
the maximum gradient attained (`max_d`), the corresponding projection
distance (`max_dist`), and the squintability metric as their products.

# Conclusion {#sec-conclusion}

This paper has presented new metrics to mathematically define desirable
features of PP indexes, squintability and smoothness, and used these to
assess the performance of the new jellyfish search optimizer. The
metrics will be generally useful for characterizing PP indexes, and help
with developing new indexes.

In the comparison of the JSO against the currently used CRS, as expected
the JSO vastly outperforms CRS, and provides a high probability of
finding the global optimum. The JSO obtains the maximum more cleanly,
with a slightly higher index value, and plot of the projected data
showing the structure more clearly.

The JSO performance is affected by the hyper-parameters, with a higher
chance of reaching the global optimum when more jellyfish are used and
the maximum number of tries is increased. However, it comes at a
computational cost, as expected. The performance declines if the
projection dimension increases and if the PP index has low
squintability. The higher the squintability the better chance the JSO
can find the optimum. However, interestingly smoothness does not affect
the JSO performance.

One future direction of this work is to test the JSO, along with its
variations and other swarm-based optimizers, on a broader range of
indexes, for example, scagnostic indexes.

```{=html}
<!--
- Summarise the result of the simulation: JSO immense improvement on current methods, ...
- JSO for a guided tour, what changes are needed? Do we follow one tentacle, do we perform off-line and then pull tentacles out to watch, how does it integrate with ferrn

Main findings:

  - JSO tends to obtian a clearer projection and higher index value when searching higher dimension spaces than CRS (and other current optimizers) - Figure 4

  - Increasing JS hyper-parameter enhances its performance, but the computational time required for the optimization can quickly become intensive, especially when evaluating the index function (such as scagnostic indexes) multiple times across numerous iterations. - Figure 5

  - Smoothness and squintability can be calculated and compared across different problems to understand their relative complexity. Once a projection pursuit problem is fully characterised (by the shape-to-find, the index function used, and the data dimension), increasing JSO hyper-parameters can enhance the search effectiveness, however, the resulting increase in computational complexity will also need to be considered. - modelling
-->
```
# Acknowledgement

The article has been created using Quarto [@Allaire_Quarto_2022] in R
[@R]. The source code for reproducing the work reported in this paper
can be found at: <https://github.com/huizezhang-sherry/paper-jso>. The
simulation data produced in Section 5 can be found at
<https://figshare.com/articles/dataset/Simulated_raw_data/26039506>.
Nicolas Langrené acknowledges the partial support of the Guangdong
Provincial/Zhuhai Key Laboratory IRADS (2022B1212010006) and the UIC
Start-up Research Fund UICR0700041-22.

The R packages used in this work include: `tidyr` [@tidyr], `dplyr` [@dplyr], 
`ggplot2` [@ggplot2], `knitr` [@knitr],`gtsummary` [@gtsummary],
`patchwork` [@patchwork], `ggh4x` [@ggh4x], `broom` [@broom], 
`kableExtra` [@kabelextra], `ferrn` [@RJ-2021-105], and `cassowaryr` [@cassowaryr].

# Supplementary materials {.unnumbered}

The supplementary materials available at
<https://github.com/huizezhang-sherry/paper-jso> include: 1) details of
the indexes used in the simulation study, 2) the script to get started
with using JSO in a PPGT and calculating smoothness and squintability,
as explained in @sec-discussion, and (3) the full code to reproduce the
plots and summaries in this paper.

# References